# Roxbury HOBO and Mod-PM

## STOP

**HAVE YOU RUN `IB_modpm_all_stats` and `IB_hobo_stats`?**
*This file loads dataframes created from running these two scripts. Make sure you run them first (if you haven't already) before running this file.*

This script is used to gain insight into correlation between purifier use and percentage reduction in PMs. 

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Roxbury/IB/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
modpm_df <- read_csv(paste0(path_to_data, "s_IB_M_quants.csv"))
hobo_df <- read_csv(paste0(path_to_data, "s_IB_H_props.csv"))
```
```{r}
merged_df <- modpm_df %>%
  filter(case == "redu", environment == "ratio", type == "pm1", 
         participant_id != "all") %>%
  left_join(hobo_df, join_by(participant_id)) %>%
  drop_na() %>%
  mutate(match_room = case_when(!is_br & room == "living" ~ TRUE,
                                is_br & room == "bedroom" ~ TRUE,
                                participant_id == "303" ~ TRUE,
                                .default = FALSE)) %>%
  filter(match_room) %>%
  filter(!(participant_id == "303" & room == "living"))
```

```{r}
merged_df %>%
  ggplot(aes(x = mean, y = prop_on, label = participant_id)) +
    geom_point() +
    geom_text(vjust =1.5) +
    labs(x = "Mean % Reduction", y = "Proportion of Time Purifier On")
```

