# What are the relationships?
In the `analysis` stage, correlation coefficients were computed for every variable for each case for every participant. This notebook explores what we can learn about the relationship (and lack thereof) between different key variables

```{r}
library(tidyverse)
library(corrr)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/HAFTRAP/OH/"
```

## Summarizing participants
Currently, we have individual correlation coefficients for each participant. Using the merged dataframe from `OH_modpm_cpc_corrs`, we can calculate overall coefficients.

```{r}
# Read overall merged CPC + Mod-PM data
load(paste0(path_to_data, "merged.RData"))

# Read correlation coefficient data
corr_df <- read.csv(paste0(path_to_data, "s_OH_MC_corr.csv"))
```

## Indoor CPC vs. Mod-PM
```{r}
# Compare concent indoor to Mod-PM indoor readings
corr_df %>%
  filter(x == "concent_indoor", str_detect(y, "^pm"), 
         str_detect(y, "indoor$"), participant_id == "all") %>%
  arrange(case)
```
You can see that the correlation between CPC and Mod-PM is slightly higher for sham than for hepa when looking at overall indoor data. Let's investigate this further with scatterplots

## Plot CPC vs. Mod-PM for Indoor
### Constants
Used to map variable to appropriate meaning while plotting
```{r}
# List mapping variable to appropriate meaning
title_map <- list(pm1 = "PM 1", pm25 = "PM 2.5", pm10 = "PM 10", 
                  pm1num = "PM 1 Count", concent = "UFP")

# List mapping variable to appropriate meaning
axes_map <- list(pm1 = "PM 1 Concentration", pm25 = "PM 2.5 Concentration",
               pm10 = "PM 10 Concentration", pm1num = "PM 1 Particle Count",
               concent = "CPC Particle Count")
```

### Plotting functions:
#### Single Line of Best Fit Scatterplot
The following function is a general function that takes any two variables in the provided dataframe, filters for specified case and environment, and scatterplots them with their R-squared value
```{r}
one_line <- function(df, constraint, var_x, var_y, case_str, env_str) {
  # Filter by specified case and environment
  plot_df <- filter(df, case == case_str, environment == env_str)
  
  # Filter by specified constraint (no filtering if overall)
  if (constraint == "Overall") {  } else{
    plot_df <- filter(plot_df, participant_id == constraint)
  }
  
  # Get r-squared value
  r_val <- summary(lm(plot_df[[var_y]] ~ plot_df[[var_x]]))$r.squared
  
  # Scatterplot provided x vs. provided y with line of best fit
  ggplot(plot_df, aes(x = .data[[var_x]], y = .data[[var_y]])) +
    # Scatterplot
    geom_point(alpha = 0.1) +
    # Line of best fit
    geom_smooth(method=lm, se=FALSE) +
    
    # Paste in R-square value
    annotate("text", x = 0.8*max(plot_df[[var_x]]), 
             y = 0.8*max(plot_df[[var_y]]), 
             label = bquote(R^2 == .(round(r_val[[1]], 3)))) +
    
    # Add axes names, title
    labs(title = paste(title_map[[var_y]], "vs.", title_map[[var_x]],
                       "in", constraint, str_to_title(case_str), "Period"), 
         x = axes_map[[var_x]], y = axes_map[[var_y]]) -> my_plot
    
  # Save plot
  ggsave(filename = paste("artifacts/HAFTRAP/OH/scatterplot/OH", constraint,
                           var_x, var_y, case_str, env_str, "one.png", 
                          sep = "_"), plot = my_plot)
  # Display plot
  my_plot
}
```

#### Two lines of best fit Scatterplot (CPC vs. Mod-PM only)
The following function plots a provided Mod-PM reading against CPC particle count at a provided level of constraint. The provided data is filtered and split between two clusters. Two lines of best fit are calculated for each of the clusters.

Note that (by nature of this analysis) this function only only makes sense to be used between CPC particle count and Mod-PM readings - the legend labels reflect that. However, those are the only things that would need to be modified (as well as the cluster splitting equation) in case this function should be used beyond this analysis.
```{r}
# List mapping variable to constant for clustering in two_line
cluster_val <- list(pm1 = 0.002, pm25 = 0.002, pm10 = 0.002, pm1num = 0.002)

two_line <- function(df, constraint, var_x, var_y, case_str, env_str) {
  # Grouped column name
  var_grouped <- paste0(var_y, "_grouped")
  
  # Filter by specified case and environment
  plot_df <- filter(df, case == case_str, environment == env_str) %>%
    # Add column for splitting
    mutate(!!var_grouped := as.factor(
      ifelse(.data[[var_y]] < cluster_val[[var_y]] * .data[[var_x]], 
             "Mod-PM Missed", "Mod-PM + CPC Detected")))
  
  # Filter by specified constraint (no filtering if overall)
  if (constraint == "Overall") {  } else{
    plot_df <- filter(plot_df, participant_id == constraint)
  }
  
  # Set model
  my_model <- paste(var_y, "~", var_x)
  
  # Calculate r-squared, proportion of points
  props <- plot_df %>% 
    # For each group
    group_by(across(last_col())) %>% 
    # Calculate proportion of points, r-squared
    summarize(proportion = 100 * n() / nrow(plot_df),
              rsq = summary(lm(eval(parse(text = my_model))))$r.squared) %>%
    # Format appropriately for annotation
    mutate(proportion = paste0("(", round(proportion, 1), "%)"), 
           rsq = round(rsq, 3))
  
  # Recode values to include proportion in factor, pipe into plotting
  plot_df %>%
    mutate((!!var_grouped) := 
             recode(.data[[var_grouped]], 
                    "Mod-PM Missed" = 
                      paste("Mod-PM Missed", props$proportion[2]),
                    "Mod-PM + CPC Detected" = 
                      paste("Mod-PM + CPC Detected", props$proportion[1]))) %>%
  
  # Scatterplot with two lines of best fit for separate clusters
  ggplot(aes(x = .data[[var_x]], y = .data[[var_y]], 
             color = .data[[var_grouped]])) +
    # Scatter-plot
    geom_point(alpha = 0.1) +
    # Lines of best fit
    geom_smooth(method=lm, se=FALSE) +
    # Paste in R-square value for Mod-PM + CPC Detected
    annotate("text", x = 0.5*max(plot_df[[var_x]]), 
             y = 0.8*max(plot_df[[var_y]]), 
               label = bquote(R^2 == .(props$rsq[1]))) +
    # Paste in R-square value for Mod-PM Missed
    annotate("text", x = 0.8*max(plot_df[[var_x]]), 
             y = 0.2*max(plot_df[[var_y]]), 
               label = bquote(R^2 == .(props$rsq[2]))) +
    # Add axes labels, title
    labs(title = paste("Split", title_map[[var_y]], "vs.", title_map[[var_x]],
                       "in", constraint, str_to_title(case_str), "Period"), 
         x = axes_map[[var_x]], y = axes_map[[var_y]], 
         color = "Group") -> my_plot
    
  # Save plot
  ggsave(filename = paste("artifacts/HAFTRAP/OH/scatterplot/OH", constraint,
                           var_x, var_y, case_str, env_str, "two.png", 
                          sep = "_"), plot = my_plot)
  # Display plot
  my_plot
}
```

#### Time-series plot to compare two variables
Plots two time-series overlaying each other to compare two different measurements for a particular participant in a specified case. Usually used to compare CPC (UFP Particle Count) to a Mod-PM reading. If using for a different set of variables, change the scaling constant (sc) appropriately (e.g.- sc is ~1 for comparison between Mod-PM variables).
```{r}
series_compare <- function(df, person, var_a, var_b, case_str, env_str, sc){
  # Filter dataframe
  df %>% 
    filter(participant_id == person, case == case_str, 
                     environment == env_str) %>%
  # Plot time-series of UFP with PM 2.5
  ggplot(aes(x = date)) + 
    # Create second (right) y-axis with 200 times the scale of first
    scale_y_continuous(sec.axis = sec_axis(~.*sc, name = axes_map[[var_a]])) +
    # Plot CPC reading against right axis
    geom_line(aes(y = .data[[var_a]]/sc, color = title_map[[var_a]])) +
    # Plot PM 2.5 reading against left axis
    geom_line(aes(y = .data[[var_b]], color = title_map[[var_b]])) +
    
    # Set colors
    scale_color_manual(name = "Particles", 
                       values = setNames(c('#56B4E9', '#E69F00'),
                                         c(title_map[[var_a]],
                                           title_map[[var_b]]))) +
    
    # Set axes labels, title
    labs(title = paste("CPC vs. Mod-PM Readings for Participant", person), 
         x = "Date", y = axes_map[[var_b]]) -> my_plot
    
  # Save plot
  ggsave(filename = paste("artifacts/HAFTRAP/OH/scatterplot/OH", person,
                           var_a, var_b, case_str, env_str, "timeseries.png",
                          sep = "_"), plot = my_plot)
  # Display plot
  my_plot
}
```




CPC Particle Count Indoor vs. PM 2.5 Concentration Indoor (sham)
```{r}
for (case in c("sham", "hepa")) {
  for (pm_type in c("pm1", "pm25", "pm10", "pm1num")) {
    # Plot scatterplot with line of best fit
    print(one_line(main_df, "Overall", "concent", pm_type, case, "indoor"))
    print(paste("Plotted", pm_type, case))
  }
}
```

```{r}
for (case in c("sham", "hepa")) {
  for (pm_type in c("pm1", "pm25", "pm10", "pm1num")) {
    # Plot scatterplot with line of best fit
    print(two_line(main_df, "Overall", "concent", pm_type, case, "indoor"))
    print(paste("Plotted", pm_type, case))
  }
}
```


### Zoom into single participant
Looking at the same data but for a single participant so that a time-series comparison can also be plotted
```{r}
# Looking into PM 2.5 for Participant 42281
one_line(main_df, "42281", "concent", "pm25", "sham", "indoor")
two_line(main_df, "42281", "concent", "pm25", "sham", "indoor")
```

Plotting a time-series for that participant in that situation provides us with additional context on where the readings agreed and where they did not (see helper function series_compare)
```{r}
series_compare(main_df, "42281", "concent", "pm25", "sham", "indoor", 200)
```

