# Sham vs. HEPA - is there a difference?
This notebook investigates whether the installation of air purifiers had a notable impact on indoor air quality across various summary metrics. It uses summary data found in `summary/HAFTRAP/OH`

```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
path_to_data <- "summary/HAFTRAP/OH/"
```

## Plotting Function
```{r}
plot_heatmap <- function(joined_data, title) {
  # Sets order for displaying particles in heatmap
  particle_order <- c("PM 1 Conc.", "PM 2.5 Conc.", "PM 10 Conc.", 
                     "PM 1 Count", "UFP Count")
  # Sets order for displaying summary stats in heatmap
  summary_order <- c("Mean", "95th Percentile", "75th Percentile", "Median",
                     "25th Percentile", "5th Percentile")
  
  # Pre-process data before ploting
  joined_data %>%
    # Reshape for heatmap
    pivot_longer(mean:q95, names_to = "sum_type", values_to = "stat") %>%
    
    # Recode for accurate labels
    mutate(type = recode(type, "pm1" = "PM 1 Conc.", "pm25" = "PM 2.5 Conc.", 
                     "pm10" = "PM 10 Conc.", "pm1num" = "PM 1 Count",
                     "concent" = "UFP Count"),
         sum_type = recode(sum_type, "mean" = "Mean", "median" = "Median",
                           "q5" = "5th Percentile", "q25" = "25th Percentile",
                           "q75" = "75th Percentile", 
                           "q95" = "95th Percentile")) %>%
    
    # Arrange axes in specified order
    mutate(type = factor(type, levels = particle_order)) %>%
    mutate(sum_type = factor(sum_type, levels = summary_order)) %>%

    # Plot heatmap
    ggplot(aes(x = type, y = sum_type)) +
      # Fill with value of stat
      geom_tile(aes(fill = stat)) +
      # Print in value of stat
      geom_text(aes(label = paste0(round(stat, 1), "%"))) +
    
      # Set color scale
      scale_fill_gradient2(low = "#ff7f7f", high = "#DeF7E9", mid = "white") +
      # Set axis labels, title
      labs(title = title, x = "Particle Type", y = "Summary Statistic", 
           fill = "Reduction (%)") +
      # Set theme
      theme_minimal()
}
```

## Load Data
### CPC
```{r}
filepath <- paste0(path_to_data, "s_OH_C_quants.csv")
cpc_stats <- read_csv(filepath, show_col_types = FALSE)
```

### Mod-PM
```{r}
filepath <- paste0(path_to_data, "s_OH_M_quants.csv")
modpm_stats <- read_csv(filepath, show_col_types = FALSE)
```

### Reduction in raw Indoor Concentrations
```{r}
joined_indoor <- rbind(modpm_stats, cpc_stats) %>%
  filter(participant_id == "all", environment == "indoor", case == "redu") %>%
  filter(type %in% c("pm1", "pm25", "pm10", "pm1num", "concent"))
```

### Reduction in Indoor/Outdoor Ratio
```{r}
# Calculate CPC, Mod-PM averages for I/O ratio
joined_ratio <- rbind(modpm_stats, cpc_stats) %>%
  filter(participant_id == "all", environment == "ratio", case == "redu") %>%
  filter(type %in% c("pm1", "pm25", "pm10", "pm1num", "concent"))
```

## Making pretty tables
```{r}
plot_heatmap(joined_indoor, "Reduction in Raw Indoor Readings")
plot_heatmap(joined_ratio, "Reduction in Indoor/Outdoor Ratio")
```



## Save to file
Save the two tables with efficacy in CSV files
```{r}
# Save summary statistic files to CSVs
write.csv(joined_indoor, "artifacts/indoor_reduction.csv")
write.csv(joined_ratio, "artifacts/ratio_reduction.csv")
```

