# Create bar plot
This script uses summary statistic data to create a bar plot.

## STOP
Have you run `Community_All_modpm_stats.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
summary_df <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
```

```{r}
var_type <- "pm1"
stat_order <- c("q5", "q25", "median", "q75", "q95")
summary_df %>% filter(environment == "indoor", type == var_type, case == "redu",
                      participant_id == "all") %>%
  pivot_longer(cols = mean:sd, names_to = "stat") %>%
  filter(room_type != "all", stat %in% c("q5", "q25", "median", "q75", "q95", "mean")) %>%
  mutate(stat = fct_relevel(stat, "q5", "q25", "median", "q75", "q95")) %>%
  ggplot() + geom_col(aes(x = room_type, y = value, fill = stat), position = "dodge") +
  labs(x = "Deployment Class", y = paste("Percentage Reduction", var_type), fill = "Summary Statistic") +
  scale_fill_discrete(labels = c("5th Percentile", "25th Percentile", "Median", 
                                 "75th Percentile", "95th Percentile", "Mean"))
```

```{r}
var_type <- "pm1"

plot_df <- summary_df %>% 
  filter(environment == "indoor", type == var_type, case == "redu", 
         room_type == "all") %>%
  pivot_longer(cols = mean:sd, names_to = "stat") %>%
  filter(stat %in% c("q5", "q25", "median", "q75", "q95", "mean")) %>%
  mutate(stat = fct_relevel(stat, "q5", "q25", "median", "q75", "q95"))

mean_redu <- plot_df %>% filter(stat == "mean") %>% select(value) %>% as.double()
plot_df %>%
  filter(stat != "mean") %>%
  ggplot() + 
    geom_col(aes(x = stat, y = value), position = "dodge", fill = "#c3e7fd") +
    labs(x = "Summary Statistic Across All Deployments", 
         y = bquote("Percentage Reduction in"~PM[1])) +
    scale_x_discrete(labels = c("5th Percentile", "25th Percentile", "Median", 
                                   "75th Percentile", "95th Percentile")) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +
    geom_hline(yintercept = mean_redu, linetype = "dashed") + 
    geom_text(label = "Mean", x = "q95", y = mean_redu, vjust = -0.5) +
    theme_bw()

```

```{r}
var_type <- "pm1"
stat_order <- c("q5", "q25", "median", "q75", "q95")
summary_df %>% 
  filter(environment %in% c("indoor", "ratio"), type == var_type, participant_id == "all",
         room_type != "all", case %in% c("off", "on")) %>%
  ggplot() + 
  geom_col(aes(x = room_type, y = mean, fill = case), position = "dodge") +
  labs(x = "Deployment Class", y = paste("Mean", var_type, "Comparison"), 
       fill = "Purifier State") +
  facet_wrap(~environment, nrow = 2, scales = "free_y") +
  theme_bw()
```
```{r}
off_points <- 5 + 1*scale(rnorm(100))

off_quants <- quantile(off_points, c(0.05, 0.25, 0.5, 0.75, 0.95))

on_points <- 2 + scale(rnorm(100))

on_quants <- quantile(on_points, c(0.05, 0.25, 0.5, 0.75, 0.95))

reduction <- (off_quants - on_quants)*100/ off_quants

reduction %>%
  enframe() %>%
  mutate(name = factor(name, levels = c("5%", "25%", "50%", "75%", "95%"))) %>%
ggplot(aes(x = name, y = value)) +
  geom_bar(stat = "identity", fill = "#c3e7fd") +
  labs(x = "Quantile", y = "Percentage Reduction")
```

