# Create violin plot
This script uses summary statistic data to create a violin plot.

## STOP
Have you run `Community_All_modpm_stats.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "all.RData"))
summary_df <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
```


```{r}
current_redu <- summary_df %>% 
  filter(environment == "indoor", type == "pm25", case == "redu", 
         participant_id == "all", room_type != "all")

mean_plot <- summary_df %>% 
  filter(environment == "indoor", type == "pm25", case != "redu", 
         participant_id == "all", room_type != "all")

plot_df <- all_df %>% filter(environment == "indoor")

d_width <- position_dodge(width = 0.8)

violin_range <- quantile(all_df$pm25, c(0.05, 0.95), na.rm = TRUE)
                         
plot_df %>%
  ggplot(aes(x = room_type, y = pm25)) +
  geom_violin(aes(fill = case), position = d_width) +
  geom_point(data = mean_plot, aes(y = mean, fill = case, size = "Mean"), 
             position = d_width) +
  scale_y_continuous(limits = violin_range) +
  labs(fill = "Purifier State", x = "Room Type", y = bquote(PM[2.5])) +
  scale_fill_discrete(labels = c("Off", "On")) +
  geom_text(aes(y = violin_range[2], label = paste0(round(mean, 1), "%")), 
                data = current_redu, fontface = "bold") +
  scale_size_manual(name = "Statistic", values = c("Mean" = 2)) +
  guides(size = guide_legend(title = "Statistic"))
  
```

```{r}
current_redu <- summary_df %>% 
  filter(environment == "ratio", type == "pm25", case == "redu", 
         participant_id == "all", room_type != "all")

mean_plot <- summary_df %>% 
  filter(environment == "ratio", type == "pm25", case != "redu", 
         participant_id == "all", room_type != "all")

plot_df <- all_df %>% filter(environment == "ratio")

d_width <- position_dodge(width = 0.8)

violin_range <- quantile(all_df$pm25, c(0.05, 0.95), na.rm = TRUE)
                         
plot_df %>%
  ggplot(aes(x = room_type, y = pm25)) +
  geom_violin(aes(fill = case), position = d_width) +
  geom_point(data = mean_plot, aes(y = mean, fill = case, size = "Mean"), 
             position = d_width) +
  scale_y_continuous(limits = violin_range, trans = "log10") +
  labs(fill = "Purifier State", x = "Room Type", y = bquote(PM[2.5])) +
  scale_fill_discrete(labels = c("Off", "On")) +
  geom_text(aes(y = violin_range[2], label = paste0(round(mean, 1), "%")), 
                data = current_redu, fontface = "bold") +
  scale_size_manual(name = "Statistic", values = c("Mean" = 2)) +
  guides(size = guide_legend(title = "Statistic"))
  
```