# Create box plot
This script uses summary statistic data to create a box and whisker plot.

## STOP
Have you run `Community_All_modpm_stats.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "all.RData"))
summary_df <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
```


```{r}
current_redu <- summary_df %>% 
  filter(environment == "indoor", type == "pm1", case == "redu", 
         participant_id == "all", room_type != "all")

plot_df <- all_df %>% filter(environment == "indoor")

whisker_range <- quantile(plot_df$pm1, c(0.01, 0.99), na.rm = TRUE)

plot_df %>%
  ggplot(aes(x = room_type, y = pm1)) +
  geom_boxplot(aes(fill = case), outlier.shape = NA) +
  scale_y_continuous(limits = whisker_range, trans = "log10") +
  labs(fill = "Purifier State", x = "Room Type", y = bquote(PM[1])) +
  scale_fill_discrete(labels = c("Off", "On")) +
  geom_text(aes(y = whisker_range[2], label = paste0(round(mean, 1), "%")), 
                data = current_redu, fontface = "bold") +
  theme_bw()
  
```

```{r}
current_redu <- summary_df %>% 
  filter(environment == "ratio", type == "pm1", case == "redu", 
         participant_id == "all", room_type != "all")

plot_df <- all_df %>% filter(environment == "ratio")

whisker_range <- quantile(plot_df$pm1, c(0.01, 0.99), na.rm = TRUE)

plot_df %>% 
  ggplot(aes(x = room_type, y = pm1)) +
  geom_boxplot(aes(fill = case), outlier.shape = NA) +
  scale_y_continuous(limits = whisker_range, trans = "log10") +
  labs(fill = "Purifier State", x = "Room Type", y = bquote(PM[1])) +
  scale_fill_discrete(labels = c("Off", "On")) +
  geom_text(aes(y = whisker_range[2], label = paste0(round(mean, 1), "%")), 
                data = current_redu, fontface = "bold") +
  theme_bw()
  
```


```{r}
format_environment <- function(df) {
  indoor_form <- deparse("Indoor Conc."~PM[1](mu~"g/cm"^3))
  ratio_form <- deparse("Indoor-Outdoor Ratio of"~PM[1])
  df %>% 
    mutate(environment_formatted = case_when(
      environment == "indoor" ~ indoor_form,
      environment == "ratio" ~ ratio_form)) %>%
    mutate(environment_formatted = factor(environment_formatted, 
                                          levels = c(indoor_form, ratio_form)))
}
```

```{r}
current_redu <- summary_df %>% 
  filter(environment != "outdoor", type == "pm1", case == "redu", 
         participant_id == "all", room_type != "all") %>%
  format_environment()

plot_df <- all_df %>% filter(environment != "outdoor") %>% format_environment()

mean_plot <- summary_df %>% 
  filter(environment != "outdoor", type == "pm1", case != "redu", 
         participant_id == "all", room_type != "all") %>% format_environment()

d_width <- position_dodge(width = 0.8)

num_participants <- summary_df %>% 
  filter(environment != "outdoor", type == "pm1", case == "redu", 
         participant_id != "all", room_type != "all") %>%
  count(room_type, environment) %>% format_environment()

whisker_range <- quantile(plot_df$pm1, c(0.01, 0.99), na.rm = TRUE)

plot_df %>%
  ggplot(aes(x = room_type, y = pm1)) +
  geom_boxplot(aes(fill = case), outlier.shape = NA) +
  scale_y_continuous(limits = whisker_range, trans = "log10") +
  labs(fill = "Purifier State", x = "Room Type", y = bquote(PM[1])) +
  scale_fill_discrete(labels = c("Off", "On")) +
  geom_text(aes(y = whisker_range[2], label = paste0(round(mean, 1), "%")), 
                data = current_redu, fontface = "bold") +
  geom_text(aes(y = whisker_range[1], label = paste("n =", n)), 
                data = num_participants, fontface = "bold", vjust = 0.1) +
  geom_point(data = mean_plot, aes(y = mean, fill = case, size = "Mean"), 
             position = d_width) -> my_plot
```

```{r}
my_plot +
  facet_wrap(~environment_formatted, nrow = 2, scales = "free_y", 
             strip.position = "left", labeller = label_parsed) +
  ylab(NULL) +
  theme_bw() +
  theme(strip.background = element_blank(),
       strip.placement = "outside") +
  guides(size = guide_legend(title=NULL)) +
  annotate("segment", x = 2.5, xend = 2.2, y = whisker_range[2] / 2, 
           yend = whisker_range[2]*0.9, arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", x = 2.5, y = whisker_range[2] / 2, 
           label = "Mean Reduction", hjust = "left", size = 3.2)
```

