# Plotting corrplots

## STOP

Have you run `Community_All_modpm-cpc_corrs.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "mod_cpc.RData"))
```

```{r}
get_corr <- function(df, place) {
  df %>% 
    filter(place_id == place, environment == "indoor") %>%
    select(-room) %>%
    drop_na() %>%
    select(where(is.numeric)) %>%
    cor()
}
```


```{r}
# Calculate correlations within each group
for (place_id in c("CH", "RH", "SS", "LF")) {
  get_corr(modcpc_df, place_id) %>%
    corrplot(method = "number", type = "upper", diag = FALSE, 
             title = paste("Correlation for", place_id), mar=c(0,0,1,0))
}
```

```{r}
modcpc_df %>% 
    filter(environment == "indoor") %>%
    select(-room) %>%
    drop_na() %>%
    select(where(is.numeric)) %>%
    cor() %>%
    corrplot(method = "number", type = "upper", diag = FALSE, 
             title = paste("Correlation overall"), mar=c(0,0,1,0))
```

