# Correlations - further digging
This script digs deeper into the correlation coefficients between measurements taken by the Mod-PM and CPC sensors.

```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

## Summarizing participants
Currently, we have individual correlation coefficients for each participant. Using the merged dataframe from `OH_modpm_cpc_corrs`, we can calculate overall coefficients.

```{r}
# Read overall merged CPC + Mod-PM data
load(paste0(path_to_data, "mod_cpc.RData"))

# Read correlation coefficient data
corr_df <- read.csv(paste0(path_to_data, "s_AllwCPC_MC_corrs.csv"))
```

## Comparing coefficients
Trying to understand which Mod-PM variables are most closely related with CPC, and in which deployments.
```{r}
corr_df %>%
  filter(x == "concent" | y == "concent") %>%
  mutate(mod_pm_var = ifelse(x == "concent", y, x)) %>%
  ggplot(aes(x = r, y = mod_pm_var, colour = mod_pm_var, label = participant_id)) +
    geom_jitter(height = 0.2) +
    xlim(0, 1) +
    theme_classic()
```



```{r}
corr_df %>%
  filter(x == "concent" | y == "concent") %>%
  mutate(mod_pm_var = ifelse(x == "concent", y, x)) %>%
  ggplot(aes(x = r, y = mod_pm_var, colour = place_id, label = participant_id)) +
    geom_jitter(height = 0.2) +
    xlim(0, 1) +
    theme_classic()
```
```{r}
corr_df %>%
  filter(x == "concent" | y == "concent") %>%
  mutate(mod_pm_var = ifelse(x == "concent", y, x)) %>%
  ggplot(aes(x = r, y = mod_pm_var, colour = case, label = participant_id)) +
    geom_jitter(height = 0.2) +
    xlim(0, 1) +
    theme_classic()
```

```{r}

```

