# Plotting corrplots

## STOP

Have you run `Community_All_modpm-cpc_corrs.Rmd`, `Community_All_modpm_stats.Rmd` and `Community_All_cpc_stats.Rmd`?
*This file loads a dataframe and two CSVs created from running the above script. Make sure you run them first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
library(corrplot)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "mod_cpc.RData"))
```

```{r}
get_corr <- function(df, place) {
  df %>% 
    filter(place_id == place, environment == "indoor") %>%
    select(-room) %>%
    drop_na() %>%
    select(where(is.numeric)) %>%
    cor()
}
```


```{r}
# Calculate correlations within each group
for (place_id in c("CH", "RH", "SS", "LF")) {
  get_corr(modcpc_df, place_id) %>%
    corrplot(method = "number", type = "upper", diag = FALSE, 
             title = paste("Correlation for", place_id), mar=c(0,0,1,0))
}
```

```{r}
modcpc_df %>% 
    filter(environment == "indoor") %>%
    select(-room) %>%
    drop_na() %>%
    select(where(is.numeric)) %>%
    cor() -> corr_plot_df
```

```{r}
rownames(corr_plot_df) <- c(":PM[1]", ":PM[2.5]", ":PM[10]", ":PM[1] (Count)", "UFP Count")

colnames(corr_plot_df) <- c(":PM[1]", ":PM[2.5]", ":PM[10]", ":PM[1] (Count)", "UFP Count")

corr_plot_df %>%
    corrplot(method = "number", type = "upper", diag = FALSE, 
             title = paste("Mod-PM to CPC Correlation"), mar=c(0,0,1,0))
```

## Part 2: Percentage reductions comparison
```{r}
modpm_summary <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
cpc_summary <- read_csv(paste0(path_to_data, "s_All_C_quants.csv"))
```

```{r}
merged_summary <- left_join(cpc_summary, modpm_summary,
                            by = c("area", "room_type", "participant_id", 
                                   "place_id", "environment", "case"),
                            relationship = "one-to-many",
                            suffix = c("_cpc", "_modpm")) %>%
  select(where(is.character), starts_with("mean")) %>%
  filter(case == "redu")
```

```{r}
merged_summary %>%
  filter(environment == "ratio", type_modpm == "pm1", area != "all") %>%
  ggplot(aes(x = mean_modpm, y = mean_cpc)) +
    geom_point(aes(color = place_id), size = 3) +
    lims(x = c(-50, 100), y = c(-50, 100)) +
    geom_hline(yintercept = 0, linetype = "dotted") +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  annotate("segment", x = 50, xend = 75, y = 90, yend = 75, 
           arrow = arrow(length = unit(0.03, "npc"))) +
  annotate("segment", x = 50, xend = 60, y = 90, yend = 60, 
           arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", x = 50, y = 95, 
           label = "Mod-PM, CPC Results Match", size = 3.2) +
    scale_color_discrete("Deployment Location",
                         labels = c(
                           "CH" = "Revere City Hall",
                           "LF" = "Little Folks Daycare",
                           "RH" = "Revere High School",
                           "SS" = "Shining Star Daycare"
                         )) +
    theme_bw() +
    labs(x = bquote("Mean % Reduction in "~PM[1]),
         y = "Mean % Reduction in UFP Count",
         title = "Comparison of Efficacy (Measured by I/O Ratio)")

```

```{r}
modpm_summary %>% 
  filter(case == "redu") %>% 
  pivot_wider(names_from = type, values_from = mean, 
              id_cols = c(participant_id, room_type, area, place_id, environment)) %>%
  filter(environment == "ratio", area != "all") %>%
  ggplot(aes(x = pm1, y = pm25)) +
    geom_point() +
    lims(x = c(-100, 100), y = c(-100, 100)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed")

```

