# Calculating proportion of PM1

## STOP
Have you run `Community_All_modpm_stats.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "all.RData"))
```

```{r}
proportion_pm1 <- function(df, grouping_vars, env_vars, case_vars) {
  df %>%
    filter(environment %in% env_vars, case %in% case_vars) %>%
    group_by(across(all_of(grouping_vars))) %>%
    mutate(pm1_to_pm25 = pm25 - pm1, pm25_to_pm10 = pm10 - pm25) %>%
    summarize(pm1_prop = mean(pm1/pm10, na.rm = TRUE),
              pm1_to_pm25_prop = mean(pm1_to_pm25/pm10, na.rm = TRUE), 
              pm25_to_pm10_prop = mean(pm25_to_pm10/pm10, na.rm = TRUE)) %>%
    pivot_longer(pm1_prop:pm25_to_pm10_prop, 
                 names_to = "size_bin", values_to = "proportion") %>%
    drop_na()
}
```

```{r}
sum_prop <- function(df) {
  df %>% 
    filter(size_bin == "pm1_prop") %>% 
    summarize(pm1_prop = mean(proportion))
}
```

```{r}
plot_label <- function(df) {
  geom_text(data = df %>% filter(size_bin == "pm1_prop"), 
              aes(y = 0.8, label = round(proportion, 3)))
}
```


```{r}
plot_df <- all_df %>%
  proportion_pm1(c("case", "place_id"), 
               env_vars = c("indoor"), case_vars = c("off", "on"))

plot_df %>%
  ggplot(aes(x = place_id, y = proportion, fill = size_bin)) +
    geom_bar(stat = "identity") +
    facet_grid(vars(case)) +
    plot_label(plot_df) +
    scale_fill_discrete(
      labels = c("< 1 Micron", "1 - 2.5 Microns", "2.5 - 10 Microns")) +
    labs(fill = "Particle Diameter") +
    theme_bw()

sum_prop(plot_df)
```

```{r}
plot_df <- all_df %>%
  proportion_pm1(c("environment", "place_id"), 
               env_vars = c("indoor", "outdoor"), case_vars = c("off", "on", NA))

plot_df %>%
  ggplot(aes(x = place_id, y = proportion, fill = size_bin)) +
    geom_bar(stat = "identity") +
    facet_grid(vars(environment)) +
    plot_label(plot_df) +
    scale_fill_discrete(
      labels = c("< 1 Micron", "1 - 2.5 Microns", "2.5 - 10 Microns")) +
    labs(fill = "Particle Diameter") +
    theme_bw()

sum_prop(plot_df)
```
```{r}
plot_df <- all_df %>%
  proportion_pm1(c("case", "room_type"), 
               env_vars = c("indoor"), case_vars = c("off", "on"))

plot_df %>%
  ggplot(aes(x = room_type, y = proportion, fill = size_bin)) +
    geom_bar(stat = "identity") +
    facet_grid(vars(case)) +
    plot_label(plot_df) +
    scale_fill_discrete(
      labels = c("< 1 Micron", "1 - 2.5 Microns", "2.5 - 10 Microns")) +
    labs(fill = "Particle Diameter") +
    theme_bw()

sum_prop(plot_df)
```
```{r}
plot_df <- all_df %>%
  proportion_pm1(c("environment", "room_type"), 
               env_vars = c("indoor", "outdoor"), case_vars = c("off", "on", NA))

plot_df %>%
  ggplot(aes(x = room_type, y = proportion, fill = size_bin)) +
    geom_bar(stat = "identity") +
    facet_grid(vars(environment)) +
    plot_label(plot_df) +
    scale_fill_discrete(
      labels = c("< 1 Micron", "1 - 2.5 Microns", "2.5 - 10 Microns")) +
    labs(fill = "Particle Diameter") +
    theme_bw()

sum_prop(plot_df)
```

