# All HOBO and Mod-PM

## STOP

**HAVE YOU RUN `Community_All_modpm_all_stats` and `Community_All_hobo_stats`?**
*This file loads dataframes created from running these two scripts. Make sure you run them first (if you haven't already) before running this file.*

This script is used to gain insight into correlation between purifier use and percentage reduction in PMs. 

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
modpm_df <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
hobo_df <- read_csv(paste0(path_to_data, "s_All_H_props.csv"))
```
## Merge summary statistics
```{r}
merged_df <- modpm_df %>%
  filter(case == "redu", environment == "indoor", type == "pm1", 
         participant_id != "all") %>%
  right_join(hobo_df, join_by(area, place_id, participant_id)) %>%
  filter(participant_id != "all")
```

```{r}
merged_df %>%
  ggplot(aes(x = prop_on, y = mean, label = participant_id, color = place_id)) +
    geom_point() +
    # geom_text(hjust = 0.5, vjust = -0.5) +
    labs(y = bquote("Mean % Reduction in "~PM[1]~"(I/O Ratio)"), 
         x = "Proportion of Time Purifier On",
         title = "Relating Purifier Usage to Efficacy") +
    ylim(-50, 100) +
    geom_hline(yintercept = 0) +
    scale_color_discrete("Deployment Location",
                         labels = c(
                           "CM" = "Cardinal Medeiros",
                           "IB" = "Indigo Block",
                           "MF" = "Mothers Out Front"
                         )) +
    theme_bw()
```


# Part 2: Isolating times when purifier is really on
```{r}
# Load Mod-PM Data
load(paste0(path_to_data, "all.RData"))

# Load HOBO Data
load(paste0(path_to_data, "all_hobo.RData"))
```

## Merge overall Mod-PM and HOBO Data
```{r}
all_df %>%
  filter(case == "on", environment == "indoor") %>%
  inner_join(on_df, join_by(date, area, place_id, participant_id, room, case)) %>%
  count(place_id)
```

