# Visualize the different percentiles

## STOP
Have you run `Community_All_modpm_stats.Rmd`?
*This file loads a dataframe created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
load(paste0(path_to_data, "all.RData"))
summary_df <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
```
```{r}
summary_df %>%
  filter(place_id == "IB", participant_id == "105", 
         environment == "indoor", type == "pm1", case != "redu") -> sum_stats

sum_stats %>% 
  filter(case == "off") -> ss_off
```



```{r}
all_df %>% 
  filter(place_id == "IB", participant_id == "105", 
         environment == "indoor") %>% 
  left_join(sum_stats) %>%
  mutate(quartile = case_when(pm1 < q5 ~ "< 5th",
                   pm1 > q5 & pm1 < q25 ~ "5th to 25th",
                   pm1 > q25 & pm1 < median ~ "25th to Median",
                   pm1 > median & pm1 < q75 ~ "Median to 75th",
                   pm1 > q75 & pm1 < q95 ~ "75th to 95th",
                   .default = "> 95th")) %>%
  ggplot(aes(x = date, y = pm1, color = quartile)) +
    geom_point(size = 0.1) +
    scale_color_manual(values = c("#9656A2", "#369ACC", "#95CF92", "#F8E16F", 
                                   "#F4895F", "#DE324C"),
                         breaks = c("< 5th", "5th to 25th", "25th to Median", 
                                    "Median to 75th", "75th to 95th", "> 95th")) +
    geom_vline(xintercept = strptime("2022-10-25 23:16:00", "%Y-%m-%d %H:%M:%S", 
                                     tz = "America/New_York"),
               linetype = "dashed", size = 0.2) +
    guides(color = guide_legend(override.aes = list(size = 2))) +
    theme_bw() -> current_plot

current_plot
```
```{r}
current_plot +
    ylim(c(0, ss_off$q95 * 3)) +
    labs(y = bquote(PM[1]~(mu~"g/cm"^3)), x = "Date", color = "Quantile Bin",
         title = "Visualizing Quantiles for Sample Deployment") +
    annotate("segment", 
             x = as.POSIXct(strptime("2022-10-27 21:00:00", "%Y-%m-%d %H:%M:%S", 
                          tz = "America/New_York")), 
             xend = as.POSIXct(strptime("2022-10-26 01:00:00", "%Y-%m-%d %H:%M:%S", 
                             tz = "America/New_York")), 
             y = 18.5, yend = 18.5, arrow = arrow(length = unit(0.02, "npc"))) +
  
  annotate(geom = "text", 
           x = as.POSIXct(strptime("2022-10-27 22:00:00", "%Y-%m-%d %H:%M:%S", 
                          tz = "America/New_York")), 
           y = 18.6, label = "Purifier Installed", hjust = "left")
```

```{r}
plot_df <- all_df %>% 
  filter(place_id == "IB", participant_id == "105", case == "off",
         environment == "indoor")

summary_df %>%
  filter(place_id == "IB", participant_id == "105", 
         environment == "indoor", type == "pm1", case == "off") -> ss_off

zone_names <- c("< 5th", "5th to 25th", "25th to Median", 
                       "Median to 75th", "75th to 95th")
rect_df <- ss_off %>% 
  pivot_longer(cols = c(q5, q25, median, q75, q95), 
                        names_to = "stat", values_to = "y_max") %>%
  mutate(y_min = c(0, head(y_max, -1)), 
         x_min = min(plot_df$date), x_max = max(plot_df$date),
         zone_name = zone_names)
```


```{r}
plot_df %>% 
  ggplot() +
    geom_line(aes(x = date, y = pm1), size = 0.1) +
    geom_hline(yintercept = c(ss_off$q5, ss_off$q25, ss_off$median, ss_off$q75,
                              ss_off$q95),
               linetype = "dashed", size = 0.2) +
    geom_rect(data = rect_df, aes(xmin = x_min, xmax = x_max, 
                                  ymin = y_min, ymax = y_max, fill = zone_name),
              alpha = 0.2) +
    scale_fill_discrete(limits = zone_names) + 
    labs(fill = "Percentile Range") +
    theme_bw() -> current_plot

current_plot
```
```{r}
current_plot +
    ylim(c(0, ss_off$q95 * 3))
```


