# Normalizing for activity

## STOP

Have you run `Community_All_modpm_stats.Rmd`?
*This file loads the CSV created from running the above script. Make sure you run it first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/Community_All/"
```

```{r}
# Load Mod-PM summary statistics
modpm_summary <- read_csv(paste0(path_to_data, "s_All_M_quants.csv"))
```

```{r}
# Load plume statistics summary
plume_summary <- read_csv(paste0(path_to_data, "participant_plume_summary.csv"))
```
## Process data
Clean summary and plume data to prepare for merging
```{r}
# Select relevant columns and filter rows from summary stats
modpm_clean <- modpm_summary %>%
  filter(place_id != "all", environment == "indoor", type == "pm1",
         case %in% c("off", "on")) %>%
  select(participant_id, place_id, case, mean)

# Select relevant columns and filter rows from plume
plume_clean <- plume_summary %>%
  filter(pm == "pm1") %>%
  select(deployment, participant_id, hepa, n) %>%
  mutate(case = case_when(hepa ~ "on", !hepa ~ "off"), .keep = "unused")
```

```{r}
stats_plumes <- inner_join(modpm_clean, plume_clean, 
          by = join_by(participant_id, case, place_id == deployment)) %>%
  mutate(mean.per.peak = mean/n)
```

```{r}
stats_plumes %>%
  pivot_longer(contains("mean"), names_to = "mean_type", values_to = "mean_value") %>%
  pivot_wider(id_cols = c(participant_id, place_id, mean_type), 
              names_from = case, values_from = mean_value) -> stats_plumes_reshaped

stats_plumes_reshaped %>%
  mutate(is_effective = off > on) %>%
  group_by(mean_type) %>%
  summarize(prop_effective = round(mean(is_effective), 3)) -> props

stats_plumes_reshaped %>%
  ggplot(aes(x = off, y = on)) +
    geom_point() +
    geom_text(data = props, mapping = aes(label = prop_effective), x = 1, y = 5, 
              hjust = "left") +
    geom_abline(slope = 1, intercept =  0, linetype = "dashed") +
    facet_wrap(~mean_type, scales = "free")
```

### Calculate percentage reduction in concentration from off to on HEPA installation
```{r}
# Spread: Create separate off, on summary statistics columns
summary_wide <- pivot_wider(stats_plumes, names_from = case, 
                            values_from = mean:mean.per.peak)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("off")) - 
  select(summary_wide, ends_with("on"))) / 
  select(summary_wide, ends_with("off"))

# Rename computed columns to 'redu' for percentage reduction
colnames(summary_redu) <- sub("off", "redu", colnames(summary_redu))

# Reshape data and pipe into variable 'summary_normal'
summary_wide %>%
  # Merge back with joined data-frame
  cbind(summary_redu) %>%
  # Gather: Return to original shape by removing the 'sham/hepa/redu'
  # suffixes by re-forming the case variable
  pivot_longer(mean_off:mean.per.peak_redu, 
               names_to = c(".value", "case"), names_sep = "_") %>%
  # Filter out all NaN and Inf values
  filter_all(all_vars(!is.infinite(.))) -> summary_normal


```

```{r}
summary_normal %>%
  filter(case == "redu") %>%
  ggplot(aes(x = mean, y = mean.per.peak)) +
    geom_point() +
    lims(x = c(-50, 100), y = c(-50, 100))
```

