# HAFTRAP Data Cleaning and Summaries - multiple (for CPC)
This script is used to calculated summary statistics for multiple participants in the HAFTRAP study, for CPC deployments by Olin. It is largely a copy of `OH_modpm_all_stats.Rmd`, but with few modifications to deal with CPC data. Ensure that each participant has indoor and outdoor data (NOT distinguished by sham and HEPA) and that all files are labelled appropriately.

Results can be found in the folder`summary/HAFTRAP/OH/cpc/`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrplot)
library(rstatix)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/vkuchhal/Documents/hepa-summer23')
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Vector of participant numbers' data to be processed
participants <- c("41271", "42281", "45361", "44621", "45451")

# Read summary table of times
time_map <- read_excel(
  paste0("data/HAFTRAP/OH/", "OH_N_summarized.xlsx")) %>%
  # Convert to appropriate variable types
  mutate(across(dt_start:dt_end, 
                  ~as.POSIXct(.x, format = "%m/%d/%Y %H:%M"))) %>%
  mutate(across(participant_id:case, as.factor))
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/cpc/"

# Set data category
data_cat <- "OH_C"
```

## Helper Functions
The following functions are not part of the main pipeline to calculate summary statistics, but are used to clean and process the data -

### Filter by relevant times
Each participant contains data for periods where both the sham and actual purifier was running. This function filters the relevant time periods and adds an added variable to track sham and HEPA (actual) readings.
```{r}
include_case <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # For sham periods
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start[cd$case == "sham"], 
              cd$dt_end[cd$case == "sham"]) ~ 'sham',
      # For HEPA periods
      between(date, cd$dt_start[cd$case == "hepa"], 
              cd$dt_end[cd$case == "hepa"]) ~ 'hepa'))
}
```



### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables. This function calculates the correlation matrix for a particular case for a single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, p, c) {
  print(paste("Correlation processed for", p, c))
  df %>%
    # Filter for case
    filter(participant_id == p, case == c) %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations
    cor_mat() %>%
    # Reshape into paired-list format
    cor_gather()
}
```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each environment
  for (env in c("indoor", "outdoor")) {
    # Set file path
    # Filename = data category + participant ID + environment
    file_name <- paste(data_cat, person, env, sep = "_")
    
    # File path = path to folder, type csv
    file_path <- paste0(path_to_data, file_name, ".csv")
    
    # Read csv
    df <- read_csv(file_path, show_col_types = FALSE) %>%
      # Filter to first four columns (for funky data from preprocessing)
      select(1:4) %>%
      # Add the environment type (e.g. indoor) and participant ID
      mutate(environment = env, participant_id = person)
    
    # Append to main dataframe
    all_df <- rbind(all_df, df)
  }
}

```

```{r}
# Warning: this code chunk may seem simple but takes FOREVER.
# Have patience. It deals with datetimes.
clean_df <- all_df %>%
  # Merge the date and time column
  unite(col = "date", `#YY/MM/DD`, `HR:MN:SC`, sep = " ") %>%
  # Convert to datetime object
  mutate(date = as.POSIXct(date, format = "%y/%m/%d %H:%M:%S", 
                           tz = "America/New_York")) %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, .keep_all = TRUE)
```



```{r}
# Filter by time to find whether sham or hepa
case_df <- data.frame()
for (person in participants) {
  # Filter indoor (dates match outdoor) and by current participant
  current_dates <- time_map %>% 
    filter(environment == "indoor") %>%
    filter(participant_id == person)
  # Filter by date for sham and hepa (see helper function)
  df <- include_case(clean_df, person, current_dates)
  # Append to overall dataframe
  case_df <- rbind(case_df, df)
}
```

```{r}
# Quick cleaning:
# Remove values that aren't in relevant time range, make factors
main_df <- case_df %>%
  drop_na() %>%
  mutate(across(c(environment, case, participant_id), as.factor)) 
  
```


### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that are then divided by each other to calculate the ratio. After calculating the ratio, the dataframe is 'gathered' back up to its original shape.
```{r}
# Spread: Create separate indoor and outdoor columns for measurements
main_df_wide <- pivot_wider(main_df,
                  names_from = environment,
                  values_from = c(concent, inlttmp))

# Divide indoor by outdoor to calculate ratios
df_ratio <- select(main_df_wide, ends_with("indoor"))/
  select(main_df_wide, ends_with("outdoor"))

# Rename computed columns to ratio
colnames(df_ratio) <- sub("indoor", "ratio", colnames(df_ratio))

# Merge back with joined data-frame
main_df_wide <- cbind(main_df_wide, df_ratio)

# Gather: Return to original shape by re-forming the environment variable
main_df <- pivot_longer(main_df_wide, 
                               concent_indoor:inlttmp_ratio, 
                               names_to = c(".value", "environment"), 
                               names_sep = "_") %>%
  mutate(environment = as.factor(environment))
```


### Calculate correlation matrices
Helps better understand the relationship between different variables in sham and true HEPA conditions for each participant. Matrices are reshaped into longer pair format for stacking all the data into single dataframe. (Warning: this code takes a while to run.)

```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()
# Loop through each participant
for (person in participants) {
  # Loop through each case
  for (case in c("sham", "hepa")) {
    # Use wide dataframe where indoor/outdoor are separate columns
    main_df_wide %>%
      # Get correlation coefficients (look at helper function)
      get_corr(person, case) %>%
      # Add the case and participant ID
      mutate(case = case, participant_id = person) -> df
    # Append to main dataframe
    corr_df <- rbind(corr_df, df)
  }
}
```


### Gather, group, and summarize
To calculate summary statistics of the data, the different measurements are gathered into one variable 'reading'. The resulting long dataframe is cleaned for NaN values and grouped. Numerous summary statistics are calculated for each type for each, case, environment, and participant.

```{r}
main_df %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(concent:inlttmp, 
               names_to = "type", values_to = "reading") %>%
  # Drop NaN, inf values for summary calculation
  filter(is.finite(reading)) %>%
  # Group by participant ID, case, environment, and type
  group_by(participant_id, case, environment, type) %>%
  
  # Calculate summary statistics and pipe to variable 'summary'
  summarise(mean = mean(reading),
            median = median(reading), 
            q5 = quantile(reading, probs = 0.05), 
            q25 = quantile(reading, probs = 0.25),
            q75 = quantile(reading, probs = 0.75),
            q95 = quantile(reading, probs = 0.95),
            sd = sd(reading),
            .groups = 'drop') -> summary
```

### Calculate percentage reduction in concentration from sham to hepa
Follows the same logic as calculating indoor-outdoor ratios. Spread to form sham/hepa columns that are used to calculate percent reduction in corresponding measurements. Then gather dataframe back to original shape.
```{r}
# Spread: Create separate sham, hepa summary statistics columns
summary_wide <- pivot_wider(summary, 
                            names_from = case, values_from = mean:sd)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("sham")) - 
  select(summary_wide, ends_with("hepa"))) / 
  select(summary_wide, ends_with("sham"))

# Rename computed columns to 'redu' for percentage reduction
colnames(summary_redu) <- sub("sham", "redu", colnames(summary_redu))

# Reshape data and pipe into variable 'summary_normal'
summary_wide %>%
  # Merge back with joined data-frame
  cbind(summary_redu) %>%
  # Gather: Return to original shape by removing the 'sham/hepa/redu'
  # suffixes by re-forming the case variable
  pivot_longer(mean_hepa:sd_redu, 
               names_to = c(".value", "case"), names_sep = "_") %>%
  # Filter out all NaN and Inf values
  filter_all(all_vars(!is.infinite(.))) %>%
  # Change to factors
  mutate(across(c(type, case), as.factor)) -> summary_normal
```

## Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(summary_normal, "summary/HAFTRAP/OH/cpc/s_OH_C_quants.csv")
write.csv(corr_df, "summary/HAFTRAP/OH/cpc/s_OH_C_corr.csv")
```