# HAFTRAP Data Cleaning and Summaries - multiple (for CPC)
This script is an initial builder to generate figures for multiple participants in the HAFTRAP study, using deployments by Olin. It is largely a copy of `OH_modpm_all_stats.Rmd`, but modified to deal with CPC data. Ensure that each participant has indoor and outdoor data, all files labelled appropriately.


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(openair)
library(corrplot)
library(ggplot2)
library(ggpmisc)
library(rstatix)
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}

# # If file path doesn't end with the root folder "hepa-summer23", set to that folder on your PC
# setwd(dir)
```

```{r}
# Vector of participant numbers' data to be processed
participants <- c("41271")

# Table of participant numbers and pilot start-end datetimes
time_map <- read_csv(paste0("data/HAFTRAP/OH/", "OH_T.csv")) %>%
  mutate(across(!participant_id, as.POSIXct)) %>%
  mutate(participant_id = factor(participant_id))


```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/cpc/"

# Set data category
data_cat <- "OH_C"
```

## Helper Functions
The following functions are not part of the main pipeline to calculate summary
statistics, but are used to clean and process the data


### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables.
This function calculates the correlation matrix for a particular case for a
single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, p, c) {
  df %>%
    # Filter for case
    filter(participant_id == p, case == c) %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations
    cor_mat() %>%
    # Reshape into paired-list format
    cor_gather()
}
```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each environment
  for (env in c("indoor", "outdoor")) {
    # Set file path
    # Filename = data category + participant ID + environment
    file_name <- paste(data_cat, person, env, sep = "_")
    
    # File path = path to folder, type csv
    file_path <- paste0(path_to_data, file_name, ".csv")
    
    # Read csv
    df <- read_csv(file_path, show_col_types = FALSE)
    
    # Add the environment type (e.g. indoor) and participant ID
    df <- mutate(df, environment = env, participant_id = person)
    
    # Append to main dataframe
    all_df <- rbind(all_df, df)
  }
}

```

```{r}
main_df <- all_df %>%
  unite(col = "date", `#YY/MM/DD`, `HR:MN:SC`, sep = " ") %>%
  mutate(date = as.POSIXct(date, format = "%y/%m/%d %H:%M:%S", tz = "America/New_York")) %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, .keep_all = TRUE)
```

```{r}

main_df %>%
  mutate(across(c(participant_id, environment), as.factor))
  
  
```

### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that
are then divided by each other to calculate the ratio. After calculating the
ratio, the dataframe is 'gathered' back up to its original shape.
```{r}
# Spread: Create separate indoor and outdoor columns for particle concentration
main_df_wide <- pivot_wider(main_df,
                  names_from = environment,
                  values_from = c(concent, inlttmp))

# Divide indoor by outdoor to calculate ratios
df_ratio <- select(main_df_wide, ends_with("indoor"))/
  select(main_df_wide, ends_with("outdoor"))

# Rename computed columns to ratio
colnames(df_ratio) <- sub("indoor", "ratio", colnames(df_ratio))

# Merge back with joined data-frame
main_df_wide <- cbind(main_df_wide, df_ratio)

# Gather: Return to original shape by removing the 'indoor'/'outdoor'/'ratio'
# suffixes by re-forming the environment variable
main_df <- pivot_longer(main_df_wide, 
                               concent_indoor:inlttmp_ratio, 
                               names_to = c(".value", "environment"), 
                               names_sep = "_")
```


### Calculate correlation matrices
Helps better understand the relationship between different variables for each participant. Matrices are reshaped into longer pair format for stacking all the 
data into single dataframe.

```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()
# Loop through each participant
for (person in participants) {
  # Use wide dataframe where indoor/outdoor are separate columns
  main_df_wide %>%
    # Get correlation coefficients (look at helper functions)
    get_corr(person) %>%
    # Add the case and participant ID
    mutate(participant_id = person) -> df
  # Append to main dataframe
  corr_df <- rbind(corr_df, df)
}
```


### Gather, group, and summarize
To calculate summary statistics of the data, the data are gathered into one variable 'reading'. The resulting long dataframe is cleaned for NaN values and grouped. Numerous summary statistics are calculated for each type for each, case, environment, and participant.

```{r}
main_df %>%
  # Gather dataframe to create single variable for particle concentrations
  pivot_longer(concent:inlttmp, names_to = "type", values_to = "reading") %>%
  # Drop NaN, inf values for summary calculation
  filter(is.finite(reading)) %>%
  # Group by participant ID, case, environment, and particle type
  group_by(participant_id, environment, type) %>%
  
  # Calculate summary statistics and pipe to variable 'summary'
  summarise(mean = mean(reading),
            median = median(reading), 
            q5 = quantile(reading, probs = 0.05), 
            q25 = quantile(reading, probs = 0.25),
            q75 = quantile(reading, probs = 0.75),
            q95 = quantile(reading, probs = 0.95),
            sd = sd(reading),
            .groups = 'drop') -> summary
```


## Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(summary_normal, "summary/HAFTRAP/OH/modpm/s_OH_M_quants.csv")
write.csv(corr_df, "summary/HAFTRAP/OH/modpm/s_OH_M_corr.csv")
```