FOR TODAY:
1) Refactor code for gathering/spreading
2) Deal with correlation coefficients, find a way to make it summarizable
3) Test with 2+ files (talk to Sanju for download)



# HAFTRAP Data Cleaning and Summaries - multiple
This script is an initial builder to generate figures for multiple participants in the HAFTRAP study, using deployments by Olin. It is largely a copy of `OH_modpm_single_stats.Rmd`, but scaled up to deal with multiple participants. Ensure that each participant has indoor and outdoor data for both sham and HEPA stages, all files labelled appropriately.


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(openair)
library(corrplot)
library(ggplot2)
library(ggpmisc)
library(rstatix)
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}

# # If file path doesn't end with the root folder "hepa-summer23", set to that folder on your PC
# setwd(dir)
```

```{r}
# Vector of participant numbers' data to be processed
participants <- c("41181")
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/modpm/"

# Set data category
data_cat <- "OH_M"
```

## Helper Functions
The following functions are not part of the main pipeline to calculate summary
statistics, but are used to clean and process the data

### Reformatting timestamps
Necessary since the timestamps from QuantAQ are in a weird format and need to be formatted to a format that works with the openair plotting package. An added column rounding the times to the nearest minute is calculated since two data-frames will be merged later in this script on the basis of time.

```{r}
# Function to reformat time-stamps to time object and round to nearest minute
improve_timestamps <- function(df) {
  df %>%
    # Reformat timestamps to sensible format
    # mutate(date = as.POSIXct(strptime(timestamp_local, 
    #   format = "%Y-%m-%dT%H:%M:%SZ", tz = "America/New_York"))) %>%
    mutate(date = as.POSIXct(timestamp, tz = "America/New_York")) %>%
    # Round times to nearest minute
    mutate(date = round_date(date, unit = "minute"))
}
```

### Particle counts and removing unnecessary data
Alongside particle masses, we are concerned with the counts of particles that fall under certain size bins. To approximate the counts under pm1, sum up bin 0 to 2. Discard all other data columns that are not relevant.
```{r}
# Function to calculate sums of particle counts, remove rest
sum_bins <- function(df) {
  df %>%
    # Sum particle counts
    mutate(pm1num = bin0 + bin1 + bin2) %>%
    # Delete unnecessary columns
    select(pm1:pm1num) %>%
    select(-ends_with("_model_id"))
}

```

### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables.
This function calculates the correlation matrix for a particular case for a
single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, participant_id, case) {
  df %>%
    # Filter for case
    filter(participant_id == participant_id, case == case) %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations
    cor_mat() %>%
    # Reshape into paired-list format
    cor_gather()
}
```

### Load all data. Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
main_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each case
  for (case in c("sham", "hepa")) {
    # Loop through each environment
    for (env in c("indoor", "outdoor")) {
      
      # Set file path
      # Filename = data category + participant ID + case 
      # + environment
      file_name <- paste(data_cat, person, case, env, sep = "_")
      
      # File path = path to folder, type csv
      file_path <- paste0(path_to_data, file_name, ".csv")
      
      # Read csv
      df <- read_csv(file_path, show_col_types = FALSE)
      
      # Add the case and environment type (e.g. sham, indoor)
      # And participant ID
      df <- mutate(df, case = case, environment = env,
              participant_id = person)
      
      # Append to main dataframe
      main_df <- rbind(main_df, df)
    }
  }
}

```

### Process all data to remove unnecessary columns and round time values 
This helps in syncing data across different sensors later - functions here can
be found in the 'Helper Functions' section.
```{r}
# Clean data (look at helper functions)
main_df <- main_df %>%
  # Reformat timestamps and round to nearest minute
  improve_timestamps() %>%

  # Calculate particle counts and remove unnecessary columns
  sum_bins()
```

### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that
are then divided by each other to calculate the ratio. After calculating the
ratio, the dataframe is 'gathered' back up to its original shape.
```{r}
# Spread: Create separate indoor and outdoor columns for particle concentration
main_df_wide <- pivot_wider(main_df,
                  names_from = environment,
                  values_from = c(pm1, pm25, pm10, pm1num))

# Divide indoor by outdoor to calculate ratios
df_ratio <- select(main_df_wide, ends_with("indoor"))/
  select(main_df_wide, ends_with("outdoor"))

# Rename computed columns to ratio
colnames(df_ratio) <- sub("indoor", "ratio", colnames(df_ratio))

# Merge back with joined data-frame
main_df_wide <- cbind(main_df_wide, df_ratio)

# Gather: Return to original shape by removing the 'indoor'/'outdoor'/'ratio'
# suffixes by re-forming the environment variable
main_df <- pivot_longer(main_df_wide, 
                               pm1_indoor:pm1num_ratio, 
                               names_to = c(".value", "environment"), 
                               names_sep = "_")
```


### Calculate correlation matrices
Helps better understand the relationship between different variables in sham and
true HEPA conditions for each participant. Matrices are reshaped into longer pair
format for stacking all the data into single dataframe.

```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()
# Loop through each participant
for (person in participants) {
  # Loop through each case
  for (case in c("sham", "hepa")) {
    # Use wide dataframe where indoor/outdoor are separate columns
    main_df_wide %>%
      # Get correlation coefficients (look at helper functions)
      get_corr(person, case) %>%
      # Add the case and participant ID
      mutate(case = case, participant_id = person) -> df
    # Append to main dataframe
    corr_df <- rbind(corr_df, df)
  }
}
```


### Gather, group, and summarize
To calculate summary statistics of the data, the different particle types (PM1, 
PM2.5, PM10, PM1_num) are gathered into one variable 'particle'. The resulting
long dataframe is cleaned for NaN values and grouped. Numerous summary
statistics are calculated for each particle type for each, case, environment,
and participant.

```{r}
main_df %>%
  # Gather dataframe to create single variable for particle concentrations
  pivot_longer(pm1:pm1num, names_to = "particle", values_to = "conc") %>%
  # Drop NaN values for summary calculation
  drop_na() %>%
  # Group by participant ID, case, environment, and particle type
  group_by(participant_id, case, environment, particle) %>%
  
  # Calculate summary statistics and pipe to variable 'summary'
  summarise(mean = mean(conc),
              median = median(conc), 
              q5 = quantile(conc, probs = 0.05), 
              q25 = quantile(conc, probs = 0.25),
              q75 = quantile(conc, probs = 0.75),
              q95 = quantile(conc, probs = 0.95),
              .groups = 'drop') -> summary
```


### Calculate percentage reduction in concentration from sham to hepa
Follows the same logic as calculating indoor-outdoor ratios. Spread to form 
sham/hepa columns that are used to calculate percent reduction in corresponding
summary pollution concentrations. Then gather dataframe back to original shape.
```{r}
# Spread: Create separate sham, hepa summary statistics columns
summary_wide <- pivot_wider(summary, names_from = case, values_from = mean:q95)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("sham")) - 
  select(summary_wide, ends_with("hepa"))) / select(summary_wide, ends_with("sham"))

# Rename computed columns to 'redu' for percentage reduction
colnames(summary_redu) <- sub("sham", "redu", colnames(summary_redu))

# Merge back with joined data-frame
summary_wide <- cbind(summary_wide, summary_redu)

# Gather: Return to original shape by removing the 'sham/hepa/redu'
# suffixes by re-forming the case variable
summary_normal <- pivot_longer(summary_wide, mean_hepa:q95_redu, 
                               names_to = c(".value", "case"), 
                               names_sep = "_")
```

### Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(summary_normal, "summary/HAFTRAP/OH/modpm/s_OH_M_quants.csv")
write.csv(corr_df, "summary/HAFTRAP/OH/modpm/s_OH_M_corr.csv")
```