# East Boston Data Correlation Calculations

## STOP

**HAVE YOU RUN `LF_modpm-cpc_cleaning`?**
*This file loads the dataframe created from running `LF_modpm-cpc_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to calculate correlation coefficients for multiple participants in the East Boston pilot, for Mod-PM and CPC deployments by Olin.

Results can be found in `summary/EastBoston/LF/s_LF_MC_corr.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/LF/"
```

## Load data
```{r}
# Load Mod-PM + CPC data from RData file
load(paste0(path_to_data, "cleaned_modpm-cpc.RData"))
```


### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables. This function calculates the correlation matrix for a particular case for a single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, case, person, environment) {
  df %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations (suppress output message)
    correlate(quiet = TRUE) %>%
    # Reshape into paired-list format
    stretch(remove.dups = TRUE) %>%
    # Add the case and participant ID
    mutate(case = case, participant_id = person, environment = environment) %>%
    # Drop NaNs
    drop_na()
}
```

### Calculate correlation matrices
Helps better understand the relationship between different variables in sham and true HEPA conditions for each participant. Matrices are reshaped into longer pair format for stacking all the data into single dataframe. (Warning: this code takes a while to run.)
```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()

# List of participants with both Mod-PM and CPC
participants <- unique(main_df$participant_id)

# Loop through each case
for (c in c("off", "on")) {
  # Loop through each environment
  for (env in c("indoor")) {
    # Loop through each participant
    for (participant in participants) {
      # Use wide dataframe where indoor/outdoor are separate columns
      main_df %>%
      # Filter for case
      filter(case == c, environment == env, participant_id == participant) %>%
      # Get correlation coefficients (look at helper function)
      get_corr(case = c, environment = env, person = participant) -> df
      
      # Append to main dataframe
      corr_df <- rbind(corr_df, df)
      print(paste("Computed correlation coefficients for",participant, c, env))
    }
    # Repeat for overall correlation coefficient
    main_df %>%
    # Filter for case
    filter(case == c, environment == env) %>%
    # Get correlation coefficients (look at helper function)
    get_corr(case = c, environment = env, person = "all") -> df
    
    # Append to main dataframe
    corr_df <- rbind(corr_df, df)
    print(paste("Computed correlation coefficients for all", c, env))
  }
  
}
```
## Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(corr_df, "summary/EastBoston/LF/s_LF_MC_corr.csv")
```
