# East Boston PM1 against CPC Count

## STOP

**HAVE YOU RUN `LF_modpm-cpc_cleaning`?**
*This file loads the dataframe created from running `LF_modpm-cpc_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

## Set up
Load libraries, define file paths
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/LF/"
```

## Load data
```{r}
# Load Mod-PM + CPC data from RData file
load(paste0(path_to_data, "cleaned_modpm-cpc.RData"))
```

### Constants
Used to map variable to appropriate meaning while plotting
```{r}
# List mapping variable to appropriate meaning
title_map <- list(pm1 = "PM 1", pm25 = "PM 2.5", pm10 = "PM 10", 
                  pm1num = "PM 1 Count", concent = "UFP")

# List mapping variable to appropriate meaning
axes_map <- list(pm1 = "PM 1 Concentration", pm25 = "PM 2.5 Concentration",
               pm10 = "PM 10 Concentration", pm1num = "PM 1 Particle Count",
               concent = "CPC Particle Count")
```

#### Time-series plot to compare two variables
Plots two time-series overlaying each other to compare two different measurements for a particular participant in a specified case. Usually used to compare CPC (UFP Particle Count) to a Mod-PM reading. If using for a different set of variables, change the scaling constant (sc) appropriately (e.g.- sc is ~1 for comparison between Mod-PM variables).
```{r}
series_compare <- function(df, person, var_a, var_b, case_str, env_str, sc){
  # Filter dataframe
  df %>% 
    filter(participant_id == person, case == case_str, 
                     environment == env_str) %>%
  # Plot time-series of UFP with PM 2.5
  ggplot(aes(x = date)) + 
    # Create second (right) y-axis with 200 times the scale of first
    scale_y_continuous(sec.axis = sec_axis(~.*sc, name = axes_map[[var_a]])) +
    # Plot CPC reading against right axis
    geom_line(aes(y = .data[[var_a]]/sc, color = title_map[[var_a]])) +
    # Plot PM 2.5 reading against left axis
    geom_line(aes(y = .data[[var_b]], color = title_map[[var_b]])) +
    
    # Set colors
    scale_color_manual(name = "Particles", 
                       values = setNames(c('#56B4E9', '#E69F00'),
                                         c(title_map[[var_a]],
                                           title_map[[var_b]]))) +
    
    # Set axes labels, title
    labs(title = paste("CPC vs. Mod-PM Readings for Participant", person), 
         x = "Date", y = axes_map[[var_b]]) -> my_plot
    
  # Save plot
  ggsave(filename = paste("artifacts/HAFTRAP/OH/scatterplot/OH", person,
                           var_a, var_b, case_str, env_str, "timeseries.png",
                          sep = "_"), plot = my_plot)
  # Display plot
  my_plot
}
```

```{r}
main_df %>%
  pivot_longer(c(pm1num, concent), names_to = "p_type", values_to = "value") %>%
  select(!pm1:pm10num) -> select_df

pm1num_df <- select_df %>% filter(p_type == "pm1num")

concent_df <- select_df %>% filter(p_type == "concent")
```


```{r}
main_df %>%
  filter(day(date) %in% c(26, 27)) %>%
  series_compare("class2", "concent", "pm1num", "off", "indoor", 700)
```


