---
editor_options: 
  markdown: 
    wrap: 72
---

# Community All Mod-PM Plume Viewer

Plot any particular plume you would like from any dataset!

To use this, you must make sure to import the correct data from whichever
deployment you want to look at. You also need to know the beginning and ending
indices for the plume so you can plot it. All of this information can be found
for each plume in the combined DataFrame created in
`Community_All_modpm_plume_plotting`.

```{r}
# import necessary libraries
library(tidyverse)
library(data.table)
```

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = "~/AirPartners/hepa-summer23")
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# SET PATH TO DATA HERE
path_to_data <- "cleaning/Revere/CH/"
```

## Load data

```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Main Code Run

```{r}
# Pre-processing to add relevant columns
my_df <- main_df %>%
  # Round date to nearest ten minutes
  mutate(date_round = round_date(date, unit = "10 mins")) %>%
  # Get day of the week as integer from 1 to 7
  mutate(wkdy = wday(date)) %>%
  # Classify as weekday or weekend
  mutate(is_wknd = wkdy %in% c(1, 7)) %>%
  # Extract time of the day from datetime
  mutate(time = as.POSIXct(as.ITime(date_round), format = "%H:%M:%S")) %>%
  # Sort the dataframe by date
  arrange(date)
my_df
```

## Displaying Peaks

We can look at the plumes with the best fits for before and after to visualize
the change of plumes from before and after the purifier was installed.

```{r}
my_df_subset <- my_df[my_df$participant_id=="engineering" & my_df$environment=="indoor",]

my_df_before <- my_df_subset[my_df_subset$case=="off",]
my_df_after  <- my_df_subset[my_df_subset$case=="on",]

```


```{r}
# ENTER BEGINNING AND ENDING INDICES FOR PEAKS HERE
peak_idx_beginning <- 5933
peak_idx_ending <- 7500
my_df_before %>%
  ggplot(aes(x = as.numeric(row.names(my_df_before)))) +
  # ENTER PM TYPE HERE
  geom_line(aes(y = pm25)) +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') +
  ylab('PM Levels (ug/m3)') +
  # ylim(0,500) +
  # Add fun theme
  theme_minimal() +
  # Scale to just around the plume
  coord_cartesian(xlim = c(peak_idx_beginning, peak_idx_ending))
```

```{r}
# ENTER BEGINNING AND ENDING INDICES FOR PEAKS HERE
peak_idx_beginning <- 22022
peak_idx_ending <- 23245
my_df_after %>%
  ggplot(aes(x = as.numeric(row.names(my_df_after)))) +
  # ENTER PM TYPE HERE
  geom_line(aes(y = pm25)) +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') +
  ylab('PM Levels (ug/m3)') +
  # ylim(0,500) +
  # Add fun theme
  theme_minimal() +
  # Scale to just around the plume
  coord_cartesian(xlim = c(peak_idx_beginning, peak_idx_ending))
```