# All Deployments CPC Data Summarization

## STOP

This script combines the data from ALL deployments with CPC data in this repository (except for HAFTRAP). Have you run ALL the relevant cleaning scripts? These are:

1. Revere - `RH_modpm-cpc_cleaning.Rmd`, `CH_modpm-cpc_cleaning.Rmd`
2. East Boston - `LF_modpm-cpc_cleaning.Rmd`, `SS_modpm-cpc_cleaning.Rmd`

*This file loads  dataframes created from running ALL the above scripts. Make sure you run those files first (if you haven't already) before running this file.*

Results can be found in `summary/Community_All/s_All_C_quants.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/"
```

## Load data
```{r}
var_list <- c("date", "environment", "case", "participant_id", "concent", 
              "area", "place_id")

cpc_df <- data.frame(matrix(ncol = length(var_list), nrow = 0))

cpc_data <- c("Revere/RH/", "Revere/CH/", "EastBoston/LF/", "EastBoston/SS/")

for (file_path in cpc_data) {
  # Load CPC data from RData file
  load(paste0(path_to_data, file_path, "cleaned_modpm-cpc.RData"))
  
  loc <- strsplit(file_path, "/")[[1]]
  
  main_df <- main_df %>%
    # Add the area and place ID
    mutate(area = loc[1], place_id = loc[2]) %>%
    # Select only relevant columns
    select(all_of(var_list)) %>%
    # Standardize case between off and on
    mutate(case = case_when(case == "after" ~ "on", case == "before" ~ "off",
                            .default = case))
                                 
  cpc_df <- rbind(cpc_df, main_df)
}
```

## Pre-processing
```{r}
all_df <- cpc_df %>%
  # Classify all the deployments
  mutate(room_type = case_when(
    # All classroom deployments
    area == "Revere" & str_detect(participant_id, "class") ~ "classroom",
    
    # All office deployments
    area == "Revere" & !str_detect(participant_id, "class") ~ "office",
    place_id %in% c("LF", "SS") & participant_id == "admin" ~ "office",
    
    # All daycare deployments
    place_id %in% c("LF", "SS") & participant_id != "admin" ~ "daycare",
    ))
```

```{r}
# Save dataframe in summary folder
save(all_df, file = "C:/Users/vkuchhal/Documents/hepa-summer23/summary/Community_All/all_cpc.RData")
```


## Helper Function: Calculate summary statistics
Provided a given grouped dataframe, calculate all relevant summary statistics
```{r}
my_summarise <- function(grouped_df) {
  grouped_df %>%
    summarise(mean = mean(reading),
            median = median(reading), 
            q5 = quantile(reading, probs = 0.05), 
            q25 = quantile(reading, probs = 0.25),
            q75 = quantile(reading, probs = 0.75),
            q95 = quantile(reading, probs = 0.95),
            sd = sd(reading),
            .groups = 'drop')
}
```


## Main Code Run
### Gather, group, and summarize
To calculate summary statistics of the data, the concent variable gathered into one variable 'reading'. The resulting long dataframe is cleaned for NaN values and grouped. Numerous summary statistics are calculated for each type for each, case, environment, and participant.

```{r}
# Calculate for indoor and ratio
main_df_long <- all_df %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(concent), 
               names_to = "type", values_to = "reading") %>%
  # Drop NaN, inf values for summary calculation
  filter(is.finite(reading))
  

# Calculate summary statistics for each participant
participant_summary <- main_df_long %>%
  drop_na() %>%
  # Group by participant ID, case, environment, and type
  group_by(participant_id, room_type, area, place_id, case, environment, type) %>%
  # Calculate summary statistics
  my_summarise()

# Calculate summary statistics for each room type
class_summary <- main_df_long %>%
  drop_na() %>%
  # Group by participant ID, case, environment, and type
  group_by(room_type, case, environment, type) %>%
  # Calculate summary statistics
  my_summarise() %>%
  # Add participant ID column as 'all'
  mutate(participant_id = "all", area = "all", place_id = "all")

# Calculate summary statistics over all the data
overall_summary <- main_df_long %>%
  drop_na() %>%
  # Group by case, environment, and type
  group_by(case, environment, type) %>%
  # Calculate summary statistics
  my_summarise() %>%
  # Add participant ID column as 'all'
  mutate(room_type = "all", participant_id = "all", area = "all", place_id = "all")
```

```{r}
# Bind together participant and overall summary statistics
summary <- rbind(participant_summary, class_summary, overall_summary)
```

### Calculate percentage reduction in concentration from off to on HEPA installation
Follows the same logic as calculating indoor-outdoor ratios. Spread to form 
off/on columns that are used to calculate percent reduction in corresponding
summary pollution concentrations. Then gather dataframe back to original shape.
```{r}
# Spread: Create separate off, on summary statistics columns
summary_wide <- pivot_wider(summary, 
                            names_from = case, values_from = mean:sd)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("off")) - 
  select(summary_wide, ends_with("on"))) / 
  select(summary_wide, ends_with("off"))

# Rename computed columns to 'redu' for percentage reduction
colnames(summary_redu) <- sub("off", "redu", colnames(summary_redu))

# Reshape data and pipe into variable 'summary_normal'
summary_wide %>%
  # Merge back with joined data-frame
  cbind(summary_redu) %>%
  # Gather: Return to original shape by removing the 'sham/hepa/redu'
  # suffixes by re-forming the case variable
  pivot_longer(mean_off:sd_redu, 
               names_to = c(".value", "case"), names_sep = "_") %>%
  # Filter out all NaN and Inf values
  filter_all(all_vars(!is.infinite(.))) -> summary_normal
```

## Save to file
Summary statistics are saved in a CSV file.
```{r}
# Save summary statistic files to CSVs
write.csv(summary_normal, "summary/Community_All/s_All_C_quants.csv")
```