# All Deployments Data Summarization

## STOP

This script combines the data from ALL deployments that contain CPC data in this repository (except for HAFTRAP). Have you run ALL the relevant cleaning scripts? These are:

1. Revere - `RH_modpm-cpc_cleaning.Rmd`, `CH_modpm-cpc_cleaning.Rmd`
2. East Boston - `LF_modpm-cpc_cleaning.Rmd`, `SS_modpm-cpc_cleaning.Rmd`

*This file loads two dataframes created from running ALL the above scripts. Make sure you run those files first (if you haven't already) before running this file.*

Results can be found in `summary/Community_All/s_AllwCPC_MC_corrs.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/"
```

## Load data
```{r}
var_list <- c("date", "environment", "case", "participant_id", "room",  
              "pm1", "pm25", "pm10",  "pm1num", "concent", "area", "place_id")

modcpc_df <- data.frame(matrix(ncol = length(var_list), nrow = 0))

modcpc_data <- c("Revere/RH/", "Revere/CH/", "EastBoston/LF/", "EastBoston/SS/")

for (file_path in modcpc_data) {
  # Load Mod-PM data from RData file
  load(paste0(path_to_data, file_path, "cleaned_modpm-cpc.RData"))
  
  loc <- strsplit(file_path, "/")[[1]]
  
  main_df <- main_df %>%
    # Make a column for room if not present, make columns for deployments
    mutate(room = ifelse("room" %in% colnames(main_df), room, NA),
           area = loc[1], place_id = loc[2]) %>%
    # Select only relevant columns
    select(all_of(var_list)) %>%
    # Standardize case between off and on
    mutate(case = case_when(case == "after" ~ "on", case == "before" ~ "off",
                            .default = case))
  
  modcpc_df <- rbind(modcpc_df, main_df)
}
```


```{r}
# Save dataframe in summary folder
save(modcpc_df, file = "C:/Users/vkuchhal/Documents/hepa-summer23/summary/Community_All/mod_cpc.RData")
```

### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables. This function calculates the correlation matrix for a particular case for a single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, case, person, environment) {
  current_place <- unique(select(df, place_id))
  if (person == "all") {
    current_place <- "all"
  }

  df %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations (suppress output message)
    correlate(quiet = TRUE) %>%
    # Reshape into paired-list format
    stretch(remove.dups = TRUE) %>%
    # Add the case and participant ID
    mutate(case = case, participant_id = person, environment = environment, 
           place_id = paste(current_place)) %>%
    # Drop NaNs
    drop_na()
}
```

### Calculate correlation matrices
Helps better understand the relationship between different variables in HEPA purifier off and on conditions for each participant. Matrices are reshaped into longer pair format for stacking all the data into single dataframe. (Warning: this code takes a while to run.)
```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()

# List of participants with both Mod-PM and CPC
participants <- unique(modcpc_df$participant_id)

# Loop through each case
for (c in c("off", "on")) {
  # Loop through each environment
  for (env in c("indoor")) {
    # Loop through each participant
    for (participant in participants) {
      # Use wide dataframe where indoor/outdoor are separate columns
      modcpc_df %>%
        # Filter for case
        filter(case == c, environment == env, participant_id == participant) %>%
        # Get correlation coefficients (look at helper function)
        get_corr(case = c, environment = env, person = participant) -> df
      
      # Append to main dataframe
      corr_df <- rbind(corr_df, df)
      print(paste("Computed correlation coefficients for",participant, c, env))
    }
    # Repeat for overall correlation coefficient
    modcpc_df %>%
      # Filter for case
      filter(case == c, environment == env) %>%
      # Get correlation coefficients (look at helper function)
      get_corr(case = c, environment = env, person = "all") -> df
    
    # Append to main dataframe
    corr_df <- rbind(corr_df, df)
    print(paste("Computed correlation coefficients for all", c, env))
  }
}
```

## Save to file
Summary statistics are saved in a CSV file.
```{r}
# Save summary statistic files to CSVs
write.csv(corr_df, "summary/Community_All/s_AllwCPC_MC_corrs.csv")
```