---
editor_options: 
  markdown: 
    wrap: 72
---

# Community All Plume Plotting (for Mod-PM)

Plot various plumes before and after HEPA installation to compare their
heights, widths, decay constants, etc.

This file compiles all plumes from all deployments, saved in
`summary/{location_name}/{deployment_name}/s_{deployment_name}_M_allplumes.csv`,
and computes summary statistics for every participant, every deployment, and
every room type.

```{r}
# import necessary libraries
library(tidyverse)
library(data.table)
library(scales)
library(pracma)
```

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = "~/AirPartners/hepa-summer23")
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

## Load data

```{r}
load_data <- function(paths_to_data) {
  res <- data.frame()
  # For each path, read the CSV file in that path
  for (path in paths_to_data) {
    main <- read_csv(path)
    # Append dataframes together
    if (length(res) > 1) {
      res <- rbind(res, main)
    } else {
      res <- data.frame(main)
    }
  }
  return(res)
}
```

```{r}
PATHS_TO_DATA <- c(
  # East Boston
  "summary/EastBoston/MF/s_MF_M_allplumes.csv",
  "summary/EastBoston/LF/s_LF_M_allplumes.csv",
  "summary/EastBoston/SS/s_SS_M_allplumes.csv",
  # Revere
  "summary/Revere/CH/s_CH_M_allplumes.csv",
  "summary/Revere/RH/s_RH_M_allplumes.csv",
  # Roxbury
  "summary/Roxbury/CM/s_CM_M_allplumes.csv",
  "summary/Roxbury/IB/s_IB_M_allplumes.csv"
)

main_df <- load_data(PATHS_TO_DATA)
```

## Filtering the Dataset
```{r}
my_df <- main_df %>%
  # Remove unnecessary index column
  mutate(...1 = NULL) %>%
  # Remove environment (always indoor)
  mutate(environment = NULL) %>%
  # Only include peaks that have a width greater than 30
  filter(peak_width >= 30) %>%
  # Get rid of outliers
  filter(peak_width < ifelse(room=="office", 2000, 6000)) %>%
  # Only include peaks with a height greater than 5
  filter(peak_hgt >= 5) %>%
  # Only include decay constants that are negative
  filter(k_val < 0) %>%
  # Include peaks where the conv_tol is less than third quantile
  filter(conv_tol <= 0.000005702) %>%
  # Remove peaks that take up entire span of data
  filter(abs(cumm_pm - total_pm) > 0.1)
my_df
```


## Characterize Decay Constants by Room Type

The rest of this file calculated statistics on plume reduction
throughout the dataset. Below summarizes all that analysis into one
function, creating reduction statistics, and performing that function on
the dataset as it is grouped by participant and room.

```{r}
plume_redus <- function(data, room_type) {  
  decay_sum <-
    data %>%
      filter(room == room_type) %>%
      group_by(pm, hepa) %>%
      summarize(
        # kval
        k_val_mean = mean(k_val, na.rm = FALSE),
        k_val_min = min(k_val, na.rm = FALSE),
        k_val_q25 = quantile(k_val, prob = 0.25, na.rm = FALSE),
        k_val_median = median(k_val, na.rm = TRUE),
        k_val_q75 = quantile(k_val, prob = 0.75, na.rm = TRUE),
        k_val_max = max(k_val, na.rm = TRUE),
        # peak heights
        peak_hgt_mean = mean(peak_hgt, na.rm = TRUE),
        peak_hgt_median = median(peak_hgt, na.rm = TRUE),
        peak_hgt_min = min(peak_hgt, na.rm = TRUE),
        peak_hgt_max = max(peak_hgt, na.rm = TRUE),
        # total peak widths
        peak_wdh_mean = mean(peak_width, na.rm = TRUE),
        peak_wdh_median = median(peak_width, na.rm = TRUE),
        peak_wdh_min = min(peak_width, na.rm = TRUE),
        peak_wdh_max = max(peak_width, na.rm = TRUE),
        # cumulative PM
        total_plume_area = sum(cumm_pm, na.rm = TRUE),
        # number of plumes
        n = n(),
        .groups = "keep"
      )
  
  decay_sum <- decay_sum %>% mutate(room = room_type)
  
  return(data.frame(decay_sum[,c(19, 1:18)]))
}
```

```{r}
plume_locs <- function(data, loc) {  
  decay_sum <-
    data %>%
      filter(deployment == loc) %>%
      group_by(pm, hepa) %>%
      summarize(
        # kval
        k_val_mean = mean(k_val, na.rm = TRUE),
        k_val_min = min(k_val, na.rm = TRUE),
        k_val_q25 = quantile(k_val, prob = 0.25, na.rm = TRUE),
        k_val_median = median(k_val, na.rm = TRUE),
        k_val_q75 = quantile(k_val, prob = 0.75, na.rm = TRUE),
        k_val_max = max(k_val, na.rm = TRUE),
        # peak heights
        peak_hgt_mean = mean(peak_hgt, na.rm = TRUE),
        peak_hgt_median = median(peak_hgt, na.rm = TRUE),
        peak_hgt_min = min(peak_hgt, na.rm = TRUE),
        peak_hgt_max = max(peak_hgt, na.rm = TRUE),
        # total peak widths
        peak_wdh_mean = mean(peak_width, na.rm = TRUE),
        peak_wdh_median = median(peak_width, na.rm = TRUE),
        peak_wdh_min = min(peak_width, na.rm = TRUE),
        peak_wdh_max = max(peak_width, na.rm = TRUE),
        # cumulative PM
        total_plume_area = sum(cumm_pm, na.rm = TRUE),
        # number of plumes
        n = n(),
        .groups = "keep"
      )
  
  decay_sum <- decay_sum %>% mutate(deployment = loc)
  
  return(data.frame(decay_sum[,c(19, 1:18)]))
}
```


```{r}
# Create empty DataFrame for storing k values
plume_reduc.data <- data.frame()

for (room in unique(my_df$room)) {
  print(paste("Calculating for room type", room))
  new_reducs <- plume_redus(my_df, room)
  # Rbind them with previous data
  plume_reduc.data <- rbind(plume_reduc.data, new_reducs)
}
plume_reduc.data
```

```{r}
# Create empty DataFrame for storing k values
plume_locs.data <- data.frame()

for (loc in unique(my_df$deployment)) {
  print(paste("Calculating for location", loc))
  new_reducs <- plume_locs(my_df, loc)
  # Rbind them with previous data
  plume_locs.data <- rbind(plume_locs.data, new_reducs)
}
plume_locs.data
```

## Visualizations Based on Room Type

These are visualizations of decay constant summary statistics for each room
type. These are not necessarily useful for the paper; I just created these for
exploratory analysis.

```{r}
plume_reduc.data %>%
  filter(pm == "pm1") %>%
  group_by(room) %>%
  summarise(
    min_reduc = (k_val_min[hepa==FALSE] - k_val_min[hepa==TRUE])/abs(k_val_min[hepa==FALSE]),
    q25_reduc = (k_val_q25[hepa==FALSE] - k_val_q25[hepa==TRUE])/abs(k_val_q25[hepa==FALSE]),
    median_reduc = (k_val_median[hepa==FALSE] - k_val_median[hepa==TRUE])/abs(k_val_median[hepa==FALSE]),
    mean_reduc = (k_val_mean[hepa==FALSE] - k_val_mean[hepa==TRUE])/abs(k_val_mean[hepa==FALSE]),
    q75_reduc = (k_val_q75[hepa==FALSE] - k_val_q75[hepa==TRUE])/abs(k_val_q75[hepa==FALSE]),
    max_reduc = (k_val_max[hepa==FALSE] - k_val_max[hepa==TRUE])/abs(k_val_max[hepa==FALSE])
  ) %>%
  pivot_longer(cols = min_reduc:max_reduc, names_to = "stat") %>%
  mutate(stat = fct_relevel(stat, "min_reduc", "q25_reduc", "median_reduc", 
                            "mean_reduc", "q75_reduc", "max_reduc")) %>%
  ggplot() + geom_col(aes(x = room, y = value, fill = stat), position = "dodge")
```

```{r}
plume_reduc.data %>%
  filter(pm == "pm25") %>%
  group_by(room) %>%
  summarise(
    min_reduc = (k_val_min[hepa==FALSE] - k_val_min[hepa==TRUE])/abs(k_val_min[hepa==FALSE]),
    q25_reduc = (k_val_q25[hepa==FALSE] - k_val_q25[hepa==TRUE])/abs(k_val_q25[hepa==FALSE]),
    median_reduc = (k_val_median[hepa==FALSE] - k_val_median[hepa==TRUE])/abs(k_val_median[hepa==FALSE]),
    mean_reduc = (k_val_mean[hepa==FALSE] - k_val_mean[hepa==TRUE])/abs(k_val_mean[hepa==FALSE]),
    q75_reduc = (k_val_q75[hepa==FALSE] - k_val_q75[hepa==TRUE])/abs(k_val_q75[hepa==FALSE]),
    max_reduc = (k_val_max[hepa==FALSE] - k_val_max[hepa==TRUE])/abs(k_val_max[hepa==FALSE])
  ) %>%
  pivot_longer(cols = min_reduc:max_reduc, names_to = "stat") %>%
  mutate(stat = fct_relevel(stat, "min_reduc", "q25_reduc", "median_reduc", 
                            "mean_reduc", "q75_reduc", "max_reduc")) %>%
  ggplot() + geom_col(aes(x = room, y = value, fill = stat), position = "dodge")
```
```{r}
plume_reduc.data %>%
  filter(pm == "pm10") %>%
  group_by(room) %>%
  summarise(
    min_reduc = (k_val_min[hepa==FALSE] - k_val_min[hepa==TRUE])/abs(k_val_min[hepa==FALSE]),
    q25_reduc = (k_val_q25[hepa==FALSE] - k_val_q25[hepa==TRUE])/abs(k_val_q25[hepa==FALSE]),
    median_reduc = (k_val_median[hepa==FALSE] - k_val_median[hepa==TRUE])/abs(k_val_median[hepa==FALSE]),
    mean_reduc = (k_val_mean[hepa==FALSE] - k_val_mean[hepa==TRUE])/abs(k_val_mean[hepa==FALSE]),
    q75_reduc = (k_val_q75[hepa==FALSE] - k_val_q75[hepa==TRUE])/abs(k_val_q75[hepa==FALSE]),
    max_reduc = (k_val_max[hepa==FALSE] - k_val_max[hepa==TRUE])/abs(k_val_max[hepa==FALSE])
  ) %>%
  pivot_longer(cols = min_reduc:max_reduc, names_to = "stat") %>%
  mutate(stat = fct_relevel(stat, "min_reduc", "q25_reduc", "median_reduc", 
                            "mean_reduc", "q75_reduc", "max_reduc")) %>%
  ggplot() + geom_col(aes(x = room, y = value, fill = stat), position = "dodge")
```

```{r}
plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    mean_reduc = (k_val_mean[hepa==FALSE] - k_val_mean[hepa==TRUE])/abs(k_val_mean[hepa==FALSE])
  ) %>%
  ggplot() + geom_col(aes(x = room, y = mean_reduc, fill = pm), position = "dodge")
```

```{r}
plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    peak_reduc = (peak_wdh_mean[hepa==FALSE] - peak_wdh_mean[hepa==TRUE])/abs(peak_wdh_mean[hepa==FALSE])
  ) %>%
  ggplot() + geom_col(aes(x = room, y = peak_reduc, fill = pm), position = "dodge")
```
```{r}
plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    peak_reduc = (peak_hgt_mean[hepa==FALSE] - peak_hgt_mean[hepa==TRUE])/abs(peak_hgt_mean[hepa==FALSE])
  ) %>%
  ggplot() + geom_col(aes(x = room, y = peak_reduc, fill = pm), position = "dodge")
```

## Decay Constant Box and Whiskers

```{r}
options(repr.plot.width=8, repr.plot.height=15)

# Peak height 5
# Peak width 30

pm1_form <- deparse(""~PM[1])
pm25_form <- deparse(""~PM[2.5])
pm10_form <- deparse(""~PM[10])

plume_data_formatted <- plume_reduc.data %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  ))

k_val_incs <-
  plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    k_val_reduc = (k_val_mean[hepa==TRUE] - k_val_mean[hepa==FALSE])/abs(k_val_mean[hepa==FALSE])
  ) %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  ))



my_df %>%
  filter(room != "classroom") %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  )) %>%
  # mutate(hepa = ifelse(hepa, "on", "off")) %>%
  group_by(room) %>%
  ggplot(aes(x = room, y = abs(k_val))) + 
  geom_boxplot(aes(fill = factor(hepa)), position = position_dodge(0.8), outlier.shape = NA) +
  scale_y_log10() +
  labs(title = "Decay Constant Magnitudes Based on Room Types",
       x = "Room Type",
       y = "Decay Constants",
       fill = "Purifier State") +
  geom_text(aes(y = 1, label = paste0(round(abs(k_val_reduc*100), 1), "%")),
            data = k_val_incs, fontface = "bold") +
  geom_point(aes(x = room, y = abs(k_val_mean), fill = hepa), data = plume_data_formatted, 
             position = position_dodge(0.8)) +
  facet_grid(. ~ factor(pm_formatted, levels = c(pm1_form, pm25_form, pm10_form)),
             labeller = label_parsed) +
  theme_bw()
```


```{r}
# Using log returns instead of percent changes
options(repr.plot.width=8, repr.plot.height=15)

# Peak height 5
# Peak width 30

pm1_form <- deparse(""~PM[1])
pm25_form <- deparse(""~PM[2.5])
pm10_form <- deparse(""~PM[10])

plume_data_formatted <- plume_reduc.data %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  ))

k_val_rets <-
  plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    k_val_reduc = log(abs(k_val_median[hepa==TRUE] / k_val_median[hepa==FALSE]), base=10)
  ) %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  ))



my_df %>%
  filter(room != "classroom") %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  )) %>%
  # mutate(hepa = ifelse(hepa, "on", "off")) %>%
  group_by(room) %>%
  ggplot(aes(x = room, y = abs(k_val))) + 
  geom_boxplot(aes(fill = factor(hepa)), position = position_dodge(0.8), outlier.shape = NA) +
  scale_y_log10() +
  labs(title = "Decay Constant Magnitudes Based on Room Types",
       x = "Room Type",
       y = "Decay Constants",
       fill = "Purifier State") +
  geom_text(aes(y = 1, label = round(abs(k_val_reduc), 2)),
            data = k_val_rets, fontface = "bold") +
  geom_point(aes(x = room, y = abs(k_val_mean), fill = hepa), data = plume_data_formatted, 
             position = position_dodge(0.8)) +
  facet_grid(. ~ factor(pm_formatted, levels = c(pm1_form, pm25_form, pm10_form)),
             labeller = label_parsed) +
  theme_bw()
```

```{r}

k_val_incs <-
  plume_reduc.data %>%
  group_by(room, pm) %>%
  summarise(
    k_val_reduc = (k_val_median[hepa==TRUE] - k_val_median[hepa==FALSE])/abs(k_val_median[hepa==FALSE])
  ) %>%
  mutate(pm_formatted = case_when(
    pm == "pm1" ~ pm1_form,
    pm == "pm25" ~ pm25_form,
    pm == "pm10" ~ pm10_form
  ))
k_val_incs
```

```{r}
my_df %>%
  filter(room != "classroom") %>%
  mutate(hepa = ifelse(hepa, "on", "off")) %>%
  group_by(deployment) %>%
  ggplot(aes(x = deployment, y = peak_hgt, fill = factor(hepa))) + 
  geom_boxplot(position = "dodge", outlier.shape = NA) +
  scale_y_log10() +
  labs(title = "Peak Heights Based on Room Types",
       x = "Room Type",
       y = "Peak Heights (ug/m3)",
       fill = "HEPA") +
  facet_grid(vars(pm)) +
  theme_minimal()
```

## Exposure Statistics


```{r}
# Peak width 30
# Peak height 5

gamma = 0.006

# Remove Mothers Out Front
my_df_temp <- my_df
  # filter(deployment != "MF")

# Create empty DataFrame for storing k values
plume_reduc.data.temp <- data.frame()

for (room in unique(my_df_temp$room)) {
  print(paste("Calculating for room type", room))
  new_reducs <- plume_redus(my_df_temp, room)
  # Rbind them with previous data
  plume_reduc.data.temp <- rbind(plume_reduc.data.temp, new_reducs)
}

total_exposure <- my_df_temp %>%
  group_by(participant_id, hepa, pm, room) %>% 
  summarize(total_exposure = median(total_pm)) %>% 
  group_by(room, pm, hepa) %>% 
  summarize(total_exposure = sum(total_exposure)) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  filter(pm == "pm1")

counts <- my_df_temp %>% group_by(room) %>% summarize(count = n_distinct(participant_id))

exposure <- plume_reduc.data.temp[plume_reduc.data.temp$pm=="pm1",] %>%
  select(room, hepa, total_plume_area, n) %>%
  group_by(room, hepa) %>%
  mutate(sum_plume = sum(total_plume_area)) %>%
  mutate(avg_plume = sum_plume/n) %>%
  distinct(room, hepa, .keep_all=TRUE) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  right_join(total_exposure, by = join_by(room, hepa)) %>%
  mutate(prop = sum_plume / total_exposure) %>%
  left_join(counts, by = join_by(room))
exposure

# Total exposure with plumes vs. nonplumes (unnormalized)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = gamma*sum_plume), stat = "identity", position = "dodge") +
  geom_bar(aes(y = gamma*total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Total Exposure for Each Room Type",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  theme_bw()

# Average exposure level per deployment for each room type
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = gamma*sum_plume/count), stat = "identity", position = "dodge") +
  geom_bar(aes(y = gamma*total_exposure/count), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Average Exposure per Deployment for Each Room Type",
       x = "Room Type",
       y = expression(paste("Average Exposure (", mu, "g)")),
       fill = "Purifier State") +
  theme_bw()

# Exposure metrics - proportions (non-annotated)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = (sum_plume/total_exposure)), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Exposure from Plumes vs. Baseline for Room Types",
       x = "Room Type",
       y = "Proportion",
       fill = "Purifier State") +
  theme_bw()

# Exposure metrics - proportions (annotated)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = (sum_plume/total_exposure)), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Exposure from Plumes vs. Baseline for Room Types",
       x = "Room Type",
       y = "Proportion",
       fill = "Purifier State") +
  # ANNOTATIONS
  annotate("text", x = 0.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[1], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 1.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[2], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 1.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[3], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 2.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[4], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 2.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[5], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 3.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[6], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  theme_bw()


# Exposure per plume per room type
ggplot(exposure, aes(x = factor(room), y = gamma*avg_plume, fill = factor(hepa))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title="Average Exposure per Plume for Each Room Type",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  theme_bw()
```



```{r}
# SAME CODE AS ABOVE BUT MF REDEFINED AS ENTIRELY RESIDENTS

my_df_temp <- my_df %>%
  mutate(room = ifelse(deployment=="MF", "residential", room))

# Create empty DataFrame for storing k values
plume_reduc.data.temp <- data.frame()

for (room in unique(my_df_temp$room)) {
  print(paste("Calculating for room type", room))
  new_reducs <- plume_redus(my_df_temp, room)
  # Rbind them with previous data
  plume_reduc.data.temp <- rbind(plume_reduc.data.temp, new_reducs)
}

total_exposure <- my_df_temp %>%
  group_by(participant_id, hepa, pm, room) %>% 
  summarize(total_exposure = median(total_pm)) %>% 
  group_by(room, pm, hepa) %>% 
  summarize(total_exposure = sum(total_exposure)) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  filter(pm == "pm1")

counts <- my_df_temp %>% group_by(room) %>% summarize(count = n_distinct(participant_id))

exposure <- plume_reduc.data.temp[plume_reduc.data.temp$pm=="pm1",] %>%
  select(room, hepa, total_plume_area, n) %>%
  group_by(room, hepa) %>%
  mutate(sum_plume = sum(total_plume_area)) %>%
  mutate(avg_plume = sum_plume/n) %>%
  distinct(room, hepa, .keep_all=TRUE) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  right_join(total_exposure, by = join_by(room, hepa)) %>%
  mutate(prop = sum_plume / total_exposure) %>%
  left_join(counts, by = join_by(room))
exposure

# Total exposure with plumes vs. nonplumes (unnormalized)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = sum_plume), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  scale_y_log10() +
  labs(title="Total Exposure for Each Room Type",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  theme_bw()

# Average exposure level per deployment for each room type
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = sum_plume/count), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/count), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Average Exposure per Deployment for Each Room Type",
       x = "Room Type",
       y = expression(paste("Average ", PM[1], " Exposure (", mu, "g/", m^3,")")),
       fill = "Purifier State") +
  theme_bw()

# Exposure metrics - proportions (non-annotated)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = (sum_plume/total_exposure)), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Exposure from Plumes vs. Baseline for Room Types",
       x = "Room",
       y = "Proportion",
       fill = "Purifier State") +
  theme_bw()

# Exposure metrics - proportions (annotated)
ggplot(exposure, aes(x = factor(room), fill = factor(hepa))) +
  geom_bar(aes(y = (sum_plume/total_exposure)), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Exposure from Plumes vs. Baseline for Room Types",
       x = "Room",
       y = "Proportion",
       fill = "Purifier State") +
  # ANNOTATIONS
  annotate("text", x = 0.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[3], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 1.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[4], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 1.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[5], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 2.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[6], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 2.78, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[1], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  annotate("text", x = 3.23, y = 0.15, 
           label = paste0(format(round(100*exposure$prop[2], 1), nsmall=1), "%"),
           vjust = 2, size = 5) +
  theme_bw()


# Exposure per plume per room type
ggplot(exposure, aes(x = factor(room), y = avg_plume, fill = factor(hepa))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title="Average Exposure per Plume for Each Room Type",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  theme_bw()
```

```{r}
# Peak width 30
# Peak height 5

total_exposure <- my_df %>%
  group_by(participant_id, hepa, pm, deployment) %>% 
  summarize(total_exposure = median(total_pm)) %>% 
  group_by(deployment, pm, hepa) %>% 
  summarize(total_exposure = sum(total_exposure)) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  filter(pm == "pm1")

counts <- my_df %>% group_by(deployment) %>% summarize(count = n_distinct(participant_id))

exposure <- plume_locs.data[plume_locs.data$pm=="pm1",] %>%
  select(deployment, hepa, total_plume_area, n) %>%
  group_by(deployment, hepa) %>%
  mutate(sum_plume = sum(total_plume_area)) %>%
  mutate(avg_plume = sum_plume/n) %>%
  distinct(deployment, hepa, .keep_all=TRUE) %>%
  mutate(hepa = ifelse(hepa==TRUE, "on", "off")) %>%
  right_join(total_exposure, by = join_by(deployment, hepa)) %>%
  left_join(counts, by = join_by(deployment))
exposure


# Create the bar plot with custom legend labels
ggplot(exposure, aes(x = factor(deployment), fill = factor(hepa))) +
  geom_bar(aes(y = sum_plume), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Total Exposure for Each Deployment",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  # scale_fill_manual(values = c("FALSE" = "red", "TRUE" = "green"),
  #                   labels = c("FALSE" = "off", "TRUE" = "on")) +
  theme_minimal()

# Create the bar plot with custom legend labels
ggplot(exposure, aes(x = factor(deployment), fill = factor(hepa))) +
  geom_bar(aes(y = sum_plume/count), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/count), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Total Exposure for Each Deployment",
       x = "Room Type",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  # scale_fill_manual(values = c("FALSE" = "red", "TRUE" = "green"),
  #                   labels = c("FALSE" = "off", "TRUE" = "on")) +
  theme_minimal()

# Exposure metrics - proportions (annotated)
ggplot(exposure, aes(x = factor(deployment), fill = factor(hepa))) +
  geom_bar(aes(y = (sum_plume/total_exposure)), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_exposure/total_exposure), stat = "identity", 
           position = "dodge", alpha = 0.2) +
  labs(title="Exposure from Plumes vs. Baseline for Room Types",
       x = "Room",
       y = "Proportion",
       fill = "Purifier State") +
  theme_bw()


ggplot(exposure, aes(x = factor(deployment), y = avg_plume, fill = factor(hepa))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title="Average Exposure per Plume for Each Deployment",
       x = "Deployment",
       y = "Total Exposure (ug/m3)",
       fill = "Purifier State") +
  # scale_fill_manual(values = c("FALSE" = "red", "TRUE" = "green"),
  #                   labels = c("FALSE" = "off", "TRUE" = "on")) +
  theme_minimal()
```



From the analysis above, it seems that organizing this data by room type does
not reveal any sort of emerging patterns. 

## Decay Constant and Exposure Curve Visualization

```{r}
# Define the exponential function
decay_func <- function(x, k, init_hgt) {
  return(init_hgt*exp(k*x))  # Exponential function e^x
}
```

```{r}
# Construct curves with set initial height of 20
init_hgt <- 43.2

# Initialize a series of x values
time <- seq(0, 250, by=0.2)

# Use filtered version of plume reductions data
decay_info <- plume_reduc.data %>% filter(room == "residential", pm == "pm1")

# Get median output from decay function
bpm_median <- decay_func(time, decay_info$k_val_mean[1], init_hgt)
bpm_25th   <- decay_func(time, decay_info$k_val_q25[1], init_hgt)
bpm_75th   <- decay_func(time, decay_info$k_val_q75[1], init_hgt)

apm_median <- decay_func(time, decay_info$k_val_mean[2], init_hgt)
apm_25th   <- decay_func(time, decay_info$k_val_q25[2], init_hgt)
apm_75th   <- decay_func(time, decay_info$k_val_q75[2], init_hgt)

# Bind them into a dataframe
plot_data <- as.data.frame(t(rbind(time,
                                   bpm_median, bpm_25th, bpm_75th,
                                   apm_median, apm_25th, apm_75th)))
plot_data.col_names <- c("time", 
                         "bpm_median", "bpm_25th", "bpm_75th",
                         "apm_median", "apm_25th", "apm_75th")
```

```{r}
# Plot!
ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = bpm_median),
            color="#F8766D",
            linewidth=1) +
  geom_ribbon(aes(ymin = bpm_25th, ymax = bpm_75th),
              alpha=0.1,       #transparency
              linetype=0,      #solid, dashed or other line types
              colour="grey70", #border line color
              fill="#F8766D") +    #fill color
  geom_line(aes(y = apm_median),
            color="#00BFC4",
            linewidth=1) +
  geom_ribbon(aes(ymin = apm_25th, ymax = apm_75th),
              alpha=0.1,       #transparency
              linetype=0,      #solid, dashed or other line types
              colour="black",  #border line color
              fill="#00BFC4") +  #fill color
  # median of PM1 q25 vals for residences
  geom_hline(yintercept = 1.975, linetype = "dotted") +
  # Annotations
  annotate("point", x = 125.9, y = 1.975, color = "black", size = 2) +
  annotate("point", x = 215.8, y = 1.975, color = "black", size = 2) +
  annotate("text", x = 126, y = 1.975, label = "125.9", vjust = -1, size = 3) +
  annotate("text", x = 216, y = 1.975, label = "215.8", vjust = -1, size = 3) +
  ggtitle(expression(paste(PM[1], " Decay Constants Across Residences"))) +
  xlab("Time (min)") +
  ylab(expression(paste(PM[1], " Concentration (", mu, "g/", m^3,")"))) +
  theme_bw()
```

```{r}
# Plot!
ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = bpm_median),
            colour="#F8766D",
            linewidth=1) +
  geom_line(aes(y = apm_median),
            colour="#00BFC4",
            linewidth=1) +
  geom_hline(yintercept = 1.975, linetype = "dotted") +
  # Annotations
  annotate("point", x = 125.9, y = 1.975, color = "black", size = 2) +
  annotate("point", x = 215.8, y = 1.975, color = "black", size = 2) +
  annotate("text", x = 125.9, y = 1.975, label = "125.9", vjust = 2, size = 3) +
  annotate("text", x = 215.8, y = 1.975, label = "215.8", vjust = 2, size = 3) +
  # Titles
  ggtitle("Average Rate of Decay for Before and After
  HEPA Installation") +
  xlab("Time (min)") +
  ylab(expression(paste(PM[1], " Concentration (", mu, "g/", m^3,")"))) +
  theme_bw()
```


```{r}
# Plot!
ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = bpm_median),
            color="#F8766D",
            linewidth=1) +
  geom_line(aes(y = apm_median),
            color="#00BFC4",
            linewidth=1) +
  
  geom_ribbon(aes(ymin = 0, ymax = bpm_median),
              alpha=0.1,       #transparency
              linetype=0,      #solid, dashed or other line types
              colour="grey70", #border line color
              fill="#F8766D") +#fill color
  geom_ribbon(aes(ymin = 0, ymax = apm_median),
              alpha=0.1,       #transparency
              linetype=0,      #solid, dashed or other line types
              colour="grey70", #border line color
              fill="#00BFC4") +#fill color
  geom_hline(yintercept = 5, linetype = "dotted") +
  # Annotations
  # annotate("point", x = 56.3763, y = 5, color = "black", size = 2) +
  # annotate("point", x = 96.8083, y = 5, color = "black", size = 2) +
  # annotate("text", x = 55, y = 5, label = "56.3763", vjust = 2, size = 3) +
  # annotate("text", x = 97, y = 5, label = "96.8083", vjust = 2, size = 3) +
  # annotate("text", x = 125, y = 17, label = "Exposure Difference: 319.31 ug/m3", size = 3) +
  ggtitle("Average Rate of Decay for Before and After
  HEPA Installation with Exposure Reduction") +
  xlab("Time (min)") +
  ylab(expression(paste(PM[1], " Concentration (", mu, "g/", m^3,")"))) +
  theme_bw()
```


## Peak Height and Decay Constants (Appendix Analysis)

```{r}
ggplot(my_df, aes(x = abs(k_val), y = peak_hgt)) + 
  geom_point(shape = 1) +
  geom_smooth(method=lm) +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Height and Decay Constants by Room Type",
       x = "Decay Constant Magnitude",
       y = "Peak Height (ug/m3)") +
  theme_bw()
```

```{r}
ggplot(my_df, aes(x = abs(k_val), y = peak_hgt, color=room)) + 
  geom_point(shape = 1) +
  geom_smooth(method=lm) +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Height and Decay Constants by Room Type",
       x = "Decay Constant Magnitude",
       y = "Peak Height (ug/m3)",
       color = "Room Type") +
  theme_bw()
```

```{r}
my_df %>%
  mutate(hepa = ifelse(hepa, "on", "off")) %>%
  ggplot(aes(x = abs(k_val), y = peak_hgt, color=hepa)) + 
  geom_point(shape = 1) +
  geom_smooth(method=lm) +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Height and Decay Constants by HEPA Status",
       x = "Decay Constant Magnitude",
       y = "Peak Height (ug/m3)",
       color = "Purifier State") +
  theme_bw()
```

```{r}
ggplot(my_df, aes(x = abs(k_val), y = peak_width)) + 
  geom_point(shape = 0, alpha = 0.2) +
  geom_smooth(method=lm, color = "#F8766D") +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Width and Decay Constants",
       x = "Decay Constant Magnitude",
       y = "Peak Width (min)") +
  theme_bw()
```

```{r}
lm(k_val ~ peak_width, data=my_df)
cor(log(abs(my_df$k_val)), log(my_df$peak_width), method = "pearson")
```

```{r}
my_df %>%
  mutate(hepa = ifelse(hepa, "on", "off")) %>%
  ggplot(aes(x = abs(k_val), y = peak_width, color=hepa)) + 
  geom_point(shape = 0, alpha = 0.2) +
  geom_smooth(method=lm) +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Height and Decay Constants by HEPA Status",
       x = "Decay Constant Magnitude",
       y = "Peak Height (ug/m3)",
       color = "Purifier State") +
  theme_bw()
```


```{r}
ggplot(my_df, aes(x = abs(k_val), y = peak_width, color=room)) + 
  geom_point(shape = 1, alpha=0.5) +
  geom_smooth(method=lm) +
  scale_x_log10() +
  scale_y_log10() +
  labs(title="Regressions for Peak Width and Decay Constants by Room Type",
       x = "Decay Constant Magnitude",
       y = "Peak Width (min)",
       color = "Room Type") +
  theme_minimal()
```



## Save Output
Save 
```{r}
write.csv(my_df, file = paste0("summary/Community_All/", "s_All_M_allplumes.csv"))
write.csv(plume_reduc.data, file = paste0("summary/Community_All/", "s_All_M_plumestats.csv"))
```
