# All Deployments HOBO Data Summarization

## STOP

This script combines the data from ALL deployments with cleaned HOBO data in this repository (except for HAFTRAP). Have you run ALL the relevant cleaning scripts? These are:

1. Roxbury - `IB_hobo_cleaning.Rmd`, `CM_hobo_cleaning.Rmd`
2. East Boston - `MF_hobo_cleaning.Rmd`

*This file loads  dataframes created from running ALL the above scripts. Make sure you run those files first (if you haven't already) before running this file.*

Results can be found in `summary/Community_All/s_All_H_props.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/"
```


```{r}
var_list <- c("date", "case", "participant_id", "room", "area", "place_id", "power")

hobo_df <- data.frame(matrix(ncol = length(var_list), nrow = 0))

hobo_data <- c("Roxbury/IB/", "Roxbury/CM/", "EastBoston/MF/")

for (file_path in hobo_data) {
  # Load HOBO data from RData file
  load(paste0(path_to_data, file_path, "cleaned_hobo.RData"))
  
  loc <- strsplit(file_path, "/")[[1]]
  
  if (str_detect(file_path, "CM")) {
    main_df <- main_df %>% mutate(room = NA)
  }
  
  main_df <- main_df %>%
    # Add the area and place ID
    mutate(area = loc[1], place_id = loc[2]) %>%
    # Select only relevant columns
    select(all_of(var_list)) %>%
    # Standardize case between off and on
    mutate(case = case_when(case == "after" ~ "on", case == "before" ~ "off",
                            .default = case))
  
  hobo_df <- rbind(hobo_df, main_df)
}
```

## Find means
```{r}
on_df <- hobo_df %>% filter(case == "on")

participant_prop <- on_df %>%
  group_by(area, place_id, participant_id, room) %>%
  summarise(prop_on = sum(power > 10)/n())

place_prop <- on_df %>%
  group_by(area, place_id) %>%
  summarise(prop_on = sum(power > 10)/n()) %>%
  mutate(participant_id = "all", room = "all")

area_prop <- on_df %>%
  group_by(area) %>%
  summarise(prop_on = sum(power > 10)/n()) %>%
  mutate(place_id = "all", participant_id = "all", room = "all")

overall_prop <- on_df %>%
  summarise(prop_on = sum(power > 10)/n()) %>%
  mutate(area = "all", place_id = "all", participant_id = "all", room = "all")

props_df <- rbind(participant_prop, place_prop, area_prop, overall_prop)
```
## Save to file
Summary statistics are saved in a CSV file.
```{r}
# Save summary statistic files to CSVs
write_csv(props_df, "summary/Community_All/s_All_H_props.csv")
```