---
title: "API Call"
author: "Cherry Pham"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file is created with the purpose of obtaining sensor data
from Quant-AQ through API calls.

There is no prerequisites for using this file. If you are just looking
to quickly retrieve data, you can simply run this file through the
command `Ctrl + Alt + R`. The result of running this file is a data
folder containing data of MOD-PM sensors as of February 2023. If you
wish to modify the data to be retrieved, please read the data guide
below. The original file is written for ambient data in [this
repository](https://github.com/airpartners/ambient-summer23), which
might be more up-to-date.

## DATA GUIDE

[write instructions here]

[sorry if i forgot to do this at the end, the code below is quite well-
documented though so good luck]

## PACKAGES AND INITIAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("httr", "jsonlite", "purrr", "dplyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages
invisible(sapply(packages, library, character.only = TRUE))

# Explicit package function calls
httr::GET
httr::content
jsonlite::fromJSON
purrr::map_df
dplyr::as_data_frame
dplyr::filter
dplyr::lag

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DEFINE VARIABLES

```{r}
# Define API info
base_url <- "https://api.quant-aq.com/device-api/v1/devices/"
api_key <- "VUOLNUDTM70QZ7Z9G0Z78XA7" # Replace this API key with yours if can

# Define the Sensor class
setClass(
  "Sensor",
  slots = c(
    sensorNum = "character",
    locationCode = "character",
    sensorCode = "character",
    ID = "integer",
    case = "character",
    environment = "character",
    startDate = "character",
    endDate = "character"
  )
)

# Constructor function for the Sensor class
Sensor <- function(sensorNum, locationCode, sensorCode, ID, case, environment, startDate, endDate) {
  new(
    "Sensor",
    sensorNum = sensorNum,
    locationCode = locationCode,
    sensorCode = sensorCode,
    ID = ID,
    case = case,
    environment = environment,
    startDate = startDate,
    endDate = endDate
  )
}

# Create Sensor objects corresponding to the table data
# Sensor 1
SENSOR_41181_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41181,
  case = "sham",
  environment = "indoor",
  startDate = "2022-10-21",
  endDate = "2022-10-28"
)

# Sensor 2
SENSOR_41181_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41181,
  case = "sham",
  environment = "outdoor",
  startDate = "2022-10-21",
  endDate = "2022-10-28"
)

# Sensor 3
SENSOR_41181_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41181,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-12-09",
  endDate = "2022-12-16"
)

# Sensor 4
SENSOR_41181_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00126",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41181,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-12-09",
  endDate = "2022-12-16"
)

# Sensor 5
SENSOR_44241_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44241,
  case = "sham",
  environment = "indoor",
  startDate = "2022-10-28",
  endDate = "2022-11-16"
)

# Sensor 6
SENSOR_44241_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44241,
  case = "sham",
  environment = "outdoor",
  startDate = "2022-10-28",
  endDate = "2022-11-16"
)

# Sensor 7
SENSOR_44241_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44241,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-12-20",
  endDate = "2023-01-06"
)

# Sensor 8
SENSOR_44241_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44241,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-12-20",
  endDate = "2023-01-06"
)

# Sensor 9
SENSOR_41271_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00126",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41271,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-10-28",
  endDate = "2022-11-04"
)

# Sensor 10
SENSOR_41271_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41271,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-10-28",
  endDate = "2022-11-04"
)

# Sensor 11
SENSOR_41271_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00114",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41271,
  case = "sham",
  environment = "indoor",
  startDate = "2023-01-06",
  endDate = "2023-01-13"
)

# Sensor 12
SENSOR_41271_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00126",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41271,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-01-06",
  endDate = "2023-01-13"
)

# Sensor 13
SENSOR_42231_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00126",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42231,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-11-04",
  endDate = "2022-11-17"
)

# Sensor 14
SENSOR_42231_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42231,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-11-04",
  endDate = "2022-11-17"
)

# Sensor 15
SENSOR_42231_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42231,
  case = "sham",
  environment = "indoor",
  startDate = "2023-01-06",
  endDate = "2023-01-18"
)

# Sensor 16
SENSOR_42231_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42231,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-01-06",
  endDate = "2023-01-18"
)

# Sensor 17
SENSOR_41211_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00716",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41211,
  case = "sham",
  environment = "indoor",
  startDate = "2022-11-10",
  endDate = "2022-11-16"
)

# Sensor 18
SENSOR_41211_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00718",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41211,
  case = "sham",
  environment = "outdoor",
  startDate = "2022-11-10",
  endDate = "2022-11-16"
)

# Sensor 19
SENSOR_41211_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00125",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41211,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-01-06",
  endDate = "2023-01-13"
)

# Sensor 20
SENSOR_41211_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41211,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-01-06",
  endDate = "2023-01-13"
)

# Sensor 21
SENSOR_42321_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00125",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42321,
  case = "sham",
  environment = "indoor",
  startDate = "2022-11-28",
  endDate = "2022-12-05"
)

# Sensor 22
SENSOR_42321_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42321,
  case = "sham",
  environment = "outdoor",
  startDate = "2022-11-28",
  endDate = "2022-12-05"
)

# Sensor 23
SENSOR_42321_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42321,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-01-25",
  endDate = "2023-02-01"
)

# Sensor 24
SENSOR_42321_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42321,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-01-25",
  endDate = "2023-02-01"
)

# Sensor 25
SENSOR_42281_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42281,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-12-07",
  endDate = "2022-12-14"
)

# Sensor 26
SENSOR_42281_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00125",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42281,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-12-07",
  endDate = "2022-12-14"
)

# Sensor 27
SENSOR_42281_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42281,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-02-08",
  endDate = "2023-02-15"
)

# Sensor 28
SENSOR_42281_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42281,
  case = "sham",
  environment = "indoor",
  startDate = "2023-02-08",
  endDate = "2023-02-15"
)

# Sensor 29
SENSOR_41421_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00125",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41421,
  case = "sham",
  environment = "outdoor",
  startDate = "2022-12-14",
  endDate = "2022-12-21"
)

# Sensor 30
SENSOR_41421_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 41421,
  case = "sham",
  environment = "indoor",
  startDate = "2022-12-14",
  endDate = "2022-12-21"
)

########################################

# Sensor 31
SENSOR_45361_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00125",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45361,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-12-14",
  endDate = "2022-12-20"
)

########################################

# Sensor 32
SENSOR_45361_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45361,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-12-14",
  endDate = "2022-12-20"
)

########################################

# Sensor 33
SENSOR_45361_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45361,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-02-09",
  endDate = "2023-02-15"
)

########################################

# Sensor 34
SENSOR_45361_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45361,
  case = "sham",
  environment = "indoor",
  startDate = "2023-02-09",
  endDate = "2023-02-15"
)

########################################

# Sensor 35
SENSOR_42301_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-000114",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42301,
  case = "hepa",
  environment = "outdoor",
  startDate = "2022-12-13",
  endDate = "2022-12-20"
)

########################################

# Sensor 36
SENSOR_42301_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00715",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42301,
  case = "hepa",
  environment = "indoor",
  startDate = "2022-12-13",
  endDate = "2022-12-20"
)

########################################

# Sensor 37
SENSOR_42301_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42301,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-02-09",
  endDate = "2023-02-15"
)

########################################

# Sensor 38
SENSOR_42301_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 42301,
  case = "sham",
  environment = "indoor",
  startDate = "2023-02-09",
  endDate = "2023-02-15"
)

########################################

# Sensor 39
SENSOR_45411_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45411,
  case = "sham",
  environment = "indoor",
  startDate = "2023-02-22",
  endDate = "2023-03-01"
)

########################################

# Sensor 40
SENSOR_45411_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45411,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-02-22",
  endDate = "2023-03-01"
)

########################################

# Sensor 41
SENSOR_45411_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45411,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-04-14",
  endDate = "2023-04-21"
)

########################################

# Sensor 42
SENSOR_45411_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45411,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-04-14",
  endDate = "2023-04-21"
)

########################################

# Sensor 43
SENSOR_44621_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44621,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-03-22",
  endDate = "2023-03-29"
)

########################################

# Sensor 44
SENSOR_44621_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44621,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-03-22",
  endDate = "2023-03-29"
)

########################################

# Sensor 45
SENSOR_44621_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44621,
  case = "sham",
  environment = "indoor",
  startDate = "2023-05-11",
  endDate = "2023-05-18"
)

########################################

# Sensor 46
SENSOR_44621_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 44621,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-05-11",
  endDate = "2023-05-18"
)

########################################

# Sensor 47
SENSOR_45451_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45451,
  case = "sham",
  environment = "indoor",
  startDate = "2023-03-31",
  endDate = "2023-04-07"
)

########################################

# Sensor 48
SENSOR_45451_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45451,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-03-31",
  endDate = "2023-04-07"
)

########################################

# Sensor 49
SENSOR_45451_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45451,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-05-18",
  endDate = "2023-05-24"
)

########################################

# Sensor 50
SENSOR_45451_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00104",
  locationCode = "OH",
  sensorCode = "M",
  ID = 45451,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-05-18",
  endDate = "2023-05-24"
)

########################################

# Sensor 51
SENSOR_46591_sham_indoor <- Sensor(
  sensorNum = "MOD-PM-00099",
  locationCode = "OH",
  sensorCode = "M",
  ID = 46591,
  case = "sham",
  environment = "indoor",
  startDate = "2023-04-05",
  endDate = "2023-04-12"
)

########################################

# Sensor 52
SENSOR_46591_sham_outdoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 46591,
  case = "sham",
  environment = "outdoor",
  startDate = "2023-04-05",
  endDate = "2023-04-12"
)

########################################

# Sensor 53
SENSOR_46591_hepa_indoor <- Sensor(
  sensorNum = "MOD-PM-00111",
  locationCode = "OH",
  sensorCode = "M",
  ID = 46591,
  case = "hepa",
  environment = "indoor",
  startDate = "2023-05-19",
  endDate = "2023-05-26"
)

########################################

# Sensor 54
SENSOR_46591_hepa_outdoor <- Sensor(
  sensorNum = "MOD-PM-00128",
  locationCode = "OH",
  sensorCode = "M",
  ID = 46591,
  case = "hepa",
  environment = "outdoor",
  startDate = "2023-05-19",
  endDate = "2023-05-26"
)
```

## RETTRIEVING DATA INTO FOLDERS

```{r}
# Create folder if not initiated
data_folder_path <- "./data/HAFTRAP/OH/modpm/"

# Check if the folder exists
if (!file.exists(data_folder_path)) {
  # Create the folder if it doesn't exist
  dir.create(data_folder_path, recursive = TRUE)
}

# Function to process a single device ID
process_device_id <- function(device_id) {
  for (date in time_range) {
    for (state in c("raw/", "")) {
      # Construct device URL for API call
      device_url <-
        paste0(device_id, "/data-by-date/", state, date, "/")
      # Send GET request
      raw_data <- GET(
        url = paste0(base_url, device_url),
        authenticate(api_key, "", type = "basic"),
        encoding = "UTF-8"
      )
      
      # Option 1: Un-comment this if you want to just ignore the faulty API responses
      # Check the HTTP status code of the response
      if (status_code(raw_data) != 200) {
        # Skip processing if the response is not successful (not 200)
        return()
      }
      
      # # Option 2: Un-comment this if you want to see the faulty API responses in error_log.txt
      # # Check the content type of the response
      # content_type <- httr::headers(raw_data)$`content-type`
      # if (!grepl("application/json", content_type, ignore.case = TRUE)) {
      #   # Handle the error
      #   error_msg <-
      #     paste0("Error: Unexpected content type for", device_id, "on", date)
      #   # Store the raw response or handle it as needed
      #   raw_response <-
      #     content(raw_data, as = "text", encoding = "UTF-8")
      #
      #   # Write the error message and raw response to a log file
      #   log_entry <- paste0(error_msg, "\n", raw_response, "\n\n")
      #   write(log_entry, file = "error_log.txt", append = TRUE)
      #
      #   return()
      # }
      
      data <- content(raw_data, as = "text", encoding = "UTF-8")
      parsed_data <- fromJSON(data)
      
      # Check if the data is empty or has zero rows
      if (is.null(parsed_data) ||
          is.null(parsed_data$data) || length(parsed_data$data) == 0) {
        # Skip processing if the data is empty
        return()
      }
      
      # Extract relevant data
      df <- as.data.frame(parsed_data$data)
    }
  }
  
  # Define device folder path
  device_folder_path <- file.path(data_folder_path, device_id)
  
  # Check if the device folder exists
  if (!dir.exists(device_folder_path)) {
    # Create the device folder if it doesn't exist
    dir.create(device_folder_path, recursive = TRUE)
  }
  
  # Save data to CSV file
  if (state == "raw/") {
    csv_file_path <-
      file.path(device_folder_path, paste0(device_id, "-", date, "-raw.csv"))
  }
  else {
    csv_file_path <-
      file.path(device_folder_path,
                paste0(device_id, "-", date, "-final.csv"))
  }
  
  # Try writing the CSV file and handle errors
  tryCatch({
    write.csv(df, 
              file = csv_file_path,
              row.names = FALSE,
              fileEncoding = "UTF-8")
  },
  error = function(e) {
    cat(
      paste0("Error writing CSV file for \n"),
      file = "error_log.txt",
      append = TRUE
    )
  })
}

# Process each device ID over time range
for (device_id in device_ids) {
  process_device_id(device_id, date, state)
  # Delete the device folder if it is empty
  device_folder_path <- file.path(data_folder_path, device_id)
  if (dir.exists(device_folder_path) &&
      length(list.files(device_folder_path)) == 0) {
    unlink(device_folder_path, recursive = TRUE)
  }
}
```