---
title: "API Call"
author: "Cherry Pham"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file is created with the purpose of obtaining sensor data
from Quant-AQ through API calls.

There is no prerequisites for using this file. If you are just looking
to quickly retrieve data, you can simply run this file through the
command `Ctrl + Alt + R`. The result of running this file is a data
folder containing data of MOD-PM sensors as of February 2023. If you
wish to modify the data to be retrieved, please read the data guide
below. The original file is written for ambient data in [this
repository](https://github.com/airpartners/ambient-summer23), which
might be more up-to-date.

## DATA GUIDE

Brief explanation of the code:

A. **PACKAGES AND INITIAL SETTINGS**

B. **DEFINE VARIABLES**

-   You can find sensor information, API key and API url here C.

C. **RETTRIEVING DATA INTO FOLDERS**

1.  Check if the data folder exists and creates it if it doesn't.

2.  `process_sensor` function:

    a.  Generate a date range based on the start and end dates specific
        to each sensor deployment.

    b.  Loop over the date range and makes API calls for each date.

    c.  The API response is checked for errors and processed
        accordingly.

    d.  After processing all the dates for a sensor, merge the raw and
        final data based on the "timestamp_local" column.

    e.  Flatten merged data by unnesting any nested data frames within
        it.

    f.  Attempt to write the flattened data to a CSV file using the
        generated file path. If an error occurs during the writing
        process, append an error message to the "error_log.txt" file.

3.  Loop over each sensor deployment in the list `sensors` to apply
    `process_sensor` to every deployment.

## PACKAGES AND INITIAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("httr", "jsonlite", "purrr", "dplyr", "tidyverse", "tidyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages
invisible(sapply(packages, library, character.only = TRUE))

# Explicit package function calls
httr::GET
httr::content
jsonlite::fromJSON
purrr::map_df
dplyr::as_data_frame
dplyr::filter
dplyr::lag

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DEFINE VARIABLES

```{r}
# Define API info
base_url <- "https://api.quant-aq.com/device-api/v1/devices/"
api_key <-
  "VUOLNUDTM70QZ7Z9G0Z78XA7" # Replace this API key with yours if can

# Define the Sensor class
setClass(
  "Sensor",
  slots = c(
    sensorNum = "character",
    locationCode = "character",
    sensorCode = "character",
    ID = "numeric",
    case = "character",
    environment = "character",
    startDate = "character",
    endDate = "character"
  )
)

# Constructor function for the Sensor class
Sensor <-
  function(sensorNum,
           locationCode,
           sensorCode,
           ID,
           case,
           environment,
           startDate,
           endDate) {
    new(
      "Sensor",
      sensorNum = sensorNum,
      locationCode = locationCode,
      sensorCode = sensorCode,
      ID = ID,
      case = case,
      environment = environment,
      startDate = startDate,
      endDate = endDate
    )
  }

# Create Sensor objects corresponding to the table data
# Sensor 1
sensors <- list(
  SENSOR_41181_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41181,
    case = "sham",
    environment = "indoor",
    startDate = "2022-10-21",
    endDate = "2022-10-28"
  ),

  # Sensor 2
  SENSOR_41181_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41181,
    case = "sham",
    environment = "outdoor",
    startDate = "2022-10-21",
    endDate = "2022-10-28"
  ),

  # Sensor 3
  SENSOR_41181_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41181,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-12-09",
    endDate = "2022-12-16"
  ),

  # Sensor 4
  SENSOR_41181_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00126",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41181,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-12-09",
    endDate = "2022-12-16"
  ),

  # Sensor 5
  SENSOR_44241_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44241,
    case = "sham",
    environment = "indoor",
    startDate = "2022-10-28",
    endDate = "2022-11-16"
  ),

  # Sensor 6
  SENSOR_44241_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44241,
    case = "sham",
    environment = "outdoor",
    startDate = "2022-10-28",
    endDate = "2022-11-16"
  ),

  # Sensor 7
  SENSOR_44241_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44241,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-12-20",
    endDate = "2023-01-06"
  ),

  # Sensor 8
  SENSOR_44241_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44241,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-12-20",
    endDate = "2023-01-06"
  ),

  # Sensor 9
  SENSOR_41271_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00126",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41271,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-10-28",
    endDate = "2022-11-04"
  ),

  # Sensor 10
  SENSOR_41271_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41271,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-10-28",
    endDate = "2022-11-04"
  ),

  # Sensor 11
  SENSOR_41271_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00114",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41271,
    case = "sham",
    environment = "indoor",
    startDate = "2023-01-06",
    endDate = "2023-01-13"
  ),

  # Sensor 12
  SENSOR_41271_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00126",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41271,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-01-06",
    endDate = "2023-01-13"
  ),

  # Sensor 13
  SENSOR_42231_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00126",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42231,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-11-04",
    endDate = "2022-11-17"
  ),

  # Sensor 14
  SENSOR_42231_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42231,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-11-04",
    endDate = "2022-11-17"
  ),

  # Sensor 15
  SENSOR_42231_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42231,
    case = "sham",
    environment = "indoor",
    startDate = "2023-01-06",
    endDate = "2023-01-18"
  ),

  # Sensor 16
  SENSOR_42231_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42231,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-01-06",
    endDate = "2023-01-18"
  ),

  # Sensor 17
  SENSOR_41211_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00716",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41211,
    case = "sham",
    environment = "indoor",
    startDate = "2022-11-10",
    endDate = "2022-11-16"
  ),

  # Sensor 18
  SENSOR_41211_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00718",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41211,
    case = "sham",
    environment = "outdoor",
    startDate = "2022-11-10",
    endDate = "2022-11-16"
  ),

  # Sensor 19
  SENSOR_41211_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00125",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41211,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-01-06",
    endDate = "2023-01-13"
  ),

  # Sensor 20
  SENSOR_41211_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41211,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-01-06",
    endDate = "2023-01-13"
  ),

  # Sensor 21
  SENSOR_42321_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00125",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42321,
    case = "sham",
    environment = "indoor",
    startDate = "2022-11-28",
    endDate = "2022-12-05"
  ),

  # Sensor 22
  SENSOR_42321_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42321,
    case = "sham",
    environment = "outdoor",
    startDate = "2022-11-28",
    endDate = "2022-12-05"
  ),

  # Sensor 23
  SENSOR_42321_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42321,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-01-25",
    endDate = "2023-02-01"
  ),

  # Sensor 24
  SENSOR_42321_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42321,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-01-25",
    endDate = "2023-02-01"
  ),

  # Sensor 25
  SENSOR_42281_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42281,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-12-07",
    endDate = "2022-12-14"
  ),

  # Sensor 26
  SENSOR_42281_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00125",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42281,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-12-07",
    endDate = "2022-12-14"
  ),

  # Sensor 27
  SENSOR_42281_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42281,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-02-08",
    endDate = "2023-02-15"
  ),

  # Sensor 28
  SENSOR_42281_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42281,
    case = "sham",
    environment = "indoor",
    startDate = "2023-02-08",
    endDate = "2023-02-15"
  ),

  # Sensor 29
  SENSOR_41421_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00125",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41421,
    case = "sham",
    environment = "outdoor",
    startDate = "2022-12-14",
    endDate = "2022-12-21"
  ),

  # Sensor 30
  SENSOR_41421_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 41421,
    case = "sham",
    environment = "indoor",
    startDate = "2022-12-14",
    endDate = "2022-12-21"
  ),

  # Sensor 31
  SENSOR_45361_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00125",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45361,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-12-14",
    endDate = "2022-12-20"
  ),

  # Sensor 32
  SENSOR_45361_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45361,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-12-14",
    endDate = "2022-12-20"
  ),

  # Sensor 33
  SENSOR_45361_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45361,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-02-09",
    endDate = "2023-02-15"
  ),

  # Sensor 34
  SENSOR_45361_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45361,
    case = "sham",
    environment = "indoor",
    startDate = "2023-02-09",
    endDate = "2023-02-15"
  ),

  # Sensor 35
  SENSOR_42301_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-000114",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42301,
    case = "hepa",
    environment = "outdoor",
    startDate = "2022-12-13",
    endDate = "2022-12-20"
  ),

  # Sensor 36
  SENSOR_42301_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00715",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42301,
    case = "hepa",
    environment = "indoor",
    startDate = "2022-12-13",
    endDate = "2022-12-20"
  ),

  # Sensor 37
  SENSOR_42301_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42301,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-02-09",
    endDate = "2023-02-15"
  ),

  # Sensor 38
  SENSOR_42301_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 42301,
    case = "sham",
    environment = "indoor",
    startDate = "2023-02-09",
    endDate = "2023-02-15"
  ),

  # Sensor 39
  SENSOR_45411_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45411,
    case = "sham",
    environment = "indoor",
    startDate = "2023-02-22",
    endDate = "2023-03-01"
  ),

  # Sensor 40
  SENSOR_45411_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45411,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-02-22",
    endDate = "2023-03-01"
  ),

  # Sensor 41
  SENSOR_45411_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45411,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-04-14",
    endDate = "2023-04-21"
  ),

  # Sensor 42
  SENSOR_45411_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45411,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-04-14",
    endDate = "2023-04-21"
  ),

  # Sensor 43
  SENSOR_44621_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44621,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-03-22",
    endDate = "2023-03-29"
  ),
  # Sensor 44
  SENSOR_44621_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44621,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-03-22",
    endDate = "2023-03-29"
  ),

  # Sensor 45
  SENSOR_44621_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44621,
    case = "sham",
    environment = "indoor",
    startDate = "2023-05-11",
    endDate = "2023-05-18"
  ),

  # Sensor 46
  SENSOR_44621_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 44621,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-05-11",
    endDate = "2023-05-18"
  ),

  # Sensor 47
  SENSOR_45451_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45451,
    case = "sham",
    environment = "indoor",
    startDate = "2023-03-31",
    endDate = "2023-04-07"
  ),

  # Sensor 48
  SENSOR_45451_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45451,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-03-31",
    endDate = "2023-04-07"
  ),

  # Sensor 49
  SENSOR_45451_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45451,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-05-18",
    endDate = "2023-05-24"
  ),

  # Sensor 50
  SENSOR_45451_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00104",
    locationCode = "OH",
    sensorCode = "M",
    ID = 45451,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-05-18",
    endDate = "2023-05-24"
  ),

  # Sensor 51
  SENSOR_46591_sham_indoor = Sensor(
    sensorNum = "MOD-PM-00099",
    locationCode = "OH",
    sensorCode = "M",
    ID = 46591,
    case = "sham",
    environment = "indoor",
    startDate = "2023-04-05",
    endDate = "2023-04-12"
  ),

  # Sensor 52
  SENSOR_46591_sham_outdoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 46591,
    case = "sham",
    environment = "outdoor",
    startDate = "2023-04-05",
    endDate = "2023-04-12"
  ),

  # Sensor 53
  SENSOR_46591_hepa_indoor = Sensor(
    sensorNum = "MOD-PM-00111",
    locationCode = "OH",
    sensorCode = "M",
    ID = 46591,
    case = "hepa",
    environment = "indoor",
    startDate = "2023-05-19",
    endDate = "2023-05-26"
  ),

  # Sensor 54
  SENSOR_46591_hepa_outdoor = Sensor(
    sensorNum = "MOD-PM-00128",
    locationCode = "OH",
    sensorCode = "M",
    ID = 46591,
    case = "hepa",
    environment = "outdoor",
    startDate = "2023-05-19",
    endDate = "2023-05-26"
))

```

## RETTRIEVING DATA INTO FOLDERS

```{r}
# Create folder if not initiated
data_folder_path <- "./data/HAFTRAP/OH/modpm/"

# Check if the folder exists and create it if it doesn't
if (!file.exists(data_folder_path)) {
  dir.create(data_folder_path, recursive = TRUE)
}

# Function to process a single sensor
process_sensor <- function(sensor) {
  merged_data_per_sensor <- data.frame()
  time_range <-
    seq(as.Date(sensor@startDate), as.Date(sensor@endDate), by = "day")
  time_range <- format(time_range, "%Y-%m-%d")

  for (date in time_range) {
    raw_df <- data.frame()
    final_df <- data.frame()
    merged_data_per_day <- data.frame()
    for (state in c("raw/", "")) {
      sensor_url <-
        paste0(sensor@sensorNum, "/data-by-date/", state, date, "/")
      api_data <- GET(url = paste0(base_url, sensor_url),
                      authenticate(api_key, "", type = "basic"),
        encoding = "UTF-8"
      )
      
      # Check the HTTP status code of the response
      if (status_code(api_data) != 200) {
        # Skip processing if the response is not successful (not 200)
        next
      }
      
      data <- content(api_data, as = "text", encoding = "UTF-8")
      parsed_data <- fromJSON(data)
      
      # Check if the data is empty or has zero rows
      if (is.null(parsed_data) ||
          is.null(parsed_data$data) || length(parsed_data$data) == 0) {
        
        # Skip processing if the data is empty
        next
      }
      
      if (state == "raw/") {
        raw_df <- as.data.frame(parsed_data$data)
      } else {
        final_df <- as.data.frame(parsed_data$data)
      }
    }
    
    # Skip processing if the merged data is empty or has zero rows
    if (is.null(raw_df) ||
        is.null(final_df) || nrow(raw_df) == 0 || nrow(final_df) == 0) {
      
      next
    }
    
    # Merge raw and final data
    merged_data_per_day <- merge(raw_df, final_df, by = "timestamp_local")
  }
  
  # Skip processing if the merged data is empty or has zero rows
  if (is.null(merged_data_per_day) || nrow(merged_data_per_day) == 0) {
    return()
  }
  # Flatten any nested dataframes
  cols_to_unnest <-
    names(merged_data_per_day)[sapply(merged_data_per_day, is.data.frame)]
  flattened_df <-
    unnest(merged_data_per_day, cols = cols_to_unnest, names_sep = ".")
  merged_data_per_sensor <- rbind(merged_data_per_sensor, flattened_df)
  
  # Save data to CSV file using naming convention from data_guide
  csv_file_path <- file.path(
    data_folder_path,
    paste0(
      sensor@locationCode,
      "_",
      sensor@sensorCode,
      "_",
      sensor@ID,
      "_",
      sensor@case,
      "_",
      sensor@environment,
      ".csv"
    )
  )
  
  tryCatch({
    write.csv(merged_data_per_sensor, csv_file_path)
  },
  error = function(e) {
    cat(
      paste0(
        "Error writing ",
        sensor@locationCode,
        "_",
        sensor@sensorCode,
        "_",
        sensor@ID,
        "_",
        sensor@case,
        "_",
        sensor@environment,
        ".csv\n"
      ),
      file = "error_log.txt",
      append = TRUE
    )
  })
}

# Process each sensor
for (sensor in sensors) {
  process_sensor(sensor)
}
```
