# Data Extraction and Cleaning
```{r}
# Imports
library(tidyverse)
library(openair)
library(corrplot)
```

```{r}
# Define indoor file to be imported, set time of HEPA purifier install
indoor_file <- 'data/CM_250_indoor.csv'
outdoor_file <- 'data/CM_250_outdoor.csv'
hepa_time <- "2023-02-10 15:30:00"
```

## Import and prepare data
```{r}
# Import indoor data
indoor_data <- read.csv(indoor_file)

# Import outdoor data
outdoor_data <- read.csv(outdoor_file)

```

```{r}
# Reformat timestamps to time object

indoor_data$date <- as.POSIXct(strptime(indoor_data$timestamp_local, format = "%Y-%m-%dT%H:%M:%SZ", tz = "America/New_York"))

outdoor_data$date <- as.POSIXct(strptime(outdoor_data$timestamp_local, format = "%Y-%m-%dT%H:%M:%SZ", tz = "America/New_York"))
```

```{r}
# # Remove duplicate rows based on date
# indoor_data <- distinct(indoor_data, indoor_data$date, .keep_all = TRUE)
# outdoor_data <- distinct(outdoor_data, outdoor_data$date, .keep_all = TRUE)

```

```{r}
# Append all variables with "outdoor_" or "indoor_" respectively
colnames(indoor_data) <- paste0('indoor_', colnames(indoor_data))
colnames(outdoor_data) <- paste0('outdoor_', colnames(outdoor_data))
```

```{r}
# Time-sync indoor, outdoor data
# Join dataframes by syncing indoor_date and outdoor_date columns
joined_data <- left_join(indoor_data, outdoor_data, by = c("indoor_date" = "outdoor_date"))

# Renaming synced date columns back to just 'date'
joined_data$date <- joined_data$indoor_date

# Keeping distinct time-stamps
joined_data %>% distinct(joined_data$indoor_date, .keep_all = TRUE)

# # Attach for easier reference while graphing
# attach(joined_data)

```

```{r}
outdoor_data <- outdoor_data %>% rename_at('outdoor_date', ~'date')
```

```{r}
# #Initial time series
# timePlot(joined_data, pollutant = c('indoor_pm1', 'indoor_pm25', 'indoor_pm10'), normalize = TRUE)

timePlot(joined_data, pollutant = c('outdoor_pm1', 'outdoor_pm25', 'outdoor_pm10'), normalize = TRUE)
```


```{r}
# Indoor data timevariation
# timeVariation(joined_data, pollutant = c("indoor_pm1", "indoor_pm25", "indoor_pm10"))

# Outdoor data timevariation
timeVariation(outdoor_data, pollutant = c("outdoor_pm1", "outdoor_pm25", "outdoor_pm10"))


```


```{r}
indoor_cordata <- select(joined_data, indoor_pm1, indoor_pm25, indoor_pm10)
N <- cor(na.omit(indoor_cordata), use = "p")
corrplot(N, method = 'number')
```

```{r}
outdoor_cordata <- select(outdoor_data, outdoor_pm1, outdoor_pm25, outdoor_pm10)
N <- cor(na.omit(outdoor_cordata), use = "p")
corrplot(N, method = 'number')
```


## Questions for Scott
1) What is sample_rh, sample_pres in mod_PM dataset
2) Do we filter for rows where device is only ACTIVE
3) No duplicate rows so far - do we need that code chunk?
4) Committing data files to github- is this a good idea and what are alternatives?
5) How does dataframe merging work if the timestamps are continuous and not exactly synced?? Aren't most values just going to be NAN then? (Update: joined data does not plot)

