# Summarize HOBO Data

## STOP

**HAVE YOU RUN `MF_hobo_cleaning`?**
*This file loads a dataframe created from running `MF_hobo_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to summarize proportion of time HEPA air purifiers are switched on using HOBO data.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```


## Load data
```{r}
# Load HOBO data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_hobo.RData"))
```

## Find means
```{r}
main_df %>%
  filter(case == "after") %>%
  group_by(participant_id, room) %>%
  summarise(prop_on = sum(power > 10)/n()) -> props_df
```
## Save to file
Summary statistics are saved in a CSV file.
```{r}
# Save summary statistic files to CSVs
write.csv(props_df, "summary/EastBoston/MF/s_MF_H_props.csv")
```