# Mothers Out Front Merging Mod-PM + HOBO

## STOP

**HAVE YOU RUN `MF_modpm_cleaning` AND `MF_hobo_cleaning`?**
*This file loads dataframes created from running `MF_modpm_cleaning` AND `MF_hobo_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script merges Mod-PM and HOBO data into one dataframe to plot them concurrently

As a by-product to calculating correlation coefficients, this script also generates a data-frame stored in `merged.RData`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(gridExtra)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs
```{r}
# Get relevant participant IDs from MF_participants
participants <- c("2", "3", "5", "6", "7", "8", "11")
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```

## Load data
Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))

modpm_df <- main_df
rm(main_df)
```

HOBO
```{r}
# Load HOBO data from RData file
load(paste0(path_to_data, "cleaned_hobo.RData"))

hobo_df <- main_df
# Delete original
rm(main_df)
```



## Main Code Run
### Plot
```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
# Re-try if incorrecto
while (!(person %in% participants)) {
  person <- readline("Error! Invalid ID, Re-enter: ")
}

plot_list <- list()

df <- modpm_df %>% filter(participant_id == person, environment == "indoor")

rooms <- unique(df$room)

if (length(rooms) > 1) {
  for (i in seq_along(rooms)) {
    r <- rooms[[i]]
    modpm_plot <-  df %>%
      filter(room == r) %>%
      ggplot() +
        geom_line(aes(x = date, y = pm25), color = "#0072B2") +
        scale_y_continuous(limits = c(0, 200)) +
        # Set axes labels, title
        labs(title = paste("PM 2.5 Concentration for Participant", person),
             y = "Particle Concentration") + 
        theme(axis.text.x = element_blank(), axis.title.x = element_blank())
    
    plot_list[[i]] <- modpm_plot
  }
} else {
  plot_list[[1]] <- ggplot(df) + 
    geom_line(aes(x = date, y = pm25), color = "#0072B2") +
    # Set axes labels, title
    labs(title = paste("PM 2.5 Concentration vs. HEPA Purifier Power Usage"),
         y = "Particle Concentration") + 
    theme(axis.text.x = element_blank(), axis.title.x = element_blank())
}

hobo_plot <- hobo_df %>%
  filter(participant_id == person) %>%
  ggplot() +
    geom_line(aes(x = date, y = power), color = "#D55E00") +
    # Set axes labels
    labs(x = "Date", y = "Power")

plot_list[[length(rooms) + 1]] <- hobo_plot
```

```{r}
grid.arrange(grobs = plot_list, ncol = 1)
```

