# Mothers Out Front Merging Mod-PM + HOBO

## STOP

**HAVE YOU RUN `MF_modpm_cleaning` AND `MF_hobo_cleaning`?**
*This file loads dataframes created from running `MF_modpm_cleaning` AND `MF_hobo_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script merges Mod-PM and HOBO data into one dataframe to plot them concurrently

As a by-product to calculating correlation coefficients, this script also generates a data-frame stored in `merged.RData`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(gridExtra)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```

## Load data
Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))

modpm_df <- main_df
rm(main_df)
```

HOBO
```{r}
# Load HOBO data from RData file
load(paste0(path_to_data, "cleaned_hobo.RData"))

hobo_df <- main_df
# Delete original
rm(main_df)
```



## Main Code Run
### Plot
```{r}
modpm_plot <- modpm_df %>%
  filter(participant_id == "3", environment == "indoor") %>%
  ggplot() +
    geom_line(aes(x = date, y = pm25))
```
```{r}
hobo_plot <- hobo_df %>%
  filter(participant_id == "3") %>%
  ggplot() +
    geom_line(aes(x = date, y = power))
```
```{r}
grid.arrange(modpm_plot, hobo_plot, ncol = 1)
```

