# Compare Kitchen and Play area for Participant 6

## STOP

**HAVE YOU RUN `MF_modpm_cleaning`?**
*This file loads a dataframe created from running `MF_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to see if there is any correlation between the air particles in two rooms in the same house, one of them is a kitchen.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```


## Load data
```{r}
# Load Mod-PM data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Filter
```{r}
df <- main_df %>%
      filter(participant_id == "6", case == "before", 
             environment == "indoor")
```

## Separate Kitchen, Play Room
```{r}
df_kitchen <- df %>% filter(room == "kitchen")
df_play <- df %>% filter(room == "play")
```

## Merge dataframes
```{r}
df_merged <- inner_join(df_kitchen, df_play, suffix = c("_kitchen", "_play"),
                   by = join_by(date, participant_id, environment, case))
```

```{r}
df_merged %>%
  mutate(time_round = as.integer(hour(round_date(date, "4 hours")))) %>%
  filter(time_round %in% c(8, 16, 20)) %>%
  ggplot(aes(x = pm25_kitchen, y = pm25_play, color = factor(time_round))) +
    # Scatter-plot
    geom_point(alpha = 0.1) +
    # # Lines of best fit
    # geom_smooth(method=lm, se=FALSE) +
    scale_color_manual(values = rainbow(6)) +
    ylim(0, 200) +
    xlim(0, 200)
  
```


