# East Boston Plume Plotting (for Mod-PM)
Plot various plumes before and after HEPA instalation to compare their heights,
widths, decay constants, etc.

## STOP

**HAVE YOU RUN `MF_modpm_cleaning`?**
*This file loads a dataframe created from running `MF_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to analyze decay constants and complete plume visualizations
for the East Boston Mod-PM data.

```{r}
# import necessary libraries
library(tidyverse)
library(data.table)
library(scales)
library(pracma)
```

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = "~/AirPartners/hepa-summer23")
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Main Code Run
```{r}
# Pre-processing to add relevant columns
my_df <- main_df %>%
  # Round date to nearest ten minutes
  mutate(date_round = round_date(date, unit = "10 mins")) %>%
  # Get day of the week as integer from 1 to 7
  mutate(wkdy = wday(date)) %>%
  # Classify as weekday or weekend
  mutate(is_wknd = wkdy %in% c(1, 7)) %>%
  # Extract time of the day from datetime
  mutate(time = as.POSIXct(as.ITime(date_round), format = "%H:%M:%S"))
  
```

```{r}
# Calculate summary statistics for ever ten minutes
graph_main_df <- my_df %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm25), 
               names_to = "particle", values_to = "reading") %>%
  # For every 10 minutes for given case, environment, particle, weekday/end
  group_by(is_wknd, time, case, environment, particle) %>%
  # Find summary statistics
  summarise(mean = mean(reading),
          median = median(reading), 
          q5 = quantile(reading, probs = 0.05), 
          q25 = quantile(reading, probs = 0.25),
          q75 = quantile(reading, probs = 0.75),
          q95 = quantile(reading, probs = 0.95),
          sd = sd(reading),
          .groups = 'drop')
```


# Helper functions for plume analysis
```{r}
# Find the first local minimum before/after the index of peak_row in y
# that occurs below the min_threshold.
get_local_valleys <- function(data, peaks, min_threshold) {
  # Iterate through rows of peaks matrix and get valleys before peaks
  for (row in 1:nrow(peaks)) {
    idx <- peaks[row, 2]
    n <- TRUE
    # create while loop to check for local minima
    while (n) {
      # if we're at the end of y, break out of loop
      if (idx == 0) {
        n <- FALSE
      } else {
        # otherwise, get slope (approximately)
        slope <- data[idx] - data[idx - 1]
        if (data[idx] <= min_threshold && slope < 0) {
          n <- FALSE
        } else {
          idx <- idx - 1
        }
      }
    }
    # set values of matrix so that column 3 is index of valley before peak
    peaks[row, 3] <- idx
  }
  # Repeat process for valleys after peaks
  for (row in 1:nrow(peaks)) {
    idx <- peaks[row, 2]
    n <- TRUE
    # create while loop to check for local minima
    while (n) {
      # if we're at the end of y, break out of loop
      if (idx == length(data)) {
        n <- FALSE
      } else {
        # otherwise, get slope (approximately)
        slope <- data[idx + 1] - data[idx]
        if (data[idx] <= min_threshold && slope > 0) {
          n <- FALSE
        } else {
          idx <- idx + 1
        }
      }
    }
    # set values of matrix so that column 4 is index of valley after peak
    peaks[row, 4] <- idx
  }
  peaks
}
```

```{r}
# Exponential curve fitting for air quality data; returns dataframe containing
# k-values of curves
# NOTE: if the curve is too small or doesn't contain enough information for this
# computation, peak will be omitted from final return
curve_fitting <- function(data, peaks) {
    # Create empty DataFrame for storing k values
  alphas.data <- data.frame(
    "peak_idx" = numeric(0),
    "valley_idx_b" = numeric(0),
    "valley_idx_a" = numeric(0),
    "peak_hgt" = numeric(0),
    "k_val" = numeric(0),
    "conv_tol" = numeric(0),
    "cumm_pm" = numeric(0)
  )
  # Define parameters for curve fitting function for each row
  for (row in 1:nrow(peaks)) {
    row <- as.double(row)
    i_range <- peaks[row, 2]:peaks[row, 4]
    # Check if time is sorted in this range; if not, reverse the indices
    time_check <- data$time[i_range]
    is_sorted_asc <- all(time_check[-1] >= time_check[-length(time_check)])
    if (is_sorted_asc) {
      sect <- rev(data$pm25[i_range])
    } else {
      # sect <- data$pm25[i_range]
      sect <- data$pm25[i_range]
    }
    # print(sect)
    t <- i_range - peaks[row, 2] + 1
    df <- data.frame(t = t, y = sect)
    # Get exponential fit
    nlc <- nls.control(maxiter = 1000)
    fit <- try(nls(y ~ SSasymp(t, yf, y0, log_alpha), data = df, control = nlc))
    if (class(fit) != "nls") {
      next
    }
    # Get parameters of the fit
    params <- coef(fit)
    # Extract the log_alpha value and put it in form e^(log(a)) to get a
    log_alpha <- as.double(params["log_alpha"])
    alpha <- exp(log_alpha)
    # Get achieved convergence tolerance as metric for accuracy of fit
    # NOTE: R^2 value can be calculated but is not a useful metric
    # for nonlinear models
    conv <- fit$convInfo$finTol
    # Add alpha to dataframe
    alphas.newdata <- data.frame(
      "peak_idx" = c(peaks[row, 2]),
      "valley_idx_b" = c(peaks[row, 3]),
      "valley_idx_a" = c(peaks[row, 4]),
      "peak_hgt" = c(peaks[row, 1]),
      "k_val" = c(alpha),
      "conv_tol" = c(conv)
    )
    alphas.data <- rbind(alphas.data, alphas.newdata)
  }
  alphas.data <- arrange_all(alphas.data)
  alphas.data <- alphas.data[order(alphas.data$peak_hgt), ]
}
```


### Main plume analysis
```{r}
# Separate my_df by before vs. after
my_df_before <- my_df[my_df$case=='before',]
my_df_after  <- my_df[my_df$case=='after',]

# Get peaks from before and after
peaks_b <- findpeaks(my_df_before$pm25,
                     nups = 1,
                     ndowns = 2,
                     minpeakheight = 50,
                     minpeakdistance = 200,
                     threshold = 0)
peaks_a <- findpeaks(my_df_after$pm25,
                   nups = 1,
                   ndowns = 2,
                   minpeakheight = 50,
                   minpeakdistance = 200,
                   threshold = 0)
```

```{r}
# Get valleys before and after to isolate each plume
peaks_b <- get_local_valleys(my_df_before$pm25,
                             peaks = peaks_b,
                             min_threshold = 50)
peaks_a <- get_local_valleys(my_df_after$pm25,
                             peaks = peaks_a,
                             min_threshold = 50)
```

```{r}
# Plot peaks and valleys on top of current time series with valleys 
# (sanity check)
plot(my_df_before$pm25,
  type = "l",
  main = "Indoor/Outdoor Ratio",
  xlab = "Time (indices)",
  ylab = "Ratio of indoor/outdoor concentrations",
  col = "navy")
  grid()
points(peaks_b[, 2], peaks_b[, 1], pch = 20, col = "maroon")
points(peaks_b[, 3], my_df_before$pm25[peaks_b[, 3]], pch = 20, col = "yellow")
points(peaks_b[, 4], my_df_before$pm25[peaks_b[, 4]], pch = 20, col = "green")
```

```{r}
# perform decay analysis
decays_before <- curve_fitting(my_df_before, peaks = peaks_b)
decays_before
```

```{r}
# perform decay analysis
decays_after <- curve_fitting(my_df_after, peaks = peaks_a)
decays_after
```


# Displaying peaks with best fits
```{r}
# Before HEPA deployment
peak_idxs <- c(
  tail(decays_before, n=1)$valley_idx_a,
  tail(decays_before, n=1)$valley_idx_b
)
my_df_before %>%
  ggplot(aes(x = as.numeric(row.names(my_df_before)))) +
  # Plot the highest pm25 plume
  geom_line(aes(y = pm25)) +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') + 
  ylab('PM2.5 Levels') +
  ylim(0,1000) +
  # Add fun theme
  theme_bw() +
  # Scale to just around the plume
  coord_cartesian(xlim = peak_idxs)
```

```{r}
# After HEPA deployment
peak_idxs <- c(
  tail(decays_after, n=1)$valley_idx_a,
  tail(decays_after, n=1)$valley_idx_b
)
my_df_after %>%
  ggplot(aes(x = as.numeric(row.names(my_df_after)))) +
  # Plot the highest pm25 plume
  geom_line(aes(y = pm25)) +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') + 
  ylab('PM2.5 Levels') +
  ylim(0,1000) +
  # Add fun theme
  theme_bw() +
  # Scale to just around the plume
  coord_cartesian(xlim = peak_idxs)
```

```{r}
# Merge decay tables, adding a column indicating whether HEPA installed or not
decays_before$hepa <- FALSE
decays_after$hepa <- TRUE
all_decays <- merge(decays_before, decays_after, all = TRUE)
all_decays

# Get index of when HEPA was installed
hepa_install_idx <- match(TRUE, all_decays$hepa)

# Create summary of decay values for before and after
all_decays["hepa_installed"] <- all_decays["peak_idx"] >= hepa_install_idx
all_decays["peak_width"] <- all_decays["valley_idx_a"] - all_decays["valley_idx_b"]

decay_sum <-
  all_decays %>%
    group_by(hepa) %>%
    summarize(
      k_val_mean = mean(k_val, na.rm = TRUE),
      k_val_median = median(k_val, na.rm = TRUE),
      k_val_min = min(k_val, na.rm = TRUE),
      k_val_max = max(k_val, na.rm = TRUE),
      peak_hgt_mean = mean(peak_hgt, na.rm = TRUE),
      peak_hgt_median = median(peak_hgt, na.rm = TRUE),
      peak_wdh_mean = mean(peak_width, na.rm = TRUE)
    )
decay_sum
```

```{r}
peak_width_sum <-
  all_decays %>%
    group_by(hepa) %>%
    summarize(
      peak_wdh_mean = mean(peak_width, na.rm = TRUE),
      peak_wdh_median = median(peak_width, na.rm = TRUE),
      peak_wdh_min = min(peak_width, na.rm = TRUE),
      peak_wdh_max = max(peak_width, na.rm = TRUE),
    )
peak_width_sum
```

TODO: Create a function here for calculating the cumulative PM of a peak and
correlate it to peak widths.
```{r}
get_cumm_PM <- function(decays_data, plume_data) {
  total_pm <- 0
  for (row in 1:nrow(decays_data)) {
    idx_b <- decays_data$valley_idx_b[row]
    idx_a <- decays_data$valley_idx_a[row]
    plume_pm <- sum(plume_data$pm25[idx_b:idx_a])
    total_pm <- total_pm + plume_pm
  }
  total_pm
}
```

```{r}
before_pm <- get_cumm_PM(decays_before, my_df_before)
after_pm <- get_cumm_PM(decays_after, my_df_after)
c(before_pm, after_pm)
```










