# EastBoston Data Summarization - multiple (for Mod-PM)

## STOP

**HAVE YOU RUN `MF_modpm_cleaning`?**
*This file loads a dataframe created from running `MF_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to calculated summary statistics for multiple participants in the East Boston pilot, for Mod-PM deployments by Olin. This is different from `MF_modpm_all_stats` in that it filters out participant 11 (the one anomalous value).

Results can be found in `summary/EastBoston/MF/s_MF_M_quants.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Filter out anomalous value
```{r}
select_df <- main_df %>% filter(participant_id == "11", day(date) != 29)
```

## Calculate summary statistics
```{r}
select_df %>%
  group_by(case, environment) %>%
  summarise(mean_pm1 = mean(pm1)) %>%
  pivot_wider(names_from = case, values_from = mean_pm1) %>%
  mutate(redu = (before - after) / before)
```

```{r}
select_df %>%
  mutate(date_plot = case_when(
    case == "before" ~ date,
    case == "after" ~ date - as.difftime(2, unit = "weeks")
  )) %>%
  filter(environment == "ratio") %>%
  ggplot(aes(x = date_plot, y = pm1, color = case)) +
    geom_line(alpha = 0.5) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
```

