# EastBoston Plotting Boxplots

## STOP

**HAVE YOU RUN `MF_modpm_cleaning`?**
*This file loads dataframes created from running `MF_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used in generating a specific kind of plot that isn't saved in `artifacts`. The original written version uses CPC data for participant 44621  to plot timeseries to look at relationship between mean, median, and 95th percentile.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```

## Load data
CPC - filter data to whole minutes
```{r}
# Load CPC data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

```{r}
# Filter desired participant and environment
plot_df <- main_df %>%
  # Choose participant ID from most CPC Q95 increase in s_MF_c_quants.csv
  filter(participant_id == "3", environment == "indoor") %>%
  select(date, pm25, case)
```

```{r}
# Summary statistics (copied over from MF_modpm_all_stats)
plot_df %>%
  # Drop NaN, inf values for summary calculation
  filter(is.finite(pm25)) %>%
  # Group by participant ID, case, environment, and type
  group_by(case) %>%
  
  # Calculate summary statistics and pipe to variable 'summary'
  summarise(mean = mean(pm25),
            median = median(pm25), 
            q5 = quantile(pm25, probs = 0.05), 
            q25 = quantile(pm25, probs = 0.25),
            q75 = quantile(pm25, probs = 0.75),
            q95 = quantile(pm25, probs = 0.95),
            .groups = 'drop') -> summary

# This is very bad janky code to calculate % reduction, sorry. Prints warning
summary <- rbind(summary, apply(select_if(summary, is.numeric), 2, 
                                function(x) 100 * (x[2] - x[1]) / x[2]))

# Specify order
summary$case <- factor(c("after", "before", "redu"))

# Plot for only before, after
summary_plot <- filter(summary, case %in% c("after", "before"))
```


```{r}
# Store plot in object
my_plot <- plot_df %>%
  # Re-order the case in levels of "before", then "after"
  mutate(across(case, \(x) factor(x, levels=c("before", "after")))) %>%
  
  # Plot time-series with 95th percentile, mean, median
  ggplot() +
    # Plot a line graph of Mod-PM reading vs. date, one line for each case
    geom_line(aes(x = date, y = pm25, color = case)) +
    # Split into two separate graphs in single row (same y axis) by case
    facet_wrap(~case, nrow = 1, scales = "free_x") +
    # Plot the horizontal line for mean
    geom_hline(data = summary_plot, 
               aes(yintercept = mean), linetype = "dashed", alpha = 0.5) +
    # Set the line type for each horizontal line
    scale_linetype_manual(name='Summaries',
                       breaks=c('95th', 'Mean', 'Median'),
                       values=c('Median'='solid',
                                'Mean'='dashed', '95th'='dotted')) +
    # Set axes labels, title
    labs(title = "PM 2.5 Concentration for Participant 3",
         x = "Date", y = "Particle Concentration")

# Show plot
my_plot
```

```{r}
# Show plot zoomed in on smaller y-range
my_plot + ylim(0, 100)
```