# Modulair-PM Data Cleaning
This file cleans, and calculates the I/O ratio for Mod-PM data. Output is saved in a `cleaned_modpm.RData` file that is *untracked* by Github. *All files that depend on using Mod-PM data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```


```{r}
# Set path to data
path_to_data <- "data/EastBoston/LF/modpm/"

# Set data category
data_cat <- "LF_M"
```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (file_name in list.files(path_to_data)) {
  # Read csv
  file_path <- paste0(path_to_data, file_name)
  df <- read_csv(file_path, show_col_types = FALSE)

  # Split file name into different parts
  name_split <- strsplit(file_name, "[_.]")[[1]] %>% head(-1)
  room <- name_split[3]
  
  
  df <- df %>%
    # Select broadly relevant columns
    select(date1min, HEPA_on, starts_with(c("bin", "pm"))) %>%
    # Filter out unnecessary columns (bin12;23, model_id, env, bin0_ratio)
    select(-contains(c(as.character(12:23), "model", "env", "bin0_"))) %>%
    
    # Add the participant ID
    mutate(participant_id = room)

  # Append to main dataframe
  all_df <- rbind(all_df, df)
  
  print(paste("Loaded room", room))
}
```

```{r}
curated_df <- all_df %>% 
  
  # Calculate number count of particles in PM1 size
  mutate(pm1num.indoor = bin0.indoor + bin1.indoor + bin2.indoor,
         pm1num.outdoor = bin0.outdoor + bin1.outdoor + bin2.outdoor,
         pm25num.indoor = pm1num.indoor + bin3.indoor + bin4.indoor + bin5.indoor,
         pm25num.outdoor = pm1num.outdoor + bin3.outdoor + bin4.outdoor + bin5.outdoor,
         pm10num.indoor = pm25num.indoor + bin6.indoor + bin7.indoor + bin8.indoor +
           bin9.indoor + bin10.indoor + bin11.indoor,
         pm10num.outdoor = pm25num.outdoor + bin6.outdoor + bin7.outdoor + bin8.outdoor +
           bin9.outdoor + bin10.outdoor + bin11.outdoor,
         
         # Convert HEPA readings to on or off
         case = case_when(HEPA_on == 0 ~ "off", HEPA_on == 1 ~ "on"), 
         
         # Drop redundant columns
         .keep = "unused") %>%
  
  mutate(pm1num_ratio = pm1num.indoor / pm1num.outdoor,
         pm25num_ratio = pm25num.indoor / pm25num.outdoor,
         pm10num_ratio = pm10num.indoor / pm10num.outdoor) %>%
  
  # Rename date column to standard
  rename(date = date1min) %>%
  
  # Change ratio column names delimiter from _ to .
  rename_with( ~ gsub("_", ".", .x, fixed = TRUE), ends_with("ratio"))
```

```{r}
# Bind together indoor, ratio data
main_df <- curated_df %>%
  
  # Gather dataframe to make environment a category
  pivot_longer(cols = contains("."), 
               names_to = c(".value", "environment"), 
               names_sep = "\\.") %>%
  
  # Remove NAs
  drop_na()
```


### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0("cleaning/EastBoston/LF/", "cleaned_modpm.RData"))
```
