# SCOPE HAFTRAP Data Cleaning and Summaries - multiple

## STOP

**HAVE YOU RUN `SC_all_cleaning`?**
*This file loads dataframes created from running `SC_all_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script calculates correlation coefficients for data collected from select sensors in the SCOPE HAFTRAP study

Results can be found in `summary/HAFTRAP/SC/s_SC_RASTC_corr.csv`.

As a by-product to calculating correlation coefficients, this script also generates a data-frame stored in `merged.RData` for use in `SC_correlations`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
library(readxl)
library(corrplot)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Path to Data
```{r}
# Set path to data
path_to_data <- "cleaning/HAFTRAP/SC/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_everything.RData"))
```


## Main Code Run

### Select Options
The below code chunk will be really the only one you need to change. Select the
participants, cases, environments, and variables you want to calculated the
correlation coefficients for. Don't forget to update the corresponding variable
names!
```{r}
# Select participant IDs from OH_participants
participants <- c("1", "2", "3","4")

# Select case to consider (purifier off, on, or both)
cases <- c("off", "on")

# Select environment to consider (indoor, outdoor, or both)
environments <- c("indoor", "outdoor")

# Select columns to calculate correlations between
col_list <- c("mod_pm25", "mod_pm10", "mod_co", "mod_no", "sems_sum100", 
              "acsm_org", "t200u_no", "t300_co", "mcpc_counts")

# IMPORTANT: Update this accordingly if you change col_list
col_list_names <- c("$PM[2.5]", "$PM[10]", "$CO[modulair]", "$NO[modulair]",
                    "$SEMS[100]", "$Organic[ACSM]", "$NO[teledyne]", 
                    "$CO[teledyne]", "$UFPs[cpc]")
```

### Data processing
Filter and select data according to above specifications
```{r}
# Filter columns by specifications
filtered_df <- main_df %>%
  filter(participant_id %in% participants, case %in% cases, 
         environment %in% environments) %>%
  
  # Select only relevant columns
  select(date, participant_id, case, environment, any_of(col_list))
```

### Calculating 10 minute averages
When correlating data, we want to make sure that there has been enough time for air to mix, therefore we calculate 10-minute averages of the data.
```{r}
# Round by 10 minutes (for air mixing)
rounded_df <- filtered_df %>% 
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean over ten minutes
  summarize_at(col_list, mean, na.rm = TRUE) %>%
  # Drop groups
  ungroup() %>%
  # Rename data column
  rename(date = date_round)
```

### Finally, calculate the correlation matrix
```{r}
# Calculate correlation matrix
corr_matrix <- rounded_df %>% 
  select(any_of(col_list)) %>% 
  cor(use = "complete.obs")

print("Calculated correlation matrix for")
print(paste("Participants:", paste(participants, collapse = ", ")))
print(paste("Environments:", paste(environments, collapse = ", ")))
print(paste("Cases (purifier):", paste(cases, collapse = ", ")))
```
### Plotting
Plot the matrix - feel free to customize
```{r}
# Create copy of matrix for plotting
plot_matrix <- corr_matrix

# Rename the axis labels in the matrix for corrplot
rownames(plot_matrix) <- col_list_names
colnames(plot_matrix) <- col_list_names

# Plot the matrix accordingly
corrplot(plot_matrix, method = "number")
```