# HAFTRAP Data Cleaning and Summaries - multiple (for CPC)

## STOP

**HAVE YOU RUN `SC_all_cleaning`?**
*This file loads dataframes created from running `SC_all_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script calculates correlation coefficients for data collected from all sensors in the SCOPE HAFTRAP study

Results can be found in `summary/HAFTRAP/SC/s_SC_RASTC_corr.csv`.

As a by-product to calculating correlation coefficients, this script also generates a data-frame stored in `merged.RData` for use in `SC_correlations`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Get relevant participant IDs from OH_participants
participants <- c("1", "2", "3","4")

# Set path to data
path_to_data <- "cleaning/HAFTRAP/SC/"
```


## Helper Functions:
### Calculating 10 minute averages
When correlating data, we want to make sure that there has been enough time for air to mix, therefore we calculate 10-minute averages of the data.
```{r}
# Round over 10 minute averages
round_10min <- function(df, vec) {
  df %>%
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean over ten minutes
  summarize(across(all_of(vec), mean), .groups = "drop") %>%
  # Rename data column
  rename(date = date_round)
}
```

### Calculating correlation coefficients
Correlation coefficients tell us a lot about relationships between variables. This function calculates the correlation matrix for a particular case for a single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, case, env, col_1, col_2) {
  df %>%
    filter(case == case, environment == env) %>%
    select(col_1, col_2) %>% 
    filter_all(all_vars(is.finite(.))) %>%
    # Calculate correlations (suppress output message)
    cor() -> cor_mat
    
  data.frame(coefficient = cor_mat[[2]], col1 = col_1, col2 = col_2,
             case = case, environment = env)
}
```

## Load data

Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_everything.RData"))
```



## Main Code Run
### Calculate correlation matrices
Helps better understand the relationship between different variables in sham and true HEPA conditions for each participant. Matrices are reshaped into longer pair format for stacking all the data into single dataframe. (Warning: this code takes a while to run.)

```{r}
# Create dataframe to hold correlation coefficients
corr_df <- data.frame()
strings <- main_df %>% select_if(is.numeric) %>% colnames()
all_cols <- t(combn(strings, 2))

# Loop through each case
for (c in c("off", "on")) {
    # Loop through each environment
    for (env in c("indoor", "outdoor")) {
      # Loop through each pair of columns
      for (i in 1:nrow(all_cols)) {
        df <- main_df %>% get_corr(c, env, 
                                   all_cols[i, 1], all_cols[i, 2])
        corr_df <- rbind(corr_df, df)
    }
    print(paste("Computed correlation coefficients for all", c, env)) 
  }
}

```

## Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(corr_df, "summary/HAFTRAP/SC/s_SC_RASTC_corr.csv")
```