# HAFTRAP Plotting Violinplots

## STOP

**HAVE YOU RUN `SC_all_cleaning`?**
*This file loads dataframes created from running `SC_all_cleaning`. Make sure you run it first (if you haven't already) before running this file.*

This script is used to plot boxplots using all the data from the SCOPE HAFTRAP study.

Plots can be found here and saved in the folder`artifacts/HAFTRAP/SC/violinplot`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Path to data
```{r}
# Set path to data
path_to_data <- "cleaning/HAFTRAP/SC/"
```

## Load data
```{r}
# Load all data from RData file
load(paste0(path_to_data, "cleaned_everything.RData"))
```


```{r}
labels_df <- read.csv("analysis/HAFTRAP/SC/plot_labels.csv", row.names = 1)
labels_list <- setNames(split(labels_df, seq(nrow(labels_df))),
                        rownames(labels_df))
```


## Plotting for Indoor and Outdoor raw concentrations
Define lists for plot titles
```{r}
titles_raw <- list(pm1 = "PM 1 Concentration", pm25 = "PM 2.5 Concentration",
               pm10 = "PM 10 Concentration", pm1num = "PM 1 Particle Count",
               concent = "CPC Particle Count")

titles_ratio <- list(pm1 = "PM 1 I/O Ratio", pm25 = "PM 2.5 I/O Ratio",
               pm10 = "PM 10 I/O Ratio", pm1num = "PM 1 Count I/O Ratio",
               concent = "CPC Count I/O Ratio")
```

## Helper functions:
### Plotting function
The following function plots the raw data for a given dataset
```{r}
plot_raw <- function(df_united, p_type, val) {
  # Filter dataset for particle type, dropping NaNs and Infs
  df_united %>% 
    filter(type == p_type, is.finite(reading)) %>%
    # Order sham then hepa by creating factors
    mutate(across(case, \(x) factor(x, levels=c("off", "on")))) %>%
    
  # Plot filtered indoor data
  ggplot() +
    # Create violinplot
    geom_violin(aes(x = situation, y = reading, fill = case)) +
    # Convert y-axis scale to log
    scale_y_continuous(trans = "log10") +
    # Set title, caption with percentage reduction
    labs(title = paste("Overall", labels_list[[p_type]][["title_raw"]]),
         caption = paste0("Mean reduction = ", round(val, 1), "%")) -> my_plot
  
  # Save plot
  ggsave(filename = paste0("artifacts/HAFTRAP/SC/violinplot/", "SC_", p_type,
                           "_raw.png"), plot = my_plot)
  
  print(paste("Plotted for", labels_list[[p_type]][["title_raw"]]))
  
  # Display plot
  my_plot
}
```

### Percentage reduction calculating
The following function calculates mean percentage reduction for a specified particle type. 
```{r}
# Calculate mean percentage reduction for particle in environment
get_redu <- function(df, env, p_type) {
  vals <- df %>% 
    filter(environment == env, type == p_type) %>%
    group_by(case, participant_id) %>% 
    filter(is.finite(reading)) %>% 
    summarize(my_mean = mean(reading))
  
  # Calculate percentage reduction for each participant and store in vector
  vec <- 100 * (vals[vals$case == "off", 3] - vals[vals$case == "on", 3])/
    vals[vals$case == "off", 3]
  
  # Return mean of vector (indexing cause vector is technically a dataframe)
  mean(vec[[1]])
}
```

### Code Run
```{r}
# Filter, order data appropriately
main_united <- main_df %>%
  filter(environment != "ratio") %>%
  unite("situation", environment, case, sep = ", ", remove = FALSE) %>%

  # Gather dataframe to create single variable for measurements
  pivot_longer(c(mod_pm25, mod_pm10, mod_pm1, mod_o3, mod_no2, mod_no, mod_co,
                 mod_pm1num, sems_sum50, sems_sum100, sems_sum1000, acsm_chl,
                 acsm_nh4, acsm_no3, acsm_org, acsm_so4, acsm_f43, acsm_f44,
                 acsm_f57, acsm_ratio_55_57, t300_co, t200u_no, t200u_no2,
                 t200u_nox, mcpc_counts), 
               names_to = "type", values_to = "reading")
```

Plotting for all data
```{r}
for (p_type in unique(main_united$type)) {
  # Calculate percentage reduction
  val <- get_redu(main_united, "indoor", p_type)
  # Plot data
  print(plot_raw(main_united, p_type, val))
}
```

## Plotting for Indoor/Outdoor Ratio
### Plotting function
The following function plots the indoor/outdoor ratio for a given dataset
```{r}
plot_ratio <- function(df_ratio, p_type, val) {
  # Filter dataset for particle type, dropping NaNs and Infs
  df_ratio %>% 
    filter(type == p_type, is.finite(reading)) %>%
    # Order off then on by creating factors
    mutate(across(case, \(x) factor(x, levels=c("off", "on")))) %>%
    
  # Plot filtered ratio data
  ggplot() +
    # Create violinplot
    geom_violin(aes(x = case, y = reading, fill = case)) +
    # Convert y-axis scale to log
    scale_y_continuous(trans = "log10") +
    # Set title, caption with percentage reduction
    labs(title = paste("Overall", labels_list[[p_type]][["title_ratio"]]),
         caption = paste0("Mean reduction = ", round(val, 1), "%")) -> my_plot
  
  # Save Plot
  ggsave(filename = paste0("artifacts/HAFTRAP/SC/violinplot/", "SC_", p_type,
                           "_ratio.png"), plot = my_plot)
  
  print(paste("Plotted for", labels_list[[p_type]][["title_ratio"]]))
  # Display plot
  my_plot
}
```

### Code Run
```{r}
# Filter, order data appropriately
main_ratio <- main_df %>%
  filter(environment == "ratio") %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(mod_pm25, mod_pm10, mod_pm1, mod_o3, mod_no2, mod_no, mod_co,
                 mod_pm1num, sems_sum50, sems_sum100, sems_sum1000, acsm_chl,
                 acsm_nh4, acsm_no3, acsm_org, acsm_so4, acsm_f43, acsm_f44,
                 acsm_f57, acsm_ratio_55_57, t300_co, t200u_no, t200u_no2,
                 t200u_nox, mcpc_counts), 
               names_to = "type", values_to = "reading")
```

Plotting for Mod-PM
```{r}
for (p_type in unique(main_ratio$type)) {
  # Calculate percentage reduction
  val <- get_redu(main_ratio, "ratio", p_type)
  # Plot data
  print(plot_ratio(main_ratio, p_type, val))
}
```