# Sanity Check Mod-PM Data

## STOP

**HAVE YOU RUN `IB_modpm_cleaning`?**
*This file loads a dataframe created from running `IB_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `IB_modpm_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `IB_modpm_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/AirPartners/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```
Participant IDs
```{r}
# Get relevant participant IDs from IB_participants
participants <- c("105", "203", "206", "207", "210", "303", "312", "404", "405", "409", "413", "502", "510", "513", "515", "606", "609", "611", "612", "613")
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/Roxbury/IB/"
```


## Load data
```{r}
# Load Mod-PM data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Plotting
### Plotting functions
Simplifies plotting of data
```{r}
plot_one <- function(plot_df, specify, t, y_check) {
  ggplot(plot_df) +
    geom_line(aes(x = date, y = .data[[t]])) +
    geom_hline(aes(yintercept = y_check)) +
    ggtitle(paste(c(specify, t), collapse = " "))
}
```

```{r}
plot_types <- function(input_df, sp, y_c) {
  # Plot time-series of subset with an appropriate horizontal line
  # PM 1
  print(plot_one(input_df, sp, "pm1", y_c))
  # PM 2.5
  print(plot_one(input_df, sp, "pm25", y_c))
  # PM 10
  print(plot_one(input_df, sp, "pm10", y_c))
}
```


```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
# Re-try if incorrect
while (!(person %in% participants)) {
  person <- readline("Error! Invalid ID, Re-enter: ")
}

# Plot stuff
for (c in c("before","after")) {
  # Filter data to appropriate subset
  df <- main_df %>%
    filter(participant_id == person, case == c, 
           environment == "indoor")
  
  if (length(unique(df$room)) > 1) {
    for (r in unique(df$room)) {
      plot_df <- df %>% filter(room == r)
      
      specify <- c(person, c, "indoor", r)
      plot_types(plot_df, specify, 10)
    }
  } else {
      specify <- c(person, c, "indoor")
      plot_types(df, specify, 10)
  }
  # Print
  print(paste("Plotted for", person, c, "indoor"))
}
```