# Modulair-PM Data Cleaning
This file cleans, and calculates the I/O ratio for Mod-PM data. Output is saved in a `modpm.RData` file that is *untracked* by Github. *All files that depend on using Mod-PM data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(readxl)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Read summary table of times
time_map <- read_excel("data/Roxbury/IB/IB_notes_summarized.xlsx") %>%
  # Convert to appropriate variable types
  mutate(across(c(dt_start, dt_hepa, dt_end), 
                  ~as.POSIXct(.x, format = "%m/%d/%Y %H:%M"))) %>%
  mutate(across(c(dt_start, dt_hepa, dt_end), 
                  ~force_tz(.x, tz = "America/New_York")))
```


```{r}
# Set path to data
path_to_data <- "data/Roxbury/IB/modpm/"

# Set data category
data_cat <- "IB_M"
```

## Helper Functions
The following two functions are used to clean and process the data

### Filter by relevant times
Each participant contains data for periods where both the sham and actual purifier was running. This function filters the relevant time periods and adds an added variable to track sham and HEPA (actual) readings.
```{r}
include_case <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # Before purifier is switched on
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start, cd$dt_hepa) ~ 'before',
      # After purifier is switched on
      between(date, cd$dt_hepa, cd$dt_end) ~ 'after'))
}
```

### Calculating 10 minute averages
When calculating the indoor/outdoor ratio, we want to calculate them over 10 minutes as opposed to every data point to account for the time it takes for the air to mix between indoor and outdoor. This function calculates that ten minute average.
```{r}
# Round over 10 minute averages
round_10min <- function(df) {
  df %>%
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, room, date_round) %>%
  # Calculate mean over ten minutes
  summarize(across(c(pm1, pm25, pm10, pm1num), mean), .groups = "drop")
}

```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (file_name in list.files(path_to_data)) {
  # Read csv
  file_path <- paste0(path_to_data, file_name)
  df <- read_csv(file_path, show_col_types = FALSE)

  # Split file name into different parts
  name_split <- strsplit(file_name, "[_.]")[[1]] %>% head(-1)
  person <- name_split[3]
  env <- name_split[4]
  place <- name_split[5]
  
  # Add the participant ID, environment, and room (if present, else NA)
  df <- mutate(df, participant_id = person, environment = env, room = place)

  # Append to main dataframe
  all_df <- rbind(all_df, df)
  
  print(paste("Loaded participant", person, env, place))
}
```

### Process all data
This involves dealing with datetimes, keeping only relevant columns, calculating a new column using bin data, and removing duplicate values.
```{r}
# Clean data (this takes a while to run)
clean_df <- all_df %>%
  
  # Reformat timestamps to sensible format
  mutate(date = as.POSIXct(timestamp, tz = "America/New_York")) %>%
  # Round times to nearest minute
  mutate(date = round_date(date, unit = "minute")) %>%
  
  # Select relevant columns
  select(date, environment, room, participant_id, pm1:pm10, bin0:bin2) %>%
  # Calculate number count of particles in PM1 size
  mutate(pm1num = bin0 + bin1 + bin2, .keep = "unused") %>%
  
  # Remove repeat readings
  distinct(participant_id, room, date, environment, .keep_all = TRUE)
```

### Add cases
Using the `IB_notes_summarized` spreadsheet, add in the category of "before" or "after" HEPA purifier installation for each deployment
```{r}
# Filter out only indoor values
indoor_df <- clean_df %>% filter(environment == "indoor")

# List of participants (by default all, feel free to set to custom)
participants <- unique(indoor_df$participant_id)

# Filter by time to find whether before or after HEPA purifier installation
# Only for INDOOR deployments
case_df <- data.frame()
for (person in participants) {
  # Filter indoor (dates match outdoor) and by current participant
  current_dates <- time_map %>%
    filter(participant_id == person) %>% 
    head(1)
  
  # Filter by date for sham and hepa (see helper function)
  df <- include_case(indoor_df, person, current_dates)
  # Append to overall dataframe
  case_df <- rbind(case_df, df)
}


# Filter out only outdoor values and add dummy case
outdoor_df <- clean_df %>% 
  filter(environment == "outdoor") %>%
  filter(complete.cases(across(pm1:pm1num))) %>%
  mutate(case = NA)
```

Run this code chunk only ONCE - cleans out all NA values
```{r}
# filter out NA and infinite values in data
case_df[sapply(case_df, is.infinite)] <- NA
case_df <- case_df %>%
  filter(complete.cases(across(-room))) %>%
  # Remove repeat readings
  distinct(case, participant_id, date, environment, room, .keep_all = TRUE)
```


### Calculate indoor-outdoor ratio
For each participant, calculate 10 minute averages, join with outdoor data and calculate the indoor/outdoor ratio
```{r}
# Round over 10 minute averages
indoor_df_round <- round_10min(case_df)
outdoor_df_round <- round_10min(outdoor_df)
```

Do a left join and calculate ratios, removing unnecessary columns
```{r}
# Calculate ratios
ratio_df <- indoor_df_round %>% 
  # Join indoor and outdoor data (outdoor duplicates over date)
  left_join(outdoor_df_round, join_by(date_round == date_round), 
            suffix = c("_indoor", "_outdoor")) %>%
  # Calculate indoor-outdoor ratio
  mutate(pm1_ratio = pm1_indoor/pm1_outdoor, 
         pm25_ratio = pm25_indoor/pm25_outdoor, 
         pm10_ratio = pm10_indoor/pm10_outdoor, 
         pm1num_ratio = pm1num_indoor/pm1num_outdoor) %>%
  # Select relevant columns
  select(date_round, participant_id_indoor, case_indoor, room_indoor, 
         ends_with("ratio")) %>%
  # Remove suffixes
  rename_with(~ gsub("_indoor|_ratio|_round", "", .x)) %>%
  # Insert environment as ratio
  mutate(environment = "ratio")
```

### Bind into one dataframe
```{r}
main_df <- rbind(case_df, outdoor_df, ratio_df) 
```

### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0("cleaning/Roxbury/IB/", "cleaned_modpm.RData"))
```
