# CPC Data Cleaning
This file cleans, and calculates the I/O ratio for CPC data. Data is from the output of `cpc_preprocessing`, but that script doesn't need to be run since the data should already be present in CSV files in the data folder. To update CPC data, run `cpc_preprocessing` *then* run this script.

Output is saved in a `cleaned_cpc.RData` file that is *untracked* by Github. *All files that depend on using CPC data must be run after running this file.*

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Read summary table of times
time_map <- read_excel("data/HAFTRAP/TU/TU_notes_summarized.xlsx")

# Get relevant participant IDs from TU_participants
participants <- unique(time_map$participant_id)
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/TU/cpc/"

# Set data category
data_cat <- "TU_C"
```

## Helper Functions:
The following functions are not part of the main pipeline to calculate summary statistics, but is used to clean and process the data. 

### Calculating 10 minute averages
When calculating the indoor/outdoor ratio, we want to calculate them over 10 minutes as oppposed to every data point to account for the time it takes for the air to mix between indoor and outdoor. This function calculates that ten minute average.
```{r}
# Round over 10 minute averages
round_10min <- function(df) {
  df %>%
    # Create new column with rounded dates
    mutate(date_round = round_date(date, "10 minutes")) %>%
    # Group by the rounded datetime
    group_by(case, environment, participant_id, date_round) %>%
    # Calculate mean over ten minutes
    summarize(across(c(cpc_count), mean), .groups = "drop")
}
```

### Filter by relevant times
Each participant contains data for periods where both the sham and actual purifier was running. This function filters the relevant time periods and adds an added variable to track sham and HEPA (actual) readings.
```{r}
include_case <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # For sham periods
      # Note: 'cd' is the table 'current_dates' in function call
      period == cd$period[cd$case == "sham (dummy)"] ~ 'sham (dummy)',
      # For HEPA periods
      period == cd$period[cd$case == "hepa (dummy)"] ~ 'hepa (dummy)'))
}
```


## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each environment
  for (env in c("indoor", "outdoor")) {
    # Loop through each period
    for (per in 1:2) {
      # Set file path
      # Filename = data category + participant ID + environment
      file_name <- paste(data_cat, person, env, per, sep = "_")
      
      # File path = path to folder, type csv
      file_path <- paste0(path_to_data, file_name, ".csv")
      
      # Read csv
      df <- read_csv(file_path, show_col_types = FALSE) %>%
        # Add the environment type (e.g. indoor) and participant ID
        mutate(environment = env, participant_id = person, period = per)
      
      # Append to main dataframe
      all_df <- rbind(all_df, df) 
      }
    }
  print(paste("Loaded participant", person))
}
```

```{r}
# Warning: this code chunk may seem simple but takes FOREVER.
# Have patience. It deals with datetimes.
clean_df <- all_df %>%
  # Merge the date and time column
  unite(col = "date", Date, Time, sep = " ") %>%
  # Convert to datetime object
  mutate(date = as.POSIXct(date, format = "%F %T", tz = "America/New_York")) %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, period, .keep_all = TRUE) %>%
  # Rename column to be more descriptive
  rename(cpc_count = Count)
```



```{r}
# Filter by time to find whether sham or hepa
case_df <- data.frame()
for (person in participants) {
  # Filter indoor (dates match outdoor) and by current participant
  current_dates <- time_map %>% 
    filter(environment == "indoor") %>%
    filter(participant_id == person)
  # Filter by date for sham and hepa (see helper function)
  df <- include_case(clean_df, person, current_dates)
  # Append to overall dataframe
  case_df <- rbind(case_df, df)
}
```

```{r}
# Crucial filtering based on sanity checks
main_df <- case_df %>%
  # Period is now redundant
  select(-period) %>%
  # Filter out cpc_count values that are < 100 (this can never be true IRL)
  filter(cpc_count > 100) %>%
  # Remove all filtered-out values
  drop_na()
```


### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that are then divided by each other to calculate the ratio. After calculating the ratio, the dataframe is 'gathered' back up to its original shape.
```{r}
# Round over 10 minute averages
df_round <- round_10min(main_df)

# Spread: Create separate indoor and outdoor columns for measurements
df_round_wide <- pivot_wider(df_round, 
                             names_from = environment, 
                             values_from = cpc_count)

# Divide indoor by outdoor to calculate ratios
df_ratio <- select(df_round_wide, indoor)/ select(df_round_wide, outdoor)


# Rename computed columns to ratio
df_ratio <- rename(df_ratio, cpc_count = indoor)

# Merge back with joined data-frame
df_ratio <- cbind(df_round_wide, df_ratio) %>%
                    select(!c(indoor, outdoor)) %>%
                    mutate(environment = "ratio") %>%
                    rename(date = date_round)
```

Run this code chunk only ONCE
```{r}
# Bind back into main dataframe
main_df <- rbind(main_df, df_ratio)

# filter out NA and infinite values in data
main_df[sapply(main_df, is.infinite)] <- NA
main_df <- drop_na(main_df)
```

### Save Data
The `main_df` dataframe is saved to `cleaned_cpc.RData`
```{r}
save(main_df, file = paste0("cleaning/HAFTRAP/TU/", "cleaned_cpc.RData"))
```
