# Script for pre-processing CPC data from text files
CPC data for these deployments came in multiple text files for each participant. This parses them into one indoor and one outdoor csv for each participant with only the relevant variables. 

Note: Folders were renamed, and all non .dat files were manually deleted from the dataset, as well as all .dat files that were empty (no data, only headers).
```{r}
library(data.table)
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path; Commented out because it wasn't working on Scott's computer
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/TU/cpc/"

# Set data category
data_cat <- "TU_C"
```

```{r}
# List all directories in raw data
folders <- list.dirs(path = paste0(path_to_data, "very_raw_data/"), full.names = TRUE, recursive = TRUE)
```

```{r}
# Loop through each subfolder
for (subfolder in folders) {
  
  # Split path into individual parts
  path_parts <- strsplit(subfolder, split = "/")
  
  # Only evaluate directories that have depth of 15 from the root (i.e. with
  # data files). Change this number according to the depth at which data files appear in the file structure. (e.g. on Vedaant's computer it's 7.)
  if (length(path_parts[[1]]) == 7) {
    
    # Isolate environment and participant ID
    id <- path_parts[[1]][6]
    env <- path_parts[[1]][7]
    # 
    # # Skip unwanted files
    # if (id == "Pre data" || grepl("bad", id)) {
    #   print(paste("Skipped", env, "data for participant", id))
    #   next
    # }
    
    print(paste("Reading", env, "data for participant", id, "..."))
    
    # Get list of files in directory
    main_df <- list.files(path = subfolder, full.names = TRUE) %>% 
      # Read into dataframe (datetime variables, concent, inlttmp, smpflow)
        map_df(~fread(.x, header = TRUE, skip = 5, select = c(1, 2, 4), 
                      fill = FALSE))
      
    # Make new file name
    new_file_name <- paste(data_cat, id, env, sep = "_")
    print(paste("Writing", new_file_name, "..."))
    
    # Write into appropriate path
    fwrite(main_df, paste0(path_to_data, new_file_name, ".csv"))
    print(paste("Downloaded", env, "data for participant ID", id))
  }
}
```