<!-- # Sanity Check Mod-PM Data -->

## STOP

**HAVE YOU RUN `CH_modpm_cleaning`?**
*This file loads a dataframe created from running `CH_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `CH_modpm_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `CH_modpm_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/Revere/CH/"
```


## Load data
```{r}
# Load Mod-PM data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

Participant IDs
```{r}
# Get relevant participant IDs from CH_participants
participants <- unique(main_df$participant_id)
```

## Plotting
### Plotting functions
Simplifies plotting of data
```{r}
plot_one <- function(plot_df, specify, t, y_check) {
  ggplot(plot_df) +
    geom_line(aes(x = date, y = .data[[t]])) +
    geom_hline(aes(yintercept = y_check)) +
    ggtitle(paste(c(specify, t), collapse = " "))
}
```

```{r}
plot_types <- function(input_df, sp, y_c) {
  # Plot time-series of subset with an appropriate horizontal line
  # PM 1
  print(plot_one(input_df, sp, "pm1", y_c))
  # PM 2.5
  print(plot_one(input_df, sp, "pm25", y_c))
  # PM 10
  print(plot_one(input_df, sp, "pm10", y_c))
}
```


```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
# Re-try if incorrect
while (!(person %in% participants)) {
  person <- readline("Error! Invalid ID, Re-enter: ")
}

if (person == "all"){
  plot_df <- main_df %>%
    filter(participant_id == person, environment == "outdoor")
  
  specify <- c(person, "outdoor")
  plot_types(plot_df, specify, 10)
  
} else {
  # Plot stuff
  for (c in c("off","on")) {
    for (env in c("indoor", "ratio")) {
      # Filter data to appropriate subset
    df <- main_df %>%
      filter(participant_id == person, case == c, environment == env)
      specify <- c(person, c, env)
      plot_types(df, specify, 10)
    # Print
    print(paste("Plotted for", person, c, env))
    }
  }
}
```