# Aethalometer Data Cleaning
This file cleans and assigns participant ID for Aethalometer data. Output is saved in a `aethalometer.RData` file that is *untracked* by Github. *All files that depend on using Aethalometer data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(readxl)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```


```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Get relevant participant IDs from OH_participants
participants <- as.character(
  na.omit(read_excel("data/HAFTRAP/OH/OH_participants.xlsx")$aethalometer))

# Read summary table of times
time_map <- read_excel("data/HAFTRAP/OH/OH_notes_summarized.xlsx") %>%
  # Convert to appropriate variable types
  mutate(across(dt_start:dt_end, 
                  ~as.POSIXct(.x, format = "%m/%d/%Y %H:%M"))) %>%
  mutate(across(dt_start:dt_end, 
                  ~force_tz(.x, tz = "America/New_York")))
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/aethalometer/"

# Set data category
data_cat <- "OH_E"
```

## Helper Functions
The following function is used to clean and process the data

### Calculating 10 minute averages
When calculating the indoor/outdoor ratio, we want to calculate them over 10 minutes as opposed to every data point to account for the time it takes for the air to mix between indoor and outdoor. This function calculates that ten minute average.
```{r}
# Round over 10 minute averages
round_10min <- function(df) {
  df %>%
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean over ten minutes
  summarize(across(c(uv_bc1, red_bc1), mean),
            .groups = "drop")
}
```

## Main Code Run
### Load all data. 
Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

for (env in c("indoor", "outdoor")) {
      
      # Set file path
      # Filename = data category + environment 
      file_name <- paste(data_cat, env, sep = "_")
      # File path = path to folder, type csv
      file_path <- paste0(path_to_data, file_name, ".csv")
      
      # Read csv
      df <- read_csv(file_path, show_col_types = FALSE) %>%
        # Add the environment
        mutate(environment = env, 
               date = force_tz(date, tzone = "America/New_York"))
      
      # Append to main dataframe
      all_df <- rbind(all_df, df)
}
```

### Add Participant ID and case (sham or hepa)
Using cross joins and filtering to find the participant ID, followed by case
```{r}
time_map_aethalometer <- time_map %>% 
  filter(participant_id %in% participants, environment == "indoor") %>%
  select(-environment)

main_df <- all_df %>%
  cross_join(time_map_aethalometer) %>%
  mutate(participant_id = if_else(between(date, dt_start, dt_end), 
                                  participant_id, NA)) %>% 
  drop_na() %>% select(-c(dt_start, dt_end, sensor_num)) %>%
  mutate(participant_id = as.character(participant_id))
```

### Filter invalid values
Note that currently only filters PM values
```{r}
# Filter out values over a certain threshold
main_df_indoor <- subset(main_df, environment == "indoor")
main_df_outdoor <- subset(main_df, environment == "outdoor")

main_df <- rbind(main_df_indoor, main_df_outdoor)
```

### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that are then divided by each other to calculate the ratio. After calculating the ratio, the dataframe is 'gathered' back up to its original shape.

Note: Always run this chunk with the previous chunk.
```{r}
# Round over 10 minute averages
df_indoor_round <- round_10min(main_df_indoor)
df_outdoor_round <- round_10min(main_df_outdoor)

# Merge
df_round <- merge(df_indoor_round, df_outdoor_round, 
                  by = c("date_round", "participant_id", "case"), 
                  suffixes = c("_indoor", "_outdoor")) %>%
            # Remove redundant environment columns
            select(!starts_with("environment"))
          

df_ratio <- df_round %>% 
  mutate(date = date_round,
         environment = "ratio", 
         uv_bc1 = uv_bc1_indoor/uv_bc1_outdoor,
         red_bc1 = red_bc1_indoor/red_bc1_outdoor,
         .keep = "unused")
```

Run this code chunk only ONCE
```{r}
# Bind back into main dataframe
main_df <- rbind(main_df, df_ratio)

# filter out NA and infinite values in data
main_df[sapply(main_df, is.infinite)] <- NA
main_df <- drop_na(main_df) %>%
  # Remove repeat readings
  distinct(case, participant_id, date, environment, .keep_all = TRUE)
```


### Save Data
The `main_df` dataframe is saved to `cleaned_aethalometer.RData`
```{r}
save(main_df, file = paste0("cleaning/HAFTRAP/OH/", "cleaned_aethalometer.RData"))
```