# Script for pre-processing Aethelometer data from raw CSVs
Aethelometer data arrives in confusing text files with too many unnecessary columns. This script simply selects the relevant columns and joins all the data into CSV files (one for indoor, another for outdoor). Note that the outdoor sensor malfunctioned so the data for that isn't probably going to be used further.

```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path; Commented out because it wasn't working on Scott's computer
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/aethalometer/"

# Set data category
data_cat <- "OH_E"
```

```{r}
# List all directories in raw data
folders <- list.dirs(path = paste0(path_to_data, "very_raw_data/"), full.names = TRUE, recursive = TRUE)
```

```{r}
# Loop through each subfolder
for (subfolder in folders) {
  # Split path into individual parts
  path_parts <- strsplit(subfolder, split = "/")
  
  # Only evaluate directories that have specific depth from root 
  # (change according to your OS/file structure
  file_depth <- 6
  if (length(path_parts[[1]]) == file_depth) {
    
    # Isolate environment and participant ID
    env <- path_parts[[1]][file_depth]
    
    print(paste("Reading", env, "data for participant"))
    
    # Get list of files in directory
    main_df <- list.files(path = subfolder, full.names = TRUE) %>% 
      # Read into dataframe (datetime variables, concent, inlttmp, smpflow)
      map_df(~read_csv(.x, show_col_types = FALSE) %>%
      
      # Select relevant variables and rename
      select(`Date / time local`, `UV BC1`, `Red BC1`) %>%
      rename(date = `Date / time local`, uv_bc1 = `UV BC1`, red_bc1 = `Red BC1`))
      # Note: pipe is inside map_df - conversion happens for every file 
      # read into dataframe
      
    # Make new file name
    new_file_name <- paste(data_cat, env, sep = "_")
    print(paste("Writing", new_file_name, "..."))
    
    # Write into appropriate CSV
    write_csv(main_df, paste0(path_to_data, new_file_name, ".csv"))
    print(paste("Combined", env, "data"))
  }
}
```