# Sanity Check CPC Data

## STOP

**HAVE YOU RUN `OH_cpc_cleaning`?**
*This file loads a dataframe created from running `OH_cpc_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `OH_cpc_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `OH_cpc_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(ggplot2)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/vkuchhal/Documents/hepa-summer23')
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs
```{r}
# Vector of participant numbers' data to be processed
participants <- c("41271", "42281", "45361", "44621", "45451")
```

Paths
```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/cpc/"
```


## Load data
```{r}
# Load CPC data from RData file
load(paste0(path_to_data, "cleaned_cpc.RData"))
```

## Plotting
```{r}
# Plot stuff
for (person in participants) {
  for (env in c("indoor", "outdoor")) {
    y_check <- case_match(
      env,
      "indoor" ~ 50000,
      "outdoor" ~ 100000,
    )
    for (c in c("sham","hepa")) {
      # Filter data to appropriate subset
      df <- main_df %>%
        filter(participant_id == person, case == c, 
               environment == env)
      
        # Plot time-series of subset with an appropriate horizontal line
        print(ggplot(df) +
          geom_line(aes(x = date, y = concent)) +
          geom_hline(aes(yintercept = y_check)) +
          ggtitle(paste(person, c, env)))
      
      # Pause and print
      Sys.sleep(2)
      print(paste("Plotted for", person, c, env))
    }
  }
}
```

