# Modulair-PM Data Cleaning
This file cleans, and calculates the I/O ratio for Mod-PM data. Output is saved in a `cleaned_modpm.RData` file that is *untracked* by Github. *All files that depend on using Mod-PM data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(readxl)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```


```{r}
# Set path to data
path_to_data <- "data/Revere/RH/modpm/"

# Set data category
data_cat <- "RH_M"
```

Data Thresholds
```{r}
# Max thresholds for filtering erroneous values
pm10_threshold <- 1000
pm25_threshold <- 100
pm1_threshold <- 50
```

## Helper Functions
The following two functions are used to clean and process the data

### Calculating 10 minute averages
When calculating the indoor/outdoor ratio, we want to calculate them over 10 minutes as opposed to every data point to account for the time it takes for the air to mix between indoor and outdoor. This function calculates that ten minute average.
```{r}
# Round over 10 minute averages
round_10min <- function(df) {
  df %>%
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, participant_id, date_round) %>%
  # Calculate mean over ten minutes
  summarize(across(c(pm1, pm25, pm10, pm1num), mean), .groups = "drop")
}

```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (file_name in list.files(path_to_data)) {
  # Read csv
  file_path <- paste0(path_to_data, file_name)
  df <- read_csv(file_path, show_col_types = FALSE)

  # Split file name into different parts
  name_split <- strsplit(file_name, "[_.]")[[1]] %>% head(-1)
  room <- name_split[3]
  subroom <- name_split[4]
  
  # Add the participant ID
  df <- df %>%
    select(date1min, HEPA_on, starts_with(c("bin0.", "bin1.", "bin2.", 
                                            "pm1.", "pm25.", "pm10."))) %>%
    mutate(participant_id = paste(room, subroom, sep = "_"))

  # Append to main dataframe
  all_df <- rbind(all_df, df)
  
  print(paste("Loaded room", room, subroom))
}
```

```{r}
main_df <- all_df %>% 
  
  # Calculate number count of particles in PM1 size
  mutate(pm1num.indoor = bin0.indoor + bin1.indoor + bin2.indoor,
         pm1num.outdoor = bin0.outdoor + bin1.outdoor + bin2.outdoor,
         # Convert HEPA readings to on or off
         case = case_when(HEPA_on == 0 ~ "off", HEPA_on == 1 ~ "on"), 
         .keep = "unused") %>%
  
  # Calculate indoor-outdoor ratio
  mutate(pm1.ratio = pm1.indoor / pm1.outdoor,
         pm25.ratio = pm25.indoor / pm25.outdoor,
         pm10.ratio = pm10.indoor / pm10.outdoor,
         pm1num.ratio = pm1num.indoor / pm1num.outdoor) %>%
  
  # Gather dataframe to make environment a category
  pivot_longer(cols = contains("."), 
                names_to = c(".value", "environment"), 
                names_sep = "\\.") %>%
  
  # Remove NAs, drop outdoor data, and rename date column to standardize
  drop_na() %>%
  filter(environment != "outdoor") %>%
  rename(date = date1min)
```

### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0("cleaning/Revere/RH/", "cleaned_modpm.RData"))
```
