# Modulair-PM Data Cleaning
This file cleans, and calculates the I/O ratio for Mod-PM data. Output is saved in a `cleaned_modpm.RData` file that is *untracked* by Github. *All files that depend on using Mod-PM data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```


```{r}
# Set path to data
path_to_data <- "data/Revere/RH/modpm/"

# Set data category
data_cat <- "RH_M"
```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (file_name in list.files(path_to_data)) {
  # Read csv
  file_path <- paste0(path_to_data, file_name)
  df <- read_csv(file_path, show_col_types = FALSE)

  # Split file name into different parts
  name_split <- strsplit(file_name, "[_.]")[[1]] %>% head(-1)
  room <- name_split[3]
  subroom <- name_split[4]
  
  
  df <- df %>%
    # Filter relevant columns
    select(date1min, HEPA_on, starts_with(c("bin0.", "bin1.", "bin2.",
                                            "pm1.", "pm25.", "pm10.")),
           pm1_ratio, pm25_ratio, pm10_ratio) %>%
    
    # Add the participant ID
    mutate(participant_id = paste(room, subroom, sep = "_"))

  # Append to main dataframe
  all_df <- rbind(all_df, df)
  
  print(paste("Loaded room", room, subroom))
}
```

```{r}
curated_df <- all_df %>% 
  
  # Calculate number count of particles in PM1 size
  mutate(pm1num.indoor = bin0.indoor + bin1.indoor + bin2.indoor,
         pm1num.outdoor = bin0.outdoor + bin1.outdoor + bin2.outdoor,
         
         # Convert HEPA readings to on or off
         case = case_when(HEPA_on == 0 ~ "off", HEPA_on == 1 ~ "on"), 
         
         # Drop redundant columns
         .keep = "unused") %>%
  
  mutate(pm1num_ratio = pm1num.indoor / pm1num.outdoor) %>%
  
  # Rename date column to standard
  rename(date = date1min) %>%
  
  # Change ratio column names delimiter from _ to .
  rename_with( ~ gsub("_", ".", .x, fixed = TRUE), ends_with("ratio"))
```

```{r}
# Bind together indoor, ratio data
main_df <- curated_df %>%
  
  # Gather dataframe to make environment a category
  pivot_longer(cols = contains("."), 
               names_to = c(".value", "environment"), 
               names_sep = "\\.") %>%
  
  # Drop outdoor data and remove NAs
  filter(environment != "outdoor") %>%
  drop_na()
```


### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0("cleaning/Revere/RH/", "cleaned_modpm.RData"))
```
