# SCOPE HAFTRAP Sanity Checking

## STOP

**HAVE YOU RUN `SC_all_cleaning`?**
*This file loads a dataframe created from running `SC_all_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to sanity check cleaning results for multiple participants in the HAFTRAP study, for all deployments in the SCOPE project.

Results can be found in `summary/HAFTRAP/OH/s_OH_M_quants.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/HAFTRAP/SC/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_everything.RData"))
```

```{r}
participants = c("1")
```

```{r}
column_names <- 
  colnames(main_df %>% select(-date, -environment, -case, -participant_id))
```


## Plot Data
```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
# Re-try if incorrect
while (!(person %in% participants)) {
  person <- readline("Error! Invalid ID, Re-enter: ")
}

# Plot stuff
for (env in c("indoor", "outdoor", "ratio")) {
  for (c in c("on","off")) {
    # Filter data to appropriate subset
    df <- main_df %>%
      filter(participant_id == person, case == c, 
             environment == env)
    for (t in column_names) {
      print(ggplot(df) +
        geom_line(aes(x = date, y = .data[[t]])) +
        ggtitle(paste(person, c, env, t)))
    }
    # Print
    print(paste("Plotted for", person, c, env))
  }
}
```
