```{r}
library(tidyverse)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```


```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/SC/"

# Set category (Modulair + ACSM + SEMS)
data_cat <- "SC_RASCT"
```

```{r}
# All participants
participants <- c("1", "2", "3", "4")

# Read relevant summary notes for time splitting
time_map <- read_excel(paste0(path_to_data, "SC_notes_summarized.xlsx"))
```

### Filter by relevant times
Each participant contains data for periods where both the purifier was off or on. This function filters the relevant time periods and adds an added variable to track off or on readings.
```{r}
# Function to include whether hepa purifier is on or off
include_case <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # For off periods
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start[cd$case == "off"], 
              cd$dt_end[cd$case == "off"]) ~ 'off',
      # For on periods
      between(date, cd$dt_start[cd$case == "on"], 
              cd$dt_end[cd$case == "on"]) ~ 'on'))
}
```

### Make sections for each pair
For data where the sensors have valve switching indoor-outdoor, this function makes a "section" for each pair. Since ratio is calculated from *previous* outdoor divided by indoor, removes the first set of indoor values. 
```{r}
# Make a "section" for I/O pair to calculate ratios
make_section <- function(df) {
  df %>%
    # Drop rows with NaNs (removes empty time values)
    drop_na() %>% 
    # Order by date - indoor-outdoor alternating sections is now clear to see
    arrange(date) %>%
    # Increase value of "section" by 1 every time environment changes
    mutate(section = cumsum(environment != lag(environment, default = ""))) %>%
    # Remove first set of "indoor" readings (why explained in get_ratios)
    filter(section > 1) %>%
    # Floor divide so that one outdoor-indoor pair is one section
    mutate(section = section %/% 2)
}
```

### Get first dates
```{r}
# Isolate single date for each section - take first in group
first_dates <- function(df) {
  df %>%
    group_by(section) %>%
    slice(1) %>%
    ungroup() %>%
    select(section, date)
}
```

### Get ratios
All the code used to calculate ratios and add it to dataframe.
```{r}
get_ratios <- function(df, val_cols) {
  # Group by, summarize, and widen
  df_wide <- df %>%
    # Group by (mainly) section and environment
    group_by(section, environment, case, participant_id) %>%
    # Calculate the mean for each indoor/outdoor for each section
    summarise(across(all_of(val_cols), mean), .groups = "drop") %>%
    # Split indoor and outdoor values into separate columns
    pivot_wider(names_from = environment, 
                values_from = val_cols)
    
  # If there is only one value column, prepend column name to correct
  if (length(val_cols) == 1) {
    df_wide <- df_wide %>% rename_with(~paste(val_cols, ., sep = "_"),
                                       indoor:outdoor)
  }
  
  # Divide indoor by outdoor to calculate ratios
  df_ratio <- select(df_wide, ends_with("indoor"))/
    select(df_wide, ends_with("outdoor"))
  
  # Rename computed columns
  colnames(df_ratio) <- sub("_indoor", "", colnames(df_ratio))
  
  # Merge back with joined data-frame
  df_ratio <- cbind(df_wide, df_ratio) %>%
    # Only keep case, section, participant
    select(!(ends_with("indoor") | ends_with("outdoor"))) %>%
    # Add environment as ratio
    mutate(environment = "ratio") %>%
    # Merge in the appropriate date
    left_join(first_dates(df), by = "section")
  
  # Bind back into main dataframe
  rbind(df, df_ratio)
}
```

### Make dataframe from specified range
This function takes a dataframe where the selected sensors have a valve switching indoor-outdoor - with readings at the same time. 
```{r}
# Makes a cleaned dataframe with ratios
make_df <- function(df, col_vec) {
  # Merge selected columns with standard set of columns
  joined_vec <- c("date", "case", "environment", "participant_id", col_vec)
  
  df %>%
    # Select all specified columns and make section
    select(all_of(joined_vec)) %>%
    make_section() %>%
    
    # Calculated indoor/outdoor ratios for each section for specified columns
    get_ratios(col_vec) %>%
    
    # Make section
    select(-section)
}
```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each environment
  for (env in c("indoor", "outdoor")) {
    # Set file path
    # Filename = data category + participant ID + environment
    file_name <- paste(data_cat, person, env, sep = "_")
    
    # File path = path to folder, type csv
    file_path <- paste0(path_to_data, file_name, ".csv")
    
    # Read csv
    df <- read_csv(file_path, show_col_types = FALSE) %>%
      # Add the environment type (e.g. indoor) and participant ID
      mutate(environment = env, participant_id = person) %>%
      # Rename columns to lowercase
      rename_all(tolower)
    
    # Remove suffixes ("_in" for indoor, "_out" for outdoor) for binding
    col_suffix <- if (env == "indoor") "_in" else "_out"
    
    if (is.null(df$stove_temp)) {
      df$stove_temp = NA
    }
    
    print(col_suffix)
    colnames(df) <- gsub(paste0(col_suffix, "$"), "", colnames(df))
    
    # Append to main dataframe
    all_df <- rbind(all_df, df)
  }
  print(paste("Loaded participant", person))
}
```
Deal with datetimes
```{r}
# Convert character to datetime
clean_df <- all_df %>%
  # Convert to datetime object
  mutate(date = as.POSIXct(acsm_local_time, format = "%m/%d/%Y %H:%M", 
                           tz = "America/New_York")) %>%
  # Drop acsm_local_time column (replaced by date)
  select(-acsm_local_time) %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, .keep_all = TRUE)
```

Add case (HEPA filter "on" or "off")
```{r}
# Filter by time to find whether hepa is off or on
case_df <- data.frame()
for (person in participants) {
  # Filter by current participant
  current_dates <- time_map %>% 
    filter(participant_id == person)
  # Filter by date for off and on (see helper function)
  df <- include_case(clean_df, person, current_dates)
  # Append to overall dataframe
  case_df <- rbind(case_df, df)
}
```

Compute additional columns
```{r}
# Compute important columns
main_df <- case_df %>%
  # Sum particle counts
  mutate(mod_pm1num = mod_bin0 + mod_bin1 + mod_bin2) %>%
  # Convert to rowwise to prevent NaNs being carried over
  rowwise() %>%
  # Create SEMS bin for < 50 nm (checking sems_bin_dia columns)
  mutate(sems_sum50 = sum(c_across(sems_bin_conc1:sems_bin_conc30))) %>%
  # Create SEMS bin for < 100 nm
  mutate(sems_sum100 = sum(c_across(sems_bin_conc1:sems_bin_conc38))) %>%
  # Create SEMS bin for < 1000 nm (effectively sum all columns)
  mutate(sems_sum1000 = sum(c_across(sems_bin_conc1:sems_bin_conc59)))
```

Split into separate dataframes, add ratios
```{r}
# Make clean dataframe, calculate ratios for each data "group"
# For SEMS data
sems_df <- make_df(main_df, c("sems_sum50", "sems_sum100", "sems_sum1000"))

# For ACSM data for specific pollutants
acsm_df1 <- make_df(main_df, c("acsm_chl", "acsm_nh4", "acsm_no3", "acsm_org", 
                               "acsm_so4"))

# For other ACSM data
acsm_df2 <- make_df(main_df, c("acsm_f43", "acsm_f44", "acsm_f57",
                               "acsm_ratio_55_57"))

# For data from Teledyne sensors
td_df <- make_df(main_df, c("t300_co", "t200u_no", "t200u_no2", "t200u_nox"))

# For CPC count
cpc_df <- make_df(main_df, c("mcpc_counts"))
```

Process Modulair data
```{r}
# Deal with modulair data separately since it has simultaneous indoor + outdoor
mod_df <- main_df %>% 
  # Select Modulair data columns
  select(date, environment, participant_id, case, mod_pm25, mod_pm10, mod_pm1,
         mod_o3, mod_no2, mod_no, mod_co, mod_pm1num, stove_temp) %>%
  # Create new column with dates rounded to ten minutes
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Order by rounded dates (remove row-wise)
  arrange(date_round) %>% 
  ungroup() %>%
  select(-date_round) %>%
  # Increment by one everytime environment changes
  mutate(section = cumsum(environment != lag(environment, default = ""))) %>%
  # Make single indoor-outdoor pair as one section (no dropping first "indoor")
  mutate(section = (section + 1) %/% 2) %>%
  
  # Get indoor-outdoor ratio for each section for specified columns
  get_ratios(c("mod_pm25", "mod_pm10", "mod_pm1","mod_o3", 
               "mod_no2", "mod_no", "mod_co", "mod_pm1num", "stove_temp")) %>%

  # Drop section column
  select(-section)
```

## Save Data
Merge all dataframes into one and save into `cleaned_everything.Rdata`
```{r}
obs_cols <- c("date", "participant_id", "case", "environment")

main_df <- mod_df %>%
  full_join(sems_df, by = obs_cols) %>%
  full_join(acsm_df1, by = obs_cols) %>%
  full_join(acsm_df2, by = obs_cols) %>%
  full_join(td_df, by = obs_cols) %>%
  full_join(cpc_df, by = obs_cols) 

save(main_df, file = 
       paste0("cleaning/HAFTRAP/SC/", "cleaned_everything.RData"))
```

