# Sanity Check Mod-PM Data

## STOP

**HAVE YOU RUN `MF_hobo_cleaning`?**
*This file loads a dataframe created from running `MF_hobo_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `MF_modpm_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `MF_modpm_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```
Participant IDs
```{r}
# Get relevant participant IDs from MF_participants
participants <- c("2", "3", "5", "6", "7", "8", "10", "11")
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```


## Load data
```{r}
# Load Mod-PM data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_hobo.RData"))
```

## Plotting
### Plotting functions
Simplifies plotting of data
```{r}
plot_one <- function(plot_df, person, t, y_check) {
  ggplot(plot_df) +
    geom_line(aes(x = date, y = .data[[t]])) +
    ggtitle(paste(person, collapse = " "))
}
```

```{r}
plot_types <- function(input_df, p) {
  # Plot time-series of subset with an appropriate horizontal line
  # PM 1
  print(plot_one(input_df, p, "voltage"))
  # PM 2.5
  print(plot_one(input_df, p, "current"))
  # PM 10
  print(plot_one(input_df, p, "power"))
  # PM 1 Count
  print(plot_one(input_df, p, "energy"))
}
```


```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
# Re-try if incorrect
while (!(person %in% participants)) {
  person <- readline("Error! Invalid ID, Re-enter: ")
}

df <- main_df %>%
      filter(participant_id == person, case == "after")
    
plot_types(df, person)
    
# Print
print(paste("Plotted for", person))

```