TODO: Change so that it sets all "before" values to 0

# Modulair-PM Data Cleaning
This file organizes HOBO data, which are in CSV files (one file has multiple deployments). Output is saved in a `hobo.RData` file that is *untracked* by Github. *All files that depend on using HOBO data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(readxl)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

## Helper functions
```{r}
# Function to reformat time-stamps to time object and round to nearest minute
improve_timestamps <- function(df) {
  df %>%
    # Reformat timestamps to sensible format
    mutate(date = as.POSIXct(date, tz = "America/New_York",
                             format = "%m/%d/%y %I:%M:%S %p")) %>%
    # Round times to nearest minute
    mutate(date = round_date(date, unit = "minute"))
}
```

```{r}
include_participant <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # Before purifier is switched on
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start, cd$dt_hepa) ~ 'before',
      # After purifier is switched on
      between(date, cd$dt_hepa, cd$dt_end) ~ 'after'))
}
```

```{r}
include_case <- function(df, person, cd) {
  df %>%
    # Select current participant
    filter(participant_id == person) %>%
    # Add case column
    mutate(case = case_when(
      # Before purifier is switched on
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start, cd$dt_hepa) ~ 'before',
      # After purifier is switched on
      between(date, cd$dt_hepa, cd$dt_end) ~ 'after'))
}
```

```{r}
# Read summary table of times
time_map <- read_excel("data/EastBoston/MF/MF_notes_summarized.xlsx") %>%
  # Convert to appropriate variable types
  mutate(across(c(dt_start, dt_hepa, dt_end), 
                  ~as.POSIXct(.x, format = "%m/%d/%Y %H:%M"))) %>%
  mutate(across(c(dt_start, dt_hepa, dt_end), 
                  ~force_tz(.x, tz = "America/New_York")))
```


```{r}
# Set path to data
path_to_data <- "data/EastBoston/MF/hobo/"

# Set data category
data_cat <- "MF_H"
```

```{r, warning=FALSE}

# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (file_name in list.files(path_to_data)) {
  # Read csv
  file_path <- paste0(path_to_data, file_name)
  df <- read_csv(file_path, skip = 1, show_col_types = FALSE)
  
  # Remove unnecessary columns
  df <- df[,2:8] %>% drop_na()
  
  # Rename columns
  colnames(df) <- c("date", "voltage", "current", "power", "energy", 
                    "power_app", "power_fact")
  
  # Convert timestamps to datetime
  df <- improve_timestamps(df)
  
  # Extract serial number from file name
  name_split <- strsplit(file_name, "[_.]")[[1]] %>% head(-1)
  serial_no <- name_split[3]
  
  # Add serial number to dataframe
  df <- mutate(df, sn_hobo = serial_no)
  
  # Append to main dataframe
  all_df <- rbind(all_df, df)
  
  print(paste("Loaded HOBO logger", serial_no))
}
```
```{r}
# Initialize master dataframe for all data
clean_df <- data.frame()

for (serial_no in unique(all_df$sn_hobo)) {
  df <- all_df %>% filter(sn_hobo == serial_no)
  parts <- time_map %>% filter(sn_hobo == serial_no)
  
  serial_df <- data.frame()
  for (person in unique(parts$participant_id)) {
    cd <- parts %>% filter(participant_id == person)
    person_df <- df %>%
      filter(between(date, cd$dt_start, cd$dt_end)) %>%
      # Add case column
      mutate(participant_id = person)
    
    serial_df <- rbind(serial_df, person_df)
  }
  clean_df <- rbind(clean_df, serial_df)
}
```

```{r}
# List of participants (by default all, feel free to set to custom)
participants <- unique(clean_df$participant_id)

# Filter by time to find whether before or after HEPA purifier installation
case_df <- data.frame()
for (person in participants) {
  # Filter indoor (dates match outdoor) and by current participant
  current_dates <- time_map %>% 
    filter(environment == "outdoor") %>%
    filter(participant_id == person)
  # Filter by date for sham and hepa (see helper function)
  df <- include_case(clean_df, person, current_dates)
  # Append to overall dataframe
  case_df <- rbind(case_df, df)
}
```

```{r}
# More cleaning to come
main_df <- case_df %>%
  select(-voltage, -current, -energy, -power_app, -power_fact) %>%
  mutate(participant_id = as.character(participant_id)) %>%
  mutate(power = case_when(
    case == "before" ~ 0,
    case == "after" ~ power))
  
```

### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0("cleaning/EastBoston/MF/", "cleaned_hobo.RData"))
```