TODO: Change so that it plots before/after separately

# Sanity Check HOBO Data

## STOP

**HAVE YOU RUN `MF_hobo_cleaning`?**
*This file loads a dataframe created from running `MF_hobo_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `MF_modpm_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `MF_modpm_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```
Participant IDs
```{r}
# Get relevant participant IDs from MF_participants
participants <- c("2", "3", "5", "6", "7", "8", "11")
```

Paths
```{r}
# Set path to data
path_to_data <- "cleaning/EastBoston/MF/"
```


## Load data
```{r}
# Load Mod-PM data from RData file; loads as main_df
load(paste0(path_to_data, "cleaned_hobo.RData"))
```

## Plotting
### Plotting functions
Simplifies plotting of data
```{r}
plot_one <- function(plot_df, person, c, t) {
  plot_df %>%
    filter(case == c, participant_id == person) %>%
    ggplot() +
      geom_line(aes(x = date, y = .data[[t]])) +
      ggtitle(paste(person, t, c))
}
```

```{r}
for (p in participants) {
  print(plot_one(main_df, p, "before", "power"))
  print(plot_one(main_df, p, "after", "power"))
  # Print
  print(paste("Plotted for", p))
}
```