# HAFTRAP Plotting Boxplots

## STOP

**HAVE YOU RUN `OH_modpm_cleaning`?**
*This file loads a dataframe created from running `OH_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to plot boxplots using all the data from the Olin HAFTRAP study.

Plots can be found here and saved in the folder`summary/HAFTRAP/OH/`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(ggplot2)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/vkuchhal/Documents/hepa-summer23')
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Vector of participant numbers' data to be processed
participants <- c("41271", "42281", "45361", "44621", "45451")

### Set paths
# # For CPC
# Set path to data
path_to_data_C <- "data/HAFTRAP/OH/cpc/"

# # For Mod-PM
# Set path to data
path_to_data_M <- "data/HAFTRAP/OH/modpm/"
```

## Load data

Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data_M, "cleaned_modpm.RData"))

# Create copy of main_df, and delete original
modpm_df <- main_df %>%
  # Filtering out 41211 for now in Mod-PM because of funky data (look at CSVs)
  filter(participant_id != "41211")
rm(main_df)
```

CPC - filter data to whole minutes
```{r}
# Load CPC data from RData file
load(paste0(path_to_data_C, "cleaned_cpc.RData"))

# Create copy of main_df
cpc_df <- main_df

# Delete original
rm(main_df)
```

## Plotting for Indoor and Outdoor raw concentrations
### CPC
```{r}
# Filter, order data appropriately
cpc_united <- cpc_df %>%
  filter(environment != "ratio") %>%
  unite("situation",environment, case, sep = ", ")
```

Plotting for concent
```{r}
cpc_united <- cpc_united %>% filter(is.finite(concent))
ggplot(cpc_united) +
    geom_boxplot(aes(x = situation, y = concent), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(cpc_united$concent, c(0.05, 0.95)))
```
### Mod-PM
```{r}
# Filter, order data appropriately
modpm_united <- modpm_df %>%
  filter(environment != "ratio") %>%
  unite("situation",environment, case, sep = ", ")
```

Plotting for PM 1
```{r}
modpm_united <- modpm_united %>% filter(is.finite(pm1))
ggplot(modpm_united) +
    geom_boxplot(aes(x = situation, y = pm1), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(modpm_united$pm1, c(0.05, 0.95)))
```

Plotting for PM 2.5
```{r}
modpm_united <- modpm_united %>% filter(is.finite(pm25))
ggplot(modpm_united) +
    geom_boxplot(aes(x = situation, y = pm25), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(modpm_united$pm25, c(0.05, 0.95)))
```

Plotting for PM 10
```{r}
modpm_united <- modpm_united %>% filter(is.finite(pm10))
ggplot(modpm_united) +
    geom_boxplot(aes(x = situation, y = pm10), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(modpm_united$pm10, c(0.05, 0.95)))
```

Plotting for PM 1 Numbers
```{r}
modpm_united <- modpm_united %>% filter(is.finite(pm1num))
ggplot(modpm_united) +
    geom_boxplot(aes(x = situation, y = pm1num), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(modpm_united$pm1num, c(0.05, 0.95)))
```
