
# Script for pre-processing CPC data from text files
CPC data for these deployments came in multiple text files for each participant. This parses them into one indoor and one outdoor csv for each participant with only the relevant variables. 

Note: I manually deleted indoor files 230331A0 and 230221A1 from participant 45451 in the data.


```{r}
library(data.table)
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/vkuchhal/Documents/hepa-summer23')
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/cpc/"

# Set data category
data_cat <- "OH_C"
```

```{r}
# List all directories in raw data
folders <- list.dirs(path = paste0(path_to_data, "very_raw_data/"), full.names = TRUE, recursive = TRUE)
```

```{r}
# Loop through each subfolder
for (subfolder in folders) {
  
  # Split path into individual parts
  path_parts <- strsplit(subfolder, split = "/")
  
  # Only evaluate directories that have depth of 7 from the root (i.e. with
  # data files)
  if (length(path_parts[[1]]) == 7) {
    
    # Isolate environment and participant ID
    env <- path_parts[[1]][6]
    id <- path_parts[[1]][7]
  
    # Skip unwanted files
    if (id == "Pre data" || grepl("bad", id)) {
      print(paste("Skipped", env, "data for participant", id))
      next
    }
    
    print(paste("Reading", env, "data for participant", id, "..."))
    
    # Get list of files in directory
    main_df <- list.files(path = subfolder, full.names = TRUE) %>% 
      # Read into dataframe (datetime variables, concent, inlttmp, smpflow)
        map_df(~fread(.x, header = TRUE, skip = 13, 
                      select = c(1, 2, 4, 11, 12)) %>%
        # Convert all concentration values to double
        # Note: pipe is inside map_df - conversion happens for every file 
        # read into dataframe
        mutate(concent = as.double(concent)))
      
    # Make new file name
    new_file_name <- paste(data_cat, id, env, sep = "_")
    print(paste("Writing", new_file_name, "..."))
    
    # Write into appropriate path
    fwrite(main_df, paste0(path_to_data, new_file_name, ".csv"))
    print(paste("Downloaded", env, "data for participant ID", id))
  }
}
```