# Modulair-PM Data Cleaning
This file cleans, and calculates the I/O ratio for Mod-PM data. Output is saved in a `modpm.RData` file that is *untracked* by Github. *All files that depend on using Mod-PM data must be run after running this file.*


## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrplot)
library(rstatix)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/sjatti/Desktop/hepa-summer23')
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Vector of participant numbers' data to be processed
# Skipping 41211 for now because of funky data (look at CSVs)
participants <- c("41181", "41271", "42231", "42281", "42301", "42321",
                  "44241", "44621", "45361", "45411", "45451", "46591")
```

```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/modpm/"

# Set data category
data_cat <- "OH_M"
```

## Helper Functions
The following two functions are used to clean and process the data

### Reformatting timestamps
Necessary since the timestamps from QuantAQ are in a weird format and need to be formatted to a format that works with the openair plotting package. An added column rounding the times to the nearest minute is calculated since two data-frames will be merged later in this script on the basis of time.

```{r}
# Function to reformat time-stamps to time object and round to nearest minute
improve_timestamps <- function(df) {
  df %>%
    # Reformat timestamps to sensible format
    mutate(date = as.POSIXct(timestamp, tz = "America/New_York")) %>%
    # Round times to nearest minute
    mutate(date = round_date(date, unit = "minute"))
}
```

### Particle counts and removing unnecessary data
Alongside particle masses, we are concerned with the counts of particles that fall under certain size bins. To approximate the counts under pm1, sum up bin 0 to 2. Discard all other data columns that are not relevant.
```{r}
# Function to calculate sums of particle counts, remove rest
sum_bins <- function(df) {
  df %>%
    # Sum particle counts
    mutate(pm1num = bin0 + bin1 + bin2) %>%
    # Delete unnecessary columns
    select(pm1:pm1num, sample_temp) %>%
    select(-ends_with("_model_id")) %>%
    rename("temp" = "sample_temp")
}

```

## Main Code Run
### Load all data. 
Ensure that all functions in the code blocks after this loop are loaded. Run this carefully, checking everything, and only once!
```{r}
# Initialize master dataframe for all data
all_df <- data.frame()

# Loop through each participant
for (person in participants) {
  # Loop through each case
  for (case in c("sham", "hepa")) {
    # Loop through each environment
    for (env in c("indoor", "outdoor")) {
      
      # Set file path
      # Filename = data category + participant ID + case 
      # + environment
      file_name <- paste(data_cat, person, case, env, sep = "_")
      
      # File path = path to folder, type csv
      file_path <- paste0(path_to_data, file_name, ".csv")
      
      # Read csv
      df <- read_csv(file_path, show_col_types = FALSE)
      
      # Add the case and environment type (e.g. sham, indoor)
      # And participant ID
      df <- mutate(df, case = case, environment = env,
              participant_id = person)
      
      # Append to main dataframe
      all_df <- rbind(all_df, df)
    }
  }
  print(paste("Loaded participant", person))
}

```

### Process all data to remove unnecessary columns and round time values 
This helps in syncing data across different sensors later - functions here can be found in the 'Helper Functions' section.
```{r}
# Clean data (look at helper functions)
main_df <- all_df %>%
  # Reformat timestamps and round to nearest minute
  improve_timestamps() %>%

  # Calculate particle counts and remove unnecessary columns
  sum_bins() %>%
  
  # Remove repeat readings
  distinct(case, participant_id, date, environment, .keep_all = TRUE)
```

```{r}
# Filter out values over a certain threshold
main_df_indoor <- subset(main_df, environment == "indoor" & pm1 < 200 & pm25 < 250 & pm10 < 1000)
main_df_outdoor <- subset(main_df, environment == "outdoor" & pm1 < 25 & pm25 < 50 & pm10 < 500)

main_df <- rbind(main_df_indoor, main_df_outdoor)
```

### Calculate ratio of indoor and outdoor concentration
To do so, the dataframe is 'spread' to form separate indoor/outdoor columns that are then divided by each other to calculate the ratio. After calculating the ratio, the dataframe is 'gathered' back up to its original shape.

Note: Always run this chunk with the previous chunk.
```{r}
# Spread: Create separate indoor, outdoor columns for particle concentration
main_df_wide <- pivot_wider(main_df,
                  names_from = environment,
                  values_from = c(pm1, pm25, pm10, pm1num, temp))

# Divide indoor by outdoor to calculate ratios
df_ratio <- select(main_df_wide, ends_with("indoor"))/
  select(main_df_wide, ends_with("outdoor"))

# Rename computed columns to ratio
colnames(df_ratio) <- sub("indoor", "ratio", colnames(df_ratio))

# Merge back with joined data-frame
main_df_wide <- cbind(main_df_wide, df_ratio)

# Gather: Return to original shape by re-forming the environment variable
main_df <- pivot_longer(main_df_wide, 
                               pm1_indoor:temp_ratio, 
                               names_to = c(".value", "environment"), 
                               names_sep = "_")
```


### Save Data
The `main_df` dataframe is saved to `cleaned_modpm.RData`
```{r}
save(main_df, file = paste0(path_to_data, "cleaned_modpm.RData"))
```
