# HAFTRAP Diurnal Plotting (for Mod-PM)

## STOP

**HAVE YOU RUN `OH_modpm_cleaning`?**
*This file loads a dataframe created from running `OH_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to create diurnal plots for the HAFTRAP Mod-PM data. It can be modified to create diurnal plots for a single participant vs. multiple/all participants. 

```{r}
# import necessary libraries
library(tidyverse)
library(ggplot2)
library(data.table)
library(lubridate)
```

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = '/home/sjatti/Desktop/hepa-summer23')
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# set path to data
path_to_data <- "data/HAFTRAP/OH/modpm/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

```{r}
# vector of participant numbers' data to be processed (modify to analyze specific participants)
participants <- c("41181", "41211", "41271", "42231", "42281", "42301", "42321",
                  "44241", "44621", "45361", "45411", "45451", "46591")

# filter dataframe based on participants
main_df <- filter(main_df, participant_id %in% participants)
```

```{r}
# function to create a new dataframe with summary data
summary_dataframe <- function(data) {
  data %>%
  group_by(is_wknd, time, case, environment) %>%
  # calculate mean, median, and quartiles for all groups
  summarise(mean_1 = mean(pm1), mean_2.5 = mean(pm25), mean_10 = mean(pm10), median_1 = median(pm1), median_2.5 = median(pm25), median_10 = median(pm10), percent5_1 = quantile(pm1, 0.05), percent25_1 = quantile(pm1, 0.25), percent75_1 = quantile(pm1, 0.75), percent95_1 = quantile(pm1, 0.95), percent5_2.5 = quantile(pm25, 0.05), percent25_2.5 = quantile(pm25, 0.25), percent75_2.5 = quantile(pm25, 0.75), percent95_2.5 = quantile(pm25, 0.95), percent5_10 = quantile(pm10, 0.05), percent25_10 = quantile(pm10, 0.25), percent75_10 = quantile(pm10, 0.75), percent95_10 = quantile(pm10, 0.95))
}
```

```{r}
# round all times to 10 mins
main_df$date_round <- round_date(main_df$date, unit = "10 mins")
      # filter(participant_id == person, case == c, 
      #        environment == env)
```

```{r}
# add column based on day of the week
main_df$wkdy <- wday(main_df$date)

# add boolean column based on whether or not it is a wknd
main_df$is_wknd <- replace(main_df$wkdy, main_df$wkdy > 1 & main_df$wkdy < 7, "FALSE")
main_df$is_wknd <- replace(main_df$is_wknd, main_df$is_wknd == 1 | main_df$is_wknd == 7, "TRUE")
```

```{r}
# isolate time
main_df$time <- as.ITime(main_df$date_round)

# convert back into datetime object (assigns today's date to all)
main_df$time <- as.POSIXct(main_df$time, format="%H:%M:%S")
```

```{r}
# create a new main_df with summary stats
sum_main_df <- summary_dataframe(main_df)

# restructure main_df to make it easier to graph
graph_main_df <- pivot_longer(sum_main_df, mean_1:percent95_10, names_to = c(".value", "particle"), names_sep = "_")
```

```{r}
# function to create diurnal plots + save as an svg
diurnal_plot <- function(data, case_str, env_str, pm_str) {
  
  # create labeller for faceted graphs
  wknd_labels <- as_labeller(c("TRUE" = "Weekend",
                    "FALSE" = "Weekday"))
  
  # filter data for pm, case, and env
  data <- filter(data, particle == pm_str & case == case_str & environment == env_str)
  
  # vector for all graph labels (only way to do f-string + subscript at the same time)
  graph_labels <- c("Daily", str_to_title(env_str), pm_str, "Trends", "During", str_to_title(case_str), "Period", "[ug/m^3]", "Concentration", "MOD-PM")
  
  # # svg file to save graph
  # svg(paste(case_str, env_str, pm_str, sep = "_"))
  
  ggplot(data) +
    
    # plot the 5th-95th and 25th-75th percentiles
    geom_ribbon(aes(x = time, ymin = percent5, ymax = percent95, fill = "#cfcfcf")) +
    geom_ribbon(aes(x = time, ymin = percent25, ymax = percent75, fill = "#9a9a9a")) +
    
    # plot the mean and median
    geom_line(aes(x = time, y = mean, color = "red")) +
    geom_line(aes(x = time, y = median, color = "blue")) +
    
    # facet based on wkdy vs. wknd
    facet_grid(is_wknd ~ ., labeller = wknd_labels) +
    
    # label legend for mean and median
    scale_color_identity(name = "Averages",
                            breaks = c("red", "blue"),
                            labels = c("Mean", "Median"),
                            guide = "legend") +
    
    # label legend for percentiles
    scale_fill_identity(name = "Percentiles",
                            breaks = c("#cfcfcf", "#9a9a9a"),
                            labels = c("5th - 95th", "25th - 75th"),
                            guide = "legend") +
    
    # modifiy y-axis scaling
    scale_x_datetime(date_breaks = "2 hours", date_labels = "%H:00") +
    
    # label title and axes using vector above
    labs(title = bquote(.(graph_labels[10]) ~ .(graph_labels[1]) ~ .(graph_labels[2]) ~ PM[.(graph_labels[3])] ~ .(graph_labels[4]) ~ .(graph_labels[5]) ~ .(graph_labels[6]) ~ .(graph_labels[7])),
         x = "Time of Day (hrs)",
         y = bquote(PM[.(graph_labels[3])] ~ .(graph_labels[9]) ~ .(graph_labels[8]))) +
    
    # # scale y-axis (log vs. linear)
    scale_y_log10() +
    scale_y_log10(limits = c(1e-2,1e2)) +
    # ylim(0,75) +
    
    # add fun theme
    theme_bw()
  
  # # close the graphics device
  # dev.off()
}
```

```{r}
# plot all combos for diurnals
diurnal_plot(graph_main_df, "hepa", "indoor", "1")
diurnal_plot(graph_main_df, "hepa", "indoor", "2.5")
diurnal_plot(graph_main_df, "hepa", "indoor", "10")

diurnal_plot(graph_main_df, "hepa", "outdoor", "1")
diurnal_plot(graph_main_df, "hepa", "outdoor", "2.5")
diurnal_plot(graph_main_df, "hepa", "outdoor", "10")

diurnal_plot(graph_main_df, "hepa", "ratio", "1")
diurnal_plot(graph_main_df, "hepa", "ratio", "2.5")
diurnal_plot(graph_main_df, "hepa", "ratio", "10")

diurnal_plot(graph_main_df, "sham", "indoor", "1")
diurnal_plot(graph_main_df, "sham", "indoor", "2.5")
diurnal_plot(graph_main_df, "sham", "indoor", "10")

diurnal_plot(graph_main_df, "sham", "outdoor", "1")
diurnal_plot(graph_main_df, "sham", "outdoor", "2.5")
diurnal_plot(graph_main_df, "sham", "outdoor", "10")

diurnal_plot(graph_main_df, "sham", "ratio", "1")
diurnal_plot(graph_main_df, "sham", "ratio", "2.5")
diurnal_plot(graph_main_df, "sham", "ratio", "10")
```