# HAFTRAP Data Cleaning and Summaries - multiple (for CPC)

## STOP

**HAVE YOU RUN `OH_modpm_cleaning` AND `OH_cpc_cleaning`?**
*This file loads dataframes created from running `OH_modpm_cleaning` AND `OH_cpc_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script calculates correlation coefficients for data collected from CPC and Mod-PM sensors. 

Results can be found in `summary/HAFTRAP/OH/s_OH_MC_corr.csv` and are further analyzed in `insight/HAFTRAP/OH/OH_correlations`.

As a by-product to calculating correlation coefficients, this script also generates a data-frame stored in `merged.RData` for use in `OH_correlations`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(corrr)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Get relevant participant IDs from OH_participants
participants <- as.character(
  na.omit(read_excel("data/HAFTRAP/OH/OH_participants.xlsx")$modpm_cpc))

# Set path to data
path_to_data <- "cleaning/HAFTRAP/OH/"
```


## Helper Functions:
### Calculating 10 minute averages
When correlating Mod-PM and CPC data, we want to make sure that there has been enough time for air to mix, therefore we calculate 10-minute averages of the data.
```{r}
# Round over 10 minute averages
round_10min <- function(df, vec) {
  df %>%
  # Create new column with rounded dates
  mutate(date_round = round_date(date, "10 minutes")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean over ten minutes
  summarize(across(all_of(vec), mean), .groups = "drop") %>%
  # Rename data column
  rename(date = date_round)
}
```

### Calculating correlation coefficients
The following function is not part of the main pipeline to calculate summary statistics, but is used to clean and process the data. 

Correlation coefficients tell us a lot about relationships between variables. This function calculates the correlation matrix for a particular case for a single participant and reshapes it into a paired-list format
```{r}
# Function to calculate correlation matrices
get_corr <- function(df, p, c) {
  df %>%
    # Filter for case
    filter(participant_id == p, case == c) %>%
    # Select all columns except date
    select_if(is.numeric) %>%
    # Remove NaN values
    drop_na() %>%
    # Calculate correlations (suppress output message)
    correlate(quiet = TRUE) %>%
    # Reshape into paired-list format
    stretch(remove.dups = TRUE)
}
```

## Load data

Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))

# Create copy of main_df, and delete original
modpm_df <- main_df %>%
  # # Average Mod-PM data over 10 minutes
  round_10min(c("pm1", "pm25", "pm10", "pm1num", "temp"))

rm(main_df)
```

CPC - filter data to whole minutes
```{r}
# Load CPC data from RData file
load(paste0(path_to_data, "cleaned_cpc.RData"))

# Create copy of main_df
cpc_df <- main_df %>%
  # Average CPC data over 10 minutes
  round_10min(c("concent", "inlttmp"))

# Delete original
rm(main_df)
```



## Main Code Run

### Merge CPC and Mod-PM Data
Merge is conducted over matching dates for each participant in a particular environment and case
```{r}
main_df <- merge(cpc_df, modpm_df, 
                   by = c("date", "participant_id", "environment", "case"))
```

### Calculate correlation matrices
Helps better understand the relationship between different variables in sham and true HEPA conditions for each participant. Matrices are reshaped into longer pair format for stacking all the data into single dataframe. (Warning: this code takes a while to run.)

```{r}
# Spread: Create separate indoor, outdoor columns for particle concentration
main_df_wide <- pivot_wider(main_df,
                  names_from = environment,
                  values_from = c(concent:temp))

# Create dataframe to hold correlation coefficients
corr_df <- data.frame()
# Loop through each participant
for (person in participants) {
  # Loop through each case
  for (case in c("sham", "hepa")) {
    # Use wide dataframe where indoor/outdoor are separate columns
    main_df_wide %>%
      # Get correlation coefficients (look at helper function)
      get_corr(person, case) %>%
      # Add the case and participant ID
      mutate(case = case, participant_id = person) %>%
      # Drop NaNs
      drop_na() -> df
    # Append to main dataframe
    corr_df <- rbind(corr_df, df)
    print(paste("Computed correlation coefficients for", person, case))
  }
}
```

## Save to file
Summary statistics are saved in a series of csv files. The naming convention of these files is specified by in `data_guide.md`
```{r}
# Save summary statistic files to CSVs
write.csv(corr_df, "summary/HAFTRAP/OH/s_OH_MC_corr.csv")
```

### Save Data
The `main_df` dataframe with merged CPC and Mod-PM data is saved to `merged.RData`
```{r}
save(main_df, file = paste0("summary/HAFTRAP/OH/merged.RData"))
```