# Sanity Check Mod-PM Data

## STOP

**HAVE YOU RUN `OH_modpm_cleaning`?**
*This file loads a dataframe created from running `OH_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to sanity check data produced from `OH_modpm_cleaning`. Time-series plots are created of each sensor and given a quick visual check. If required, appropriate edits are to be made to `OH_modpm_cleaning`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(ggplot2)
```
Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/sjatti/Desktop/hepa-summer23')
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```
Participant IDs
```{r}
# Vector of participant numbers' data to be processed
participants <- c("44241", "41181", "41211", "41271", "42231", "42281", "42301", "42321", "44621", "45361", "45411", "45451", "46591")
```

Paths
```{r}
# Set path to data
path_to_data <- "data/HAFTRAP/OH/modpm/"
```


## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Plotting
```{r}
# Input participant ID
person <- readline("Enter participant ID: ")
while (!(person %in% participants)) {
  # Input participant ID
  person <- readline("Error! Invalid ID, Re-enter: ")
}

# Plot stuff
for (env in c("indoor", "outdoor", "ratio")) {
  y_check <- case_match(
    env,
    "indoor" ~ 10,
    "outdoor" ~ 10,
    "ratio" ~ 10
  )
  for (c in c("sham","hepa")) {
    # Filter data to appropriate subset
    df <- main_df %>%
      filter(participant_id == person, case == c, 
             environment == env)
    
      # Plot time-series of subset with an appropriate horizontal line
      print(ggplot(df) +
        geom_line(aes(x = date, y = pm1)) +
        geom_hline(aes(yintercept = y_check)) +
        ggtitle(paste(person, c, env, "pm1")))
      
      print(ggplot(df) +
        geom_line(aes(x = date, y = pm25)) +
        geom_hline(aes(yintercept = y_check)) +
        ggtitle(paste(person, c, env, "pm25")))
      
      print(ggplot(df) +
        geom_line(aes(x = date, y = pm10)) +
        geom_hline(aes(yintercept = y_check)) +
        ggtitle(paste(person, c, env, "pm10")))
      
      print(ggplot(df) +
        geom_line(aes(x = date, y = pm1num)) +
        geom_hline(aes(yintercept = y_check)) +
        ggtitle(paste(person, c, env, "pm1_num")))
    
    # Pause and print
    Sys.sleep(2)
    print(paste("Plotted for", person, c, env))
  }
}
```