# HAFTRAP Temperature Collection

## STOP

**HAVE YOU RUN `OH_modpm_cleaning` and `OH_modulair_cleaning`?**
*This file loads dataframes created from running `OH_modpm_cleaning` and `OH_modulair_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to collate the temperature data for all HAFTRAP deployments for Olin.

Results can be found in `summary/HAFTRAP/OH/s_OH_M_quants.csv`.

## Set up
Load libraries, define file paths
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/HAFTRAP/OH/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))

modpm_df <- main_df
```

```{r}
# Load Modulair data from RData file
load(paste0(path_to_data, "cleaned_modulair.RData"))

modulair_df <- main_df
```

## Isolate and join temperature data
```{r}
var_list <- c("date", "temp", "environment", "participant_id", "case")

rbind(select(modpm_df, all_of(var_list)), select(modulair_df, all_of(var_list))) %>%
  filter(environment %in% c("indoor", "outdoor")) -> temperature_df
```

## Export data as CSV
Make sure to create the `artifacts/HAFTRAP/OH/temperature` folder beforehand!
```{r}
write_csv(temperature_df, "artifacts/HAFTRAP/OH/temperature/temperature_data.csv")
```

