# HAFTRAP Plotting Violinplots

## STOP

**HAVE YOU RUN `OH_modpm_cleaning` AND `OH_cpc_cleaning`?**
*This file loads dataframes created from running `OH_modpm_cleaning` AND `OH_cpc_cleaning`. Make sure you run those files first (if you haven't already) before running this file.*

This script is used to plot boxplots using all the data from the Olin HAFTRAP study.

Plots can be found here and saved in the folder`artifacts/HAFTRAP/OH/violinplot`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Path to data
```{r}
# Set path to data
path_to_data <- "cleaning/HAFTRAP/OH/"
```

## Load data

Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))

# Create copy of main_df, and delete original
modpm_df <- main_df
rm(main_df)
```

CPC - filter data to whole minutes
```{r}
# Load CPC data from RData file
load(paste0(path_to_data, "cleaned_cpc.RData"))

# Create copy of main_df
cpc_df <- main_df

# Delete original
rm(main_df)
```

## Plotting for Indoor and Outdoor raw concentrations
Define lists for plot titles
```{r}
titles_raw <- list(pm1 = "PM 1 Concentration", pm25 = "PM 2.5 Concentration",
               pm10 = "PM 10 Concentration", pm1num = "PM 1 Particle Count",
               concent = "CPC Particle Count")

titles_ratio <- list(pm1 = "PM 1 I/O Ratio", pm25 = "PM 2.5 I/O Ratio",
               pm10 = "PM 10 I/O Ratio", pm1num = "PM 1 Count I/O Ratio",
               concent = "CPC Count I/O Ratio")
```

## Helper functions:
### Plotting function
The following function plots the raw data for a given dataset
```{r}
plot_raw <- function(df_united, p_type, val) {
  # Filter dataset for particle type, dropping NaNs and Infs
  df_united %>% 
    filter(type == p_type, is.finite(reading)) %>%
    # Order sham then hepa by creating factors
    mutate(across(case, \(x) factor(x, levels=c("sham", "hepa")))) %>%
    
  # Plot filtered indoor data
  ggplot() +
    # Create violinplot
    geom_violin(aes(x = situation, y = reading, fill = case)) +
    # Convert y-axis scale to log
    scale_y_continuous(trans = "log10") +
    # Set title, caption with percentage reduction
    labs(title = paste("Overall", titles_raw[[p_type]]),
         caption = paste0("Mean reduction = ", round(val, 1), "%")) -> my_plot
  
  # Save plot
  ggsave(filename = paste0("artifacts/HAFTRAP/OH/violinplot/", "OH_", p_type,
                           "_raw.png"), plot = my_plot)
  
  # Display plot
  my_plot
}
```

### Percentage reduction calculating
The following function calculates mean percentage reduction for a specified particle type. 

Note: this function doesn't calculate overall reduction, but the mean of the mean reductions for each participant. Done for consistency with output of `OH_efficacy`.
```{r}
# Calculate mean percentage reduction for particle in environment
get_redu <- function(df, env, p_type) {
  vals <- df %>% 
    # Filter relevant environment
    filter(environment == env) %>%
    # Calculate mean for each case for each participant
    group_by(case, participant_id) %>% 
    summarize(across(all_of(p_type), mean))
  
  # Calculate percentage reduction for each participant and store in vector
  vec <- 100 * (vals[vals$case == "sham", 3] - vals[vals$case == "hepa", 3])/
    vals[vals$case == "sham", 3]
  
  # Return mean of vector (indexing cause vector is technically a dataframe)
  mean(vec[[1]])
}
```


### CPC
```{r}
# Filter, order data appropriately
cpc_united <- cpc_df %>%
  filter(environment != "ratio") %>%
  unite("situation", environment, case, sep = ", ", remove = FALSE) %>%

  # Gather dataframe to create single variable for measurements
  pivot_longer(concent:inlttmp, names_to = "type", values_to = "reading")
```

Plotting for concent
```{r}
# Calculate percentage reduction
val <- get_redu(cpc_df, "indoor", "concent")
# Plot data
plot_raw(cpc_united, "concent", val)
```
### Mod-PM
```{r}
# Filter, order data appropriately
modpm_united <- modpm_df %>%
  filter(environment != "ratio") %>%
  unite("situation",environment, case, sep = ", ", remove = FALSE) %>%

  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm10, pm1num, temp), 
               names_to = "type", values_to = "reading")
```

Plotting for Mod-PM
```{r}
for (p_type in c("pm1", "pm25", "pm10", "pm1num")) {
  # Calculate percentage reduction
  val <- get_redu(modpm_df, "indoor", p_type)
  # Plot data
  print(plot_raw(modpm_united, p_type, val))
}
```
## Plotting for Indoor/Outdoor Ratio
### Plotting function
The following function plots the indoor/outdoor ratio for a given dataset
```{r}
plot_ratio <- function(df_ratio, p_type, val) {
  # Filter dataset for particle type, dropping NaNs and Infs
  df_ratio %>% 
    filter(type == p_type, is.finite(reading)) %>%
    # Order sham then hepa by creating factors
    mutate(across(case, \(x) factor(x, levels=c("sham", "hepa")))) %>%
    
  # Plot filtered ratio data
  ggplot() +
    # Create violinplot
    geom_violin(aes(x = case, y = reading, fill = case)) +
    # Convert y-axis scale to log
    scale_y_continuous(trans = "log10") +
    # Set title, caption with percentage reduction
    labs(title = paste("Overall", titles_ratio[[p_type]]),
         caption = paste0("Mean reduction = ", round(val, 1), "%")) -> my_plot
  
  # Save Plot
  ggsave(filename = paste0("artifacts/HAFTRAP/OH/violinplot/", "OH_", p_type,
                           "_ratio.png"), plot = my_plot)
  
  # Display plot
  my_plot
}
```

### CPC
```{r}
# Filter, order data appropriately
cpc_ratio <- cpc_df %>%
  filter(environment == "ratio") %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(concent:inlttmp, names_to = "type", values_to = "reading")
```

Plotting for concent
```{r}
# Calculate percentage reduction
val <- get_redu(cpc_df, "ratio", "concent")
# Plot data
plot_ratio(cpc_ratio, "concent", val)
```

### Mod-PM
```{r}
# Filter, order data appropriately
modpm_ratio <- modpm_df %>%
  filter(environment == "ratio") %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm10, pm1num, temp), 
               names_to = "type", values_to = "reading")
```

Plotting for Mod-PM
```{r}
for (p_type in c("pm1", "pm25", "pm10", "pm1num")) {
  # Calculate percentage reduction
  val <- get_redu(modpm_df, "ratio", p_type)
  # Plot data
  print(plot_ratio(modpm_ratio, p_type, val))
}
```