# HAFTRAP Plotting Boxplots

## STOP

**HAVE YOU RUN `OH_modpm_cleaning`?**
*This file loads a dataframe created from running `OH_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to plot boxplots using all the data from the Olin HAFTRAP study.

Plots can be found here and saved in the folder`summary/HAFTRAP/OH/`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
library(ggplot2)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/vkuchhal/Documents/hepa-summer23')
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

Participant IDs and corresponding time map tables
```{r}
# Vector of participant numbers' data to be processed
participants <- c("41271", "42281", "45361", "44621", "45451")

### Set paths
# # For CPC
# Set path to data
path_to_data_C <- "data/HAFTRAP/OH/cpc/"

# # For Mod-PM
# Set path to data
path_to_data_M <- "data/HAFTRAP/OH/modpm/"
```

## Load data

Mod-PM
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data_M, "cleaned_modpm.RData"))

# Create copy of main_df, and delete original
modpm_df <- main_df %>%
  # Filtering out 41211 for now in Mod-PM because of funky data (look at CSVs)
  filter(participant_id != "41211")
rm(main_df)
```

CPC - filter data to whole minutes
```{r}
# Load CPC data from RData file
load(paste0(path_to_data_C, "cleaned_cpc.RData"))

# Create copy of main_df
cpc_df <- main_df

# Delete original
rm(main_df)
```

## Plotting for Indoor and Outdoor raw concentrations
### Titles and axis labels
Define lists to use in labelling axes and plot titles
```{r}
titles_raw <- list(pm1 = "PM 1 Concentration", pm25 = "PM 2.5 Concentration",
               pm10 = "PM 10 Concentration", pm1num = "PM 1 Particle Count",
               concent = "CPC Particle Count")

titles_ratio <- list(pm1 = "PM 1 I/O Ratio", pm25 = "PM 2.5 I/O Ratio",
               pm10 = "PM 10 I/O Ratio", pm1num = "PM 1 Count I/O Ratio",
               concent = "CPC Count I/O Ratio")
```

## Helper functions:
### Plotting function
The following function plots the raw data for a given dataset
```{r}
plot_raw <- function(df_select, p_type, val) {
  my_plot <- ggplot(df_select) +
    geom_violin(aes(x = situation, y = reading, fill = case)) +
    scale_y_continuous(trans = "log10") +
    labs(title = paste("Overall", titles_raw[[p_type]]),
         caption = paste0("Mean indoor reduction = ", round(val, 1), "%"))
  
  # ggsave(filename = paste0("artifacts/", p_type, "_raw.png"), plot = my_plot)
  my_plot
}
```

### Percentage reduction calculating
The following function calculates mean percentage reduction for a specified particle type
```{r}
get_redu <- function(df, env, p_type) {
  vals <- df %>% 
    filter(environment == env) %>% 
    group_by(case, participant_id) %>% 
    summarize(across(all_of(p_type), mean))
  
  vec <- 100 * (vals[vals$case == "sham", 3] - vals[vals$case == "hepa", 3])/
    vals[vals$case == "sham", 3]
  
  mean(vec[[1]])
}
```


### CPC
```{r}
# Filter, order data appropriately
cpc_united <- cpc_df %>%
  filter(environment != "ratio") %>%
  unite("situation", environment, case, sep = ", ", remove = FALSE) %>%

  # Gather dataframe to create single variable for measurements
  pivot_longer(concent:inlttmp, names_to = "type", values_to = "reading")


```

Plotting for concent
```{r}
# Calculate percentage reduction
val <- get_redu(cpc_df, "indoor", "concent")

cpc_select <- cpc_united %>% filter(type == "concent", is.finite(reading))
plot_raw(cpc_select, "concent", val)
```
### Mod-PM
```{r}
# Filter, order data appropriately
modpm_united <- modpm_df %>%
  filter(environment != "ratio") %>%
  unite("situation",environment, case, sep = ", ", remove = FALSE) %>%

  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm10, pm1num, temp), 
               names_to = "type", values_to = "reading")
```

Plotting for Mod-PM
```{r}
for (p_type in c("pm1", "pm25", "pm10", "pm1num")) {
  # Calculate percentage reduction
  val <- get_redu(modpm_df, "indoor", p_type)

  modpm_select <- modpm_united %>% filter(type == p_type, is.finite(reading))
  print(plot_raw(modpm_select, p_type, val))
}
```
## Plotting for Indoor/Outdoor Ratio
### Plotting function
The following function plots the indoor/outdoor ratio for a given dataset
```{r}
plot_ratio <- function(df_select, p_type, val) {
  my_plot <- ggplot(df_select) +
    geom_violin(aes(x = case, y = reading, fill = case)) +
    scale_y_continuous(trans = "log10") +
    labs(title = paste("Overall", titles_ratio[[p_type]]),
         caption = paste0("Mean reduction = ", round(val, 1), "%"))
  
  # ggsave(filename = paste0("artifacts/", p_type, "_ratio.png"), 
  #        plot = my_plot)
  my_plot
}
```

### CPC
```{r}
# Filter, order data appropriately
cpc_ratio <- cpc_df %>%
  filter(environment == "ratio") %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(concent:inlttmp, names_to = "type", values_to = "reading")
```

Plotting for concent
```{r}
val <- get_redu(cpc_df, "ratio", "concent")

cpc_select <- cpc_ratio %>% filter(type == "concent", is.finite(reading))
plot_ratio(cpc_select, "concent", val)
```

### Mod-PM
```{r}
# Filter, order data appropriately
modpm_ratio <- modpm_df %>%
  filter(environment == "ratio") %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm10, pm1num, temp), 
               names_to = "type", values_to = "reading")
```

Plotting for Mod-PM
```{r}
for (p_type in c("pm1", "pm25", "pm10", "pm1num")) {
  val <- get_redu(modpm_df, "ratio", p_type)
  
  modpm_select <- modpm_ratio %>% filter(type == p_type, is.finite(reading))
  print(plot_ratio(modpm_select, p_type, val))
}
```