# Sham vs. HEPA - is there a difference?
This notebook investigates whether the installation of air purifiers had a notable impact on indoor air quality across various summary metrics. It uses summary data found in `summary/HAFTRAP/SC`

```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
path_to_data <- "summary/HAFTRAP/SC/"
```

## Plotting Function
```{r}
## Global variables for plotting function
# Maps environment to title
env_title <- list(indoor = "Reduction in Raw Indoor Readings", 
                  ratio = "Reduction in Indoor/Outdoor Ratio")

particle_list <- c("mod_pm1", "mcpc_counts", "acsm_org", "t200u_no", "t200u_no2", "t300_co", "acsm_ratio_55_57")

# Sets order for displaying particles in heatmap
particle_order <- c("PM 1 Conc.", "UFP Count", "ACSM Org.", 
                   "NO", "NO2", "CO", "ACSM Ratio")

# Sets order for displaying summary stats in heatmap
summary_order <- c("Mean", "95th Percentile", "75th Percentile", "Median",
                     "25th Percentile", "5th Percentile")

plot_heatmap <- function(joined_data, env) {
  # Pre-process data before ploting
  joined_data %>%
    
    # Reshape for heatmap
    pivot_longer(mean:q95, names_to = "sum_type", values_to = "stat") %>%
    
    # Recode for accurate labels
    mutate(type = recode(type, "pm1" = "PM 1 Conc.", "mcpc_counts" = "UFP Count", "acsm_org" = "ACSM Org.", 
                   "t200u_no" = "NO", "t200u_no2" = "NO2", "t300_co" = "CO", "acsm_ratio_55_57" = "ACSM Ratio"),
         sum_type = recode(sum_type, "mean" = "Mean", "median" = "Median",
                           "q5" = "5th Percentile", "q25" = "25th Percentile",
                           "q75" = "75th Percentile", 
                           "q95" = "95th Percentile")) %>%
    
    # Arrange axes in specified order
    mutate(type = factor(type, levels = particle_order)) %>%
    mutate(sum_type = factor(sum_type, levels = summary_order)) %>%
    
    drop_na() %>%
    
    # Plot heatmap
    ggplot(aes(x = type, y = sum_type)) +
      # Fill with value of stat
      geom_tile(aes(fill = stat)) +
      # Print in value of stat
      geom_text(aes(label = paste0(round(stat, 1), "%"))) +
    
      # Set color scale
      scale_fill_gradient2(low = "#ff7f7f", high = "#DeF7E9", mid = "white",
                           limits = c(-100, 100), guide = "colorbar",
                           oob = scales::squish) +
      # Set axis labels, title
      labs(title = env_title[[env]], 
           x = "Particle Type", y = "Summary Statistic", 
           fill = "Reduction (%)") +
      # Set theme
      theme_minimal() -> my_plot
  # Save plot
  ggsave(filename = paste0("artifacts/HAFTRAP/SC/reduction/", "SC_", env,
                           "_heatmap.png"), plot = my_plot)
  # Display plot
  my_plot
}
```

## Load Data
### Mod-PM
```{r}
filepath <- paste0(path_to_data, "s_SC_RASCT_quants.csv")
all_stats <- read_csv(filepath, show_col_types = FALSE)
```

### Reduction in raw Indoor Concentrations
```{r}
joined_indoor <- all_stats %>%
  filter(participant_id == "all", environment == "indoor", case == "redu") %>%
  filter(type %in% particle_list)
```

### Reduction in Indoor/Outdoor Ratio
```{r}
# Calculate CPC, Mod-PM averages for I/O ratio
joined_ratio <- all_stats %>%
  filter(participant_id == "all", environment == "ratio", case == "redu") %>%
  filter(type %in% particle_list)
```

## Making pretty tables
```{r}
plot_heatmap(joined_indoor, "indoor")
plot_heatmap(joined_ratio, "ratio")
```



## Save to file
Save the two tables with efficacy in CSV files
```{r}
# Save summary statistic files to CSVs
write.csv(joined_indoor,
          "artifacts/HAFTRAP/SC/reduction/SC_indoor_reduction.csv")
write.csv(joined_ratio,
          "artifacts/HAFTRAP/SC/reduction/SC_ratio_reduction.csv")
```

