# HAFTRAP Summary Statistics for Decay Data (for Modulair)

## STOP

**HAVE YOU RUN `SC_mod_decay_analysis`?**
*This file loads a dataframe created from running `SC_mod_decay_analysis`. Make sure you run that file first (if you haven't already) before running this file.*

This file is used to calculate summary statistics from the decay constants for the SCOPE HAFTRAP Modulair data.

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = '/home/sjatti/Desktop/hepa-summer23')
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "summary/HAFTRAP/SC/"

# Load modpm data from csv file
decay_constants <- read.csv(paste0(path_to_data, "SC_M_decay.csv"))
```

```{r}
# calculate peak width w/ peak and valley indices
decay_constants$peak_wdth <- decay_constants$valley_idx - decay_constants$peak_idx
```

```{r}
# filter out decays constants greater than 2
decay_constants <- decay_constants[decay_constants$k_val < 2, ]  
```

```{r}
# function to calculate mean, median, and sd for k-value, peak width, and peak height
my_summarise <- function(grouped_df) {
  grouped_df %>%
    summarise(mean_k = mean(k_val),
            median_k = median(k_val), 
            sd_k = sd(k_val),
            mean_hgt = mean(peak_hgt),
            median_hgt = median(peak_hgt), 
            sd_hgt = sd(peak_hgt),
            mean_wdth = mean(peak_wdth),
            median_wdth = median(peak_wdth), 
            sd_wdth = sd(peak_wdth),
            .groups = 'drop')
}
```

```{r}
# Calculate summary statistics for each participant
participant_summary <- decay_constants %>%
  # Group by participant ID, case, and particle
  group_by(participant, case, particle) %>%
  # Calculate summary statistics
  my_summarise()
```

```{r}
# Calculate summary statistics over all the data
particle_summary <- decay_constants %>%
  # Group by case and particle
  group_by(case, particle) %>%
  # Calculate summary statistics
  my_summarise() %>%
  # Add participant ID column as 'all'
  mutate(participant = "all")
```

```{r}
# Calculate summary statistics over all the data
overall_summary <- decay_constants %>%
  # Group by case
  group_by(case) %>%
  # Calculate summary statistics
  my_summarise() %>%
  # Add participant column as 'all'
  mutate(participant = "all") %>%
  # Add particle column as 'all'
  mutate(particle = "all")
```

```{r}
# Bind together participant, particle, and overall summary statistics
summary <- rbind(participant_summary, particle_summary, overall_summary)
```

```{r}
# create new column for particle type
summary <- pivot_longer(summary, mean_k:sd_wdth,
               names_to = c(".value", "type"), names_sep = "_")
```


### Calculate percentage reduction in concentration from sham to hepa
Follows the same logic as calculating indoor-outdoor ratios. Spread to form 
sham/hepa columns that are used to calculate percent reduction in corresponding
summary pollution concentrations. Then gather dataframe back to original shape.
```{r}
# Spread: Create separate sham, hepa summary statistics columns
summary_wide <- pivot_wider(summary, 
                            names_from = case, values_from = mean:sd)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("sham")) - 
  select(summary_wide, ends_with("hepa"))) / 
  select(summary_wide, ends_with("sham"))

# Rename computed columns to 'redu' for percentage reduction
colnames(summary_redu) <- sub("sham", "redu", colnames(summary_redu))
```

```{r}
# Reshape data and pipe into variable 'summary_normal'
summary_wide <- summary_wide %>%
  # Merge back with joined data-frame
  cbind(summary_redu)
```

```{r}
# multiply k-val reduction values by -1 to make it percent increase
summary_wide$mean_redu[summary_wide$type == 'k'] <- summary_wide$mean_redu*(-1)
summary_wide$median_redu[summary_wide$type == 'k'] <- summary_wide$median_redu*(-1)
summary_wide$sd_redu[summary_wide$type == 'k'] <- summary_wide$sd_redu*(-1)
```

```{r}
summary_wide %>%
  # Gather: Return to original shape by removing the 'sham/hepa/redu'
  # suffixes by re-forming the case variable
  pivot_longer(mean_hepa:sd_redu,
               names_to = c(".value", "case"), names_sep = "_") %>%
  # Filter out all NaN and Inf values
  filter_all(all_vars(!is.infinite(.))) -> summary_normal
```

```{r}
# create a new decay folder (might already exist)
path_name <- "artifacts/HAFTRAP/OH/decay"
dir.create(path_name)
```

```{r}
# Save final decay summary dataframe to CSV
write.csv(summary_normal, paste0(path_name, "/OH_M_decay_summary.csv"))
```
