---
editor_options: 
  markdown: 
    wrap: 72
---

# Roxbury Plume Plotting (for Mod-PM)

Plot various plumes before and after HEPA installation to compare their
heights, widths, decay constants, etc.

## STOP

**HAVE YOU RUN `IB_modpm_cleaning`?** *This file loads a dataframe
created from running `IB_modpm_cleaning`. Make sure you run that file
first (if you haven't already) before running this file.*

This file is used to analyze decay constants and complete plume
visualizations for the East Boston Mod-PM data.

```{r}
# import necessary libraries
library(tidyverse)
library(data.table)
library(scales)
library(pracma)
```

```{r, setup, include=FALSE}
# set working directory
knitr::opts_knit$set(root.dir = "~/AirPartners/hepa-summer23")
```

```{r}
# check file path for working directory
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# set path to data
path_to_data <- "cleaning/Roxbury/IB/"
```

## Load data

```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Main Code Run

```{r}
# Pre-processing to add relevant columns
my_df <- main_df %>%
  # Round date to nearest ten minutes
  mutate(date_round = round_date(date, unit = "10 mins")) %>%
  # Get day of the week as integer from 1 to 7
  mutate(wkdy = data.table::wday(date)) %>%
  # Classify as weekday or weekend
  mutate(is_wknd = wkdy %in% c(1, 7)) %>%
  # Extract time of the day from datetime
  mutate(time = as.POSIXct(as.ITime(date_round), format = "%H:%M:%S")) %>%
  # Arrange by date
  arrange(dat)
my_df
```

```{r}
# Calculate summary statistics for every ten minutes
graph_main_df <- my_df %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(c(pm1, pm25, pm10), 
               names_to = "particle", values_to = "reading") %>%
  # For every 10 minutes for given case, environment, particle, weekday/end
  group_by(is_wknd, time, case, environment, particle) %>%
  # Find summary statistics
  summarise(mean = mean(reading),
          median = median(reading), 
          q5 = quantile(reading, probs = 0.05), 
          q25 = quantile(reading, probs = 0.25),
          q75 = quantile(reading, probs = 0.75),
          q95 = quantile(reading, probs = 0.95),
          sd = sd(reading),
          .groups = 'drop')
graph_main_df
```

```{r}
# Calculate average baseline parameter and mean parameters, which will be used
# to define the baseline and the minimum height of peaks
pm25_indoor <- graph_main_df[graph_main_df$environment=='indoor' & graph_main_df$particle=='pm25',]
pm_baseline <- pm25_indoor %>%
  group_by(case) %>%
  summarise(mean = mean(mean))
pm_min_heights <- pm25_indoor %>%
  group_by(case) %>%
  summarise(mean = mean(q75))
```


# Helper functions for plume analysis

```{r}
# Find the first local minimum before/after the index of peak_row in y
# that occurs below the min_threshold.
get_local_valleys <- function(data, peaks, min_threshold, min_length) {
  # Iterate through rows of peaks matrix and get valleys before peaks
  for (row in 1:nrow(peaks)) {
    idx <- peaks[row, 2]
    n <- TRUE
    # create while loop to check for local minima
    while (n) {
      # if we're at the end of y, break out of loop
      if (idx == 1) {
        n <- FALSE
      } else {
        # otherwise, get slope (approximately)
        slope <- data[idx] - data[idx - 1]
        if (data[idx] <= min_threshold && slope < 0) {
          n <- FALSE
        } else {
          idx <- idx - 1
        }
      }
    }
    # set values of matrix so that column 3 is index of valley before peak
    peaks[row, 3] <- idx
  }
  # Repeat process for valleys after peaks
  for (row in 1:nrow(peaks)) {
    idx <- peaks[row, 2]
    n <- TRUE
    count <- as.numeric(0)
    # create while loop to check for local minima
    while (n) {
      # if we're at the end of y, break out of loop
      if (idx == length(data)) {
        n <- FALSE
      } else {
        # otherwise, get slope (approximately)
        slope <- data[idx + 1] - data[idx]
        if (data[idx] <= min_threshold && slope > 0 && count >= min_length) {
          n <- FALSE
        } else {
          idx <- idx + 1
          count <- count + 1
        }
      }
    }
    # set values of matrix so that column 4 is index of valley after peak
    peaks[row, 4] <- idx
  }
  # Convert to dataframe and remove duplicate valleys
  colnames(peaks) <- c("peak_hgt", "peak", "valley_b", "valley_a")
  peaks_df <- as.data.frame(peaks)
  
  # Sorting peaks_df by peak in descending order
  peaks_df <- peaks_df[order(-peaks_df$peak_hgt),]
  
  # Drop duplicate valleys, keeping highest peak for each duplicate
  result_peaks <- peaks_df[!duplicated(peaks_df[, c("valley_b", "valley_a")]),]
  result_peaks
}
```

```{r}
# Exponential curve fitting for air quality data; returns dataframe containing
# k-values of curves (MK II)
# NOTE: if the curve is too small or doesn't contain enough information for this
# computation, peak will be omitted from final return
curve_fitting <- function(data, peaks) {
  # Create empty DataFrame for storing k values
  alphas.data <- data.frame(
    "peak_idx" = numeric(0),
    "valley_idx_b" = numeric(0),
    "valley_idx_a" = numeric(0),
    "peak_hgt" = numeric(0),
    "k_val" = numeric(0),
    "conv_tol" = numeric(0),
    "cumm_pm" = numeric(0)
  )
  # Define parameters for curve fitting function for each row
  for (row in 1:nrow(peaks)) {
    row <- as.double(row)
    # Find time range of the peak to first valley
    i_range <- peaks[row, 2]:peaks[row, 4]
    t <- i_range - peaks[row, 2] + 1
    # Find the data points of the section
    sect <- data[i_range]
    # Create DataFrame with these points
    df <- data.frame(t = t, y = sect)
    # Get exponential fit
    nlc <- nls.control(maxiter = 1000)
    f <- function(x, a, b) {a * exp(b * x)}
    fm_lm <- lm(log(y) ~ t, data=df)
    st <- list(a = exp(coef(fm_lm)[1]), b = coef(fm_lm)[2])
    fit <- try(nls(y ~ f(t, a, b), data = df, start = st, control = nlc))
    # If try block fails, break out of loop and don't return anything
    if (sapply(fit[1], class) == "character") {
      break
    }
    # Get parameters of the fit
    params <- coef(fit)
    # Extract the decay constant value (b)
    alpha <- as.double(params["b"])
    # Get achieved convergence tolerance as metric for accuracy of fit
    # NOTE: R^2 value can be calculated but is not a useful metric
    # for nonlinear models
    conv <- fit$convInfo$finTol
    # Calculate cumulative PM for each plume
    plume_pm <- sum(data[peaks[row, 2]:peaks[row, 4]])
    
    # Add alpha to dataframe
    alphas.newdata <- data.frame(
      "peak_idx" = c(peaks[row, 2]),
      "valley_idx_b" = c(peaks[row, 3]),
      "valley_idx_a" = c(peaks[row, 4]),
      "peak_hgt" = c(peaks[row, 1]),
      "k_val" = c(alpha),
      "conv_tol" = c(conv),
      "cumm_pm" = c(plume_pm)
    )
    alphas.data <- rbind(alphas.data, alphas.newdata)
  }
  alphas.data <- arrange_all(alphas.data)
  alphas.data <- alphas.data[order(alphas.data$peak_hgt), ]
}
```

### Main Plume Analysis

```{r}
# Get data subset
my_df_subset <- my_df[my_df$participant_id=="206" & my_df$environment=="indoor",]

# Separate my_df by before vs. after
my_df_before <- my_df_subset[my_df_subset$case=='before',]
my_df_after  <- my_df_subset[my_df_subset$case=='after',]

# Remove missing data
indices <- which(my_df_before$pm25 > 0)
beforepm25 <- my_df_before$pm25[indices]

indices <- which(my_df_after$pm25 > 0)
afterpm25 <- my_df_after$pm25[indices]


# Get peaks from before and after
peaks_b <- findpeaks(beforepm25,
                     nups = 1,
                     ndowns = 2,
                     minpeakheight = as.numeric(pm_min_heights[pm_min_heights$case=='before', 'mean']),
                     minpeakdistance = 200,
                     threshold = 0)
peaks_a <- findpeaks(afterpm25,
                     nups = 1,
                     ndowns = 2,
                     minpeakheight = as.numeric(pm_min_heights[pm_min_heights$case=='after', 'mean']),
                     minpeakdistance = 200,
                     threshold = 0)
```

```{r}
# Get valleys before and after to isolate each plume
peaks_b <- get_local_valleys(beforepm25,
                             peaks = peaks_b,
                             min_threshold = as.numeric(pm_baseline[pm_baseline$case=='before', 'mean']),
                             min_length = 10)
peaks_a <- get_local_valleys(afterpm25,
                             peaks = peaks_a,
                             min_threshold = as.numeric(pm_baseline[pm_baseline$case=='after', 'mean']),
                             min_length = 10)
```

```{r}

# Plot peaks and valleys on top of current time series with valleys 
# (sanity check)
plot(beforepm25,
  type = "l",
  main = "Indoor/Outdoor Ratio",
  xlab = "Time (indices)",
  ylab = "Ratio of indoor/outdoor concentrations",
  col = "navy")
  grid()
points(peaks_b[, 2], peaks_b[, 1], pch = 20, col = "maroon")
points(peaks_b[, 3], beforepm25[peaks_b[, 3]+1], pch = 20, col = "yellow")
points(peaks_b[, 4], beforepm25[peaks_b[, 4]], pch = 20, col = "green")
```

```{r}
# perform decay analysis
decays_before <- curve_fitting(beforepm25, peaks = peaks_b)
decays_before
```

```{r}
# perform decay analysis
decays_after <- curve_fitting(afterpm25, peaks = peaks_a)
decays_after
```

# Displaying peaks with best fits

```{r}
# Before HEPA deployment
diff_time <- tail(decays_before, n=1)$valley_idx_b
peak_idxs <- c(
  tail(decays_before, n=1)$valley_idx_b - diff_time,
  tail(decays_before, n=1)$valley_idx_a - diff_time
)

p <- my_df_before %>%
  mutate(diff = as.numeric(row.names(my_df_before))-diff_time) %>%
  ggplot(aes(x = diff)) +
  # Plot the highest pm25 plume
  geom_line(aes(y = pm25), color="blue") +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') + 
  ylab('PM2.5 Levels') +
  # ylim(0,500) +
  # Add fun theme
  theme_bw() +
  # Scale to just around the plume
  coord_cartesian(xlim = peak_idxs)
p
```

```{r}
# After HEPA deployment
diff_time <- tail(decays_after, n=1)$valley_idx_b
peak_idxs <- c(
  tail(decays_after, n=1)$valley_idx_b - diff_time,
  tail(decays_after, n=1)$valley_idx_a - diff_time
)
p1 <- my_df_after %>%
  mutate(diff = as.numeric(row.names(my_df_after))-diff_time) %>%
  ggplot(aes(x = diff)) +
  # Plot the highest pm25 plume
  geom_line(aes(y = pm25), color="orange") +
  # Labels
  xlab('Time Since Sensor Deployment (10min)') + 
  ylab('PM2.5 Levels') +
  # ylim(0,500) +
  # Add fun theme
  theme_bw() +
  # Scale to just around the plume
  coord_cartesian(xlim = peak_idxs)
p1
```

```{r}
diff_time_b <- tail(decays_before, n=1)$peak_idx
diff_time_a <- tail(decays_after, n=1)$peak_idx

peak_idxs <- c(
  tail(decays_before, n=1)$valley_idx_b - diff_time_b,
  tail(decays_before, n=1)$valley_idx_a - diff_time_b+20
)

# Both curves back to back
my_df_before <- my_df_before %>%
  mutate(diff = as.numeric(row.names(my_df_before))-diff_time_b+4) %>%
  mutate(HEPA = "off")
my_df_after <- my_df_after %>%
  mutate(diff = as.numeric(row.names(my_df_after))-diff_time_a) %>%
  mutate(HEPA = "on")

my_df_combined <- rbind(my_df_before, my_df_after)

ggplot(my_df_combined, aes(x=diff, y=pm25, color=HEPA)) +
  geom_line() +
  # Labels
  ggtitle("Behavior of Largest PM2.5 Peaks in ID=206") +
  xlab('Time (10min)') + 
  ylab('PM2.5 Levels') +
  # Add fun theme
  theme_bw() +
  # Scale to just around the plume
  coord_cartesian(xlim = peak_idxs)
  
```

```{r}
# Merge decay tables, adding a column indicating whether HEPA installed or not
decays_before$hepa <- FALSE
decays_after$hepa <- TRUE
all_decays <- merge(decays_before, decays_after, all = TRUE)
all_decays

# Get index of when HEPA was installed
hepa_install_idx <- match(TRUE, all_decays$hepa)

# Create summary of decay values for before and after
all_decays["hepa_installed"] <- all_decays["peak_idx"] >= hepa_install_idx
all_decays["p_to_v_peak_width"] <- all_decays["valley_idx_a"] - all_decays["peak_idx"]
all_decays["total_peak_width"] <- all_decays["valley_idx_a"] - all_decays["valley_idx_b"]

decay_sum <-
  all_decays %>%
    group_by(hepa) %>%
    summarize(
      # kval
      k_val_mean = mean(k_val, na.rm = TRUE),
      k_val_median = median(k_val, na.rm = TRUE),
      k_val_min = min(k_val, na.rm = TRUE),
      k_val_max = max(k_val, na.rm = TRUE),
      # peak heights
      peak_hgt_mean = mean(peak_hgt, na.rm = TRUE),
      peak_hgt_median = median(peak_hgt, na.rm = TRUE),
      peak_hgt_min = min(peak_hgt, na.rm = TRUE),
      peak_hgt_max = max(peak_hgt, na.rm = TRUE),
      # p_to_v peak widths
      pv_peak_wdh_mean = mean(p_to_v_peak_width, na.rm = TRUE),
      pv_peak_wdh_median = median(p_to_v_peak_width, na.rm = TRUE),
      pv_peak_wdh_min = min(p_to_v_peak_width, na.rm = TRUE),
      pv_peak_wdh_max = max(p_to_v_peak_width, na.rm = TRUE),
      # total peak widths
      total_peak_wdh_mean = mean(total_peak_width, na.rm = TRUE),
      total_peak_wdh_median = median(total_peak_width, na.rm = TRUE),
      total_peak_wdh_min = min(total_peak_width, na.rm = TRUE),
      total_peak_wdh_max = max(total_peak_width, na.rm = TRUE),
      # cumulative PM
      total_plume_area = sum(cumm_pm, na.rm = TRUE)
    )
decay_sum
```

## Plume Analysis by Participant

The rest of this file calculated statistics on plume reduction
throughout the dataset. Below summarizes all that analysis into one
function, creating reduction statistics, and performing that function on
the dataset as it is grouped by participation and room.

```{r}
get_plumes <- function(before, after, id, pm) {
  # Remove missing data
  indices <- which(my_df_before$pm25 > 0)
  beforepm25 <- my_df_before$pm25[indices]
  
  indices <- which(my_df_after$pm25 > 0)
  afterpm25 <- my_df_after$pm25[indices]
  
  # Calculate average baseline parameter and mean parameters, which will be used
  # to define the baseline and the minimum height of peaks
  pm25_indoor <- graph_main_df[graph_main_df$environment=='indoor' & graph_main_df$particle==pm,]
  pm_baseline <- pm25_indoor %>%
    group_by(case) %>%
    summarise(mean = mean(mean))
  pm_min_heights <- pm25_indoor %>%
    group_by(case) %>%
    summarise(mean = mean(q75))

  # Get peaks from before and after
  peaks_b <- findpeaks(beforepm25,
                       nups = 1,
                       ndowns = 2,
                       minpeakheight = as.numeric(pm_min_heights[pm_min_heights$case=='before', 'mean']),
                       minpeakdistance = 200,
                       threshold = 0)
  peaks_a <- findpeaks(afterpm25,
                       nups = 1,
                       ndowns = 2,
                       minpeakheight = as.numeric(pm_min_heights[pm_min_heights$case=='after', 'mean']),
                       minpeakdistance = 200,
                       threshold = 0)

  # Get valleys before and after to isolate each plume
  peaks_b <- get_local_valleys(beforepm25,
                               peaks = peaks_b,
                               min_threshold = as.numeric(pm_baseline[pm_baseline$case=='before', 'mean']),
                               min_length = 10)
  peaks_a <- get_local_valleys(afterpm25,
                               peaks = peaks_a,
                               min_threshold = as.numeric(pm_baseline[pm_baseline$case=='after', 'mean']),
                               min_length = 10)
  
  # perform decay analysis for before and after
  decays_before <- curve_fitting(beforepm25, peaks = peaks_b)
  decays_after  <- curve_fitting(afterpm25, peaks = peaks_a)
  
  # Merge decay tables, adding a column indicating whether HEPA installed or not
  decays_before$hepa <- FALSE
  decays_after$hepa <- TRUE
  all_decays <- merge(decays_before, decays_after, all = TRUE)
  
  # Add columns of all_decays
  all_decays <- all_decays %>%
    mutate(total_pm = ifelse(hepa, 
                             sum(my_df_after[[pm]]), 
                             sum(my_df_before[[pm]]))) %>%
    mutate(peak_width = valley_idx_a - valley_idx_b) %>%
    mutate(deployment = "IB") %>%
    mutate(room = "residential") %>%
    mutate(participant_id = id) %>%
    mutate(environment = "indoor") %>%
    mutate(pm = pm)
  
  return(data.frame(all_decays[,c(11:15, 1:10)]))
}
```

```{r}
plume_redus <- function(plumes, id, pm) {
  decay_sum <-
    plumes %>%
      group_by(hepa) %>%
      summarize(
        # kval
        k_val_mean = mean(k_val, na.rm = TRUE),
        k_val_median = median(k_val, na.rm = TRUE),
        k_val_min = min(k_val, na.rm = TRUE),
        k_val_max = max(k_val, na.rm = TRUE),
        # peak heights
        peak_hgt_mean = mean(peak_hgt, na.rm = TRUE),
        peak_hgt_median = median(peak_hgt, na.rm = TRUE),
        peak_hgt_min = min(peak_hgt, na.rm = TRUE),
        peak_hgt_max = max(peak_hgt, na.rm = TRUE),
        # total peak widths
        peak_wdh_mean = mean(peak_width, na.rm = TRUE),
        peak_wdh_median = median(peak_width, na.rm = TRUE),
        peak_wdh_min = min(peak_width, na.rm = TRUE),
        peak_wdh_max = max(peak_width, na.rm = TRUE),
        # cumulative PM
        total_plume_area = sum(cumm_pm, na.rm = TRUE),
        # number of plumes
        n = n(),
        # total exposure
        total_exposure = median(total_pm, na.rm = TRUE)
      )
  
  # Add identifying columns
  decay_sum <- decay_sum %>%
    mutate(participant_id = id) %>%
    mutate(environment = "indoor") %>%
    mutate(pm = pm)
  
  return(data.frame(decay_sum[,c(17:19, 1:16)]))
}
```

```{r}
# Create new DataFrame for storing plumes
plumes.data <- data.frame(
  "deployment" = numeric(0),
  "room" = numeric(0),
  "participant_id" = numeric(0),
  "environment" = numeric(0),
  "peak_idx" = numeric(0),
  "valley_idx_b" = numeric(0),
  "valley_idx_a" = numeric(0),
  "peak_hgt" = numeric(0),
  "k_val" = numeric(0),
  "conv_tool" = numeric(0),
  "cumm_pm" = numeric(0),
  "hepa" = numeric(0),
  "hepa_installed" = numeric(0),
  "peak_width" = numeric(0)
)
# Create empty DataFrame for storing k values
plume_reduc.data <- data.frame(
  "participant_id" = numeric(0),
  "k_val_mean" = numeric(0),
  "k_val_median" = numeric(0),
  # "k_val_min" = numeric(0),
  # "k_val_max" = numeric(0),
  "peak_hgt_mean" = numeric(0),
  "peak_hgt_median" = numeric(0),
  "pv_peak_wdh_mean" = numeric(0),
  "pv_peak_wdh_median" = numeric(0),
  "total_peak_wdh_mean" = numeric(0),
  "total_peak_wdh_median" = numeric(0),
  # "peak_wdh_min" = numeric(0),
  # "peak_wdh_max" = numeric(0),
  "total_plume_area" = numeric(0),
  "hepa" = numeric(0)
)

for (id in unique(my_df$participant_id)) {
  for (pm in c("pm1", "pm25", "pm10")) {
    print(paste("Started for participant", id, "at PM type", pm))
    # Get relevant data
    sub_data <- my_df[my_df$participant_id == id & my_df$environment == "indoor",]
    if (length(sub_data$pm25) < 200) next
    
    # Separate sub data by before vs. after
    my_df_before <- sub_data[sub_data$case=='before',]
    my_df_after  <- sub_data[sub_data$case=='after',]
    if (length(my_df_before$pm25) < 100) next
    if (length(my_df_after$pm25) < 100) next
    
    new_plumes <- get_plumes(my_df_before, my_df_after, id, pm)
    new_reducs <- plume_redus(new_plumes, id, pm)
    # Rbind them with previous data
    plumes.data <- rbind(plumes.data, new_plumes)
    plume_reduc.data <- rbind(plume_reduc.data, new_reducs)
  }
}
plumes.data
plume_reduc.data
```

## Save Output
Save 
```{r}
write.csv(plumes.data, file = paste0("summary/Roxbury/IB/", "s_IB_M_allplumes.csv"))
write.csv(plume_reduc.data, file = paste0("summary/Roxbury/IB/", "s_IB_M_plumestats.csv"))
```
