# Roxbury Pie Chart Creation (for Mod-PM)

## STOP

**HAVE YOU RUN `IB_modpm_cleaning`?**
*This file loads a dataframe created from running `IB_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to create pie charts the amount of time chosen for "safe" levels of concentration for multiple participants in the Roxbury pilot, for Mod-PM deployments by Olin.

Results can be found in `summary/Roxbury/IB/s_IB_M_quants.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/Roxbury/IB/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```

## Set Thresholds
```{r}
proportion_df <- main_df  %>% 
    mutate(across(case, ~factor(., levels=c("before","after")))) %>%
    group_by(case) %>%
    # Categorize as low, medium, high
    mutate(danger_pm25 = case_when(
      between(pm25, 0, 5) ~ 'LOW',
      between(pm25, 5, 12) ~ 'MEDIUM',
      pm25 > 12 ~ 'HIGH')) %>%
    # Calculate proportions
    count(danger_pm25) %>%
    mutate(prop = 24 * n / sum(n)) %>% 
    mutate(across(danger_pm25, ~factor(., levels=c("LOW","MEDIUM", "HIGH"))))
```

## Plotting
```{r}
my_caption <- expression(PM[2.5] ~levels ~categorized ~by ~EPA ~standards)

ggplot(proportion_df, aes(x="", y=prop, fill=danger_pm25)) +
  geom_bar(stat="identity", width=1) +
  scale_fill_manual(values = c("#009E73", "#F0E442", "#D55E00")) +
  facet_grid(cols = vars(case),
             labeller = as_labeller(c("before" = "No Purifier", 
                                      "after" = "With Purifier"))) +
  labs(fill = "Pollutant Levels") + 
  theme_minimal() +
  theme(axis.text.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = "Impact of Using Air Purifiers",
       subtitle = "Average time spent per day without and with purifier",
       caption = my_caption)  + 
  geom_text(aes(label = paste(round(prop), "hrs")),
            position = position_stack(vjust = 0.5), size = 3)-> plot_base
  
```

```{r}
 plot_base + xlab("") + coord_polar("y", start=0)
```
```{r}
plot_base + ylab("Percentage of Indoor Air")
```
Below transforms the code so that the x values of the bar chart are the
pollutant levels and the y values are percentages, with two bars dodged for each
category representing before and after.

```{r}
res_df <- dcast(proportion_df, danger_pm25 ~ case)

res_df %>%
  pivot_longer(-danger_pm25, names_to = "variable", values_to = "res") %>%
  ggplot(aes(x = danger_pm25, y = res, fill = factor(variable, levels = c("before", "after")))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
  xlab("Pollutant Level") +
  ylab("Time (Hours)") +
  scale_x_discrete(labels = c("CLEAN", "LOW", "MEDIUM", "HIGH")) +
  scale_fill_discrete(name = "HEPA Installation?",
                      labels = c("no purifier", "with purifier")) +
  theme_minimal() +
  labs(title = "Impact of Using Air Purifiers",
       subtitle = "Average time spent per day without and with purifier",
       caption = my_caption)
```



