# Roxbury Pie Chart Creation (for Mod-PM)

## STOP

**HAVE YOU RUN `IB_modpm_cleaning`?**
*This file loads a dataframe created from running `IB_modpm_cleaning`. Make sure you run that file first (if you haven't already) before running this file.*

This script is used to create pie charts the amount of time chosen for "safe" levels of concentration for multiple participants in the Roxbury pilot, for Mod-PM deployments by Olin.

Results can be found in `summary/Roxbury/IB/s_IB_M_quants.csv`.

## Set up
Load libraries, define file paths, include participant IDs to be processed
```{r}
# Import relevant libraries
library(tidyverse)
```

Set working directory
NOTE: The working directory is dependent by user. 
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/hepa-summer23")
```

```{r}
# Get file path
if (!endsWith(getwd(), "hepa-summer23")) {
  stop("Incorrect working directory")
}
```

```{r}
# Set path to data
path_to_data <- "cleaning/Roxbury/IB/"
```

## Load data
```{r}
# Load Mod-PM data from RData file
load(paste0(path_to_data, "cleaned_modpm.RData"))
```